### 2024-06-28 paper/manuscript.Rmd
initialize.

---
title: |
  | The Promises and Pitfalls of Using Panel Data to Understand
  | Individual Belief Change
author:
  - Stephen Vaisey
  - Turgut Keskintürk
date: "`r format(Sys.time(), '%B %Y')`"
output: pdf_document
bibliography: ./helpers/references.bib
csl: ./helpers/asa.csl
abstract: "In this article, we investigate whether studies on political belief change can identify change trajectories at the individual level. Using simulations and empirical case studies, we propose a grid-search framework that allows researchers to evaluate the extent to which their target estimates generalize to individuals in their study population. We use simulated datasets to estimate plausible values for how many people changed, how much they changed, and who changed, based on observed survey response trajectories. Our results suggest that researchers should think carefully about the conditions under which they can make claims about belief change at the individual level."
fontsize: 11pt
fontfamily: mathpazo
geometry: margin=1in
figsintext: yes
header-includes:
- \usepackage{array}
- \usepackage{longtable}
- \usepackage{booktabs}
- \usepackage{float}
- \usepackage{caption}
- \usepackage{amsmath}
- \captionsetup{justification=raggedright,
                singlelinecheck=false}
---

```{r setup, include=FALSE}
library(tidyverse); library(knitr); library(kableExtra)
knitr::opts_chunk$set(echo = F, 
                      message = F, 
                      warning = F)
knitr::opts_chunk$set(fig.align = "center")
```

Studies on political change at the individual level---change in one's attitudes, values, or preferences---face a challenging problem: while our methods typically provide estimates for change and stability at the group level, our goal is often to understand these processes at the individual level.

This *group-to-person generalizability problem*, the claim that group-level findings may not generalize to each person [@mcmanus2023], has important implications for what we can and can't say about political change across the life-course. While the previous research quantified change with various methods---ranging from assessing the time order of longitudinal observations [@kiley2020; @vaisey2020] to multilevel decomposition of over-time variance [@lersch2023]---these studies have fallen short of disentangling whether they find a small amount of change in a large number of people, or a large amount of change in a small number of people.

This issue arises in a variety of ways in empirical practice, particularly when we want to theorize the mechanisms of political belief change. For example, the General Social Survey (GSS) shows that roughly 35% of the American adults in 2006 believed that "marijuana should be legalized." In less than 10 years, this support had increased to nearly 60%. As seen in Figure 1, this change occurred both *cross-sectionally*, where average support increased with time, and *longitudinally*, where at least some segments of the American population changed their minds on this policy.

```{=latex}
\begin{figure}[!ht]
\begin{center}
```
```{r figure1, out.width = "100%"}
knitr::include_graphics("../figures/grass01.png")
```
```{=latex}
\end{center}
\caption*{Figure 1: The Percent Support for Marijuana Legalization}
\footnotesize{\textit{Notes:} The figure depicts the percent of American adults saying "marijuana should be made legal" in the General Social Survey 2006, 2008, and 2010 panel studies. We included respondents who participated in all three waves of each panel study and weighted the results using GSS post-stratification weights. Note that panel attrition caused compositional changes in the estimates, and compared to the GSS time-series data, the estimates of support are marginally inflated.}
\end{figure}
```

However, these two processes, cross-sectional change and longitudinal change, attract competing theories of political change, and group-level estimates may not properly adjudicate the underlying generative mechanisms. Following the same example, the cross-sectional change in beliefs about marijuana legalization may have occurred via two mechanisms: people may have changed their minds following a changing _zeitgeist_, or old generations who disapproved this policy might have been replaced by new generations approving it. Similarly, the longitudinal change could result from various mechanisms: large segments of the population may have shifted their opinions positively in small amounts, a small segment of the population may have made large changes, or a mix of changes---including both approvals and disapprovals---may have altered the overall balance.

While the cross-sectional processes received strong scholarly attention in the literature [@vaisey2016; @restrepoochoa2024; @bartels2014], questions of the second sort remained largely unexplored. However, given that theories of political ideology involve theoretical expectations about individual-level outcomes, the fact that we lack information about whether our models capture person-level empirical processes is detrimental to understanding adult belief change in the political context. Therefore, our article aims to address this gap by examining whether we can identify (a) *how many people changed*, (b) *how much they changed*, and (c) *who changed*, based on observed survey response trajectories of political beliefs in empirical data.

Using simulations and empirical case studies, we show that it is indeed possible to provide plausible scenarios as to which generative models are more likely to have produced empirical observations, even though a reliable identification of *who* changed occurs only under very narrow conditions. In doing this, we provide a simple grid-search procedure to aid researchers in evaluating the extent to which their target estimates may be generalized to the population of their study. We explore how competing generative models can lead to varying assumptions, emphasizing the importance for researchers to carefully consider the conditions under which claims about personal-level change can be made. Accordingly, we claim that when the meaning of "political change" is elusive and depends too much on how we define change in the first place, our empirical strategies could potentially mislead rather than inform, especially if we overlook the underlying generative mechanisms.

We start with an overview of the "latent variable model" of beliefs and the problems of reliability and resolution in survey response measurement [@alwin2007]. We then introduce our grid-search procedure in the context of three-wave panel data with binary outcomes, examining cases from the GSS on marijuana legalization and abortion. In the next section, we relax the period assumption and investigate how linear trajectories at the individual level can assist us in understanding underlying generative mechanisms in contexts with more than three waves. In the final empirical section, we analyze whether these generative models help us in classification tasks. We conclude by discussing the promises and pitfalls of panel data for understanding individual belief change.

# Measuring Change in Political Beliefs

Alwin's [-@alwin2007] latent-variable approach to survey response is based on the assumption that there is a random variable $Y$, which captures a person $i$'s "true" position in a continuous latent construct. This random variable might stand in for any disposition, including one's political attitudes, values, and preferences, which allows us to define one's political position---at the level of one's "personal culture" [@lizardo2017]---as their *true score* $Y_i$.[^1]

[^1]: Naturally, the assumption that *there is indeed a true score* might miss cases where individuals hold contradictory ideas [@swidler1986] and are ambivalent about their true positions [@zaller1992; @zaller1992a]. In this article, we follow the latent variable model and assume that there *is* a true position for each individual.

The latent-variable approach reveals a fundamental problem in survey research: one's true score $Y_i$ is almost always realized with a certain degree of measurement error. Call this realization the *observed score* $y_i$. We often assume that $y_i$ represents the expected value of hypothetical draws from $Y_i$, with an $\epsilon$ term reflecting the measurement error in this process---i.e., $y_i = Y_i + \epsilon_i$. This error might result from various factors that introduce biases to how we extract information from individuals, given that problems like question construction and interviewer effects might contaminate how individuals reveal their true preferences [@alwin1991]. The inherent stochasticity in this measurement process implies that each draw from $Y_i$ might at least slightly differ from the true score, and this error term is difficult to circumvent with one measure, which is often the case in public opinion surveys.[^2] In this sense, there is often a degree of systematic gap between $Y_i$ and $y_i$.

[^2]: The idea that we should have repeated measures of $Y_i$ is the main intuition behind scale construction [@ansolabehere2008], with the explicit expectation that multiple measures can help us reduce measurement error.

Hence, the main problem is mapping one's true scores to their observed scores, and there are at least two ways we can fail in this context. First, our observed measures might have low *reliability*, such that a hypothetical set of repeated draws may significantly differ from one's true score, due to, e.g., poor question construction or individual idiosyncrasies. Second, survey response categories may be imperfect: the *resolution* of response categories may fail to distinguish consequential differences in people's true scores. In either case, the observed score may not reliably indicate one's true score.

## Separating True Change and Measurement Error in Panel Data

In the panel context, where we observe the same individuals across time, the measurement problem causes substantive questions: if true scores are imperfectly realized and we only have access to one's observed responses, how can we know if a person *actually changed* between time $t$ and $t + 1$, or whether the observed change is merely due to measurement error?

Scholars in public opinion research have attempted to address this issue in multiple ways, using structural equation models that explicitly quantify the measurement error component [@achen1975], strategies that integrate longitudinal surveys to exploit the information in the ordering of responses [@kiley2020], and various multilevel mixed-effects models that shrink responses based on population estimates to derive reliable trends [@lersch2023; @hout2016]. To some extent, all these strategies were fruitful to offset the problems associated with measurement error.[^3] That said, while these empirical strategies were largely successful in accounting for the amount of change *in the population*, they were rather silent about *who* changed in our observations.

[^3]: However, note that the use of repeated observations in the panel context comes with additional assumptions: it is rather straightforward to assume that multiple measures at time $t$ can be indicative of an underlying construct---e.g., we can average them to have an index---but the use of different time periods needs a *stability* assumption, that is, the reliability estimates may only be computed if one assumes that people do not really change between waves, which might confuse true change with measurement error [see, e.g., @hout2016].

Why is this important? After all, if we can produce reasonable estimates at the population level, this must surely be enough. We believe that there are at least two reasons why we should care about this problem. First, the inability to identify the individuals who truly change gives us a false sense of homogeneity. When we report an average change of, say, 0.10 standard deviations, we do not know whether this reflects uniform change across all individuals, or significant changes in some while others remain stable. This is a big substantive difference, yet often overlooked in practice. Second, if we do not know the changers in the first place, our comparisons between groups---either across some pre-determined characteristic, or across treatment status---will average true processes with false negatives and false positives. This is obviously fine if the error occurs randomly, but it is challenging to verify this if individual-level accuracy is compromised. Hence, glossing over this problem discourages us to think about what exactly we are averaging.

## The Question of Individual Belief Change

We can now clearly state our research question: given a longitudinal panel study where we observe some individuals across time, how can we determine the underlying generative process---that is, how many people changed, how much they changed, and who changed---given a level of latent score reliability, response resolution, and a longitudinal survey design?

We pose these questions in a fixed-effects panel context [@allison2009; @vaisey2017], with varying individual trajectories across time [@rüttenauer2020]. This achieves two things. First, it allows us to shift the focus from *distributional change* to *individual change*, given that our theoretical interest is to understand how people change their political beliefs over the life-course. Second, it allows us to model response trajectories, rather than individual changes in response to a shock. Instead, our focus is "classification"---the extent to which we can discriminate those who change from those who do not change in a specific time window.

## A Grid-Search Procedure for Identifying Individual Belief Change

We propose a grid-search procedure to answer our research questions. The idea is straightforward:

(1) Generate simulated datasets with several parameters of interest---rate of change, strength of change, and the direction of change, so that we know the real data-generation process (DGP);
(2) Calculate the distribution of response trajectories in both the observed and simulated data;
(3) Compare the simulated trajectories to the observed trajectories, calculating a distance measure to quantify the extent to which specific simulated values reproduce the observed values;
(4) Replicate this analysis $N$ times for all possible parameter combinations;
(5) Inspect the parameter combinations that most closely approximate to the observed values.

This is simply what Gelman and Hill [-@gelman2007] call a "fake-data simulation," and it is powerful for at least two reasons. First, it allows us to produce reasonable estimates as to what plausible generative models might have produced the real data, ensuring that we can make educated guesses about the extent to which our parameter estimates might generalize to our populations. Second, it is generative, meaning that we can ask specific questions about response trajectories and modeling assumptions.

Our fake-data simulations involve several DGP components: a random variable $Y_{it}$ for true scores, with varying realizations $y_{it}$, representing a person $i$'s beliefs at time $t$; an indicator for change, $D_{it}$, operationalized as a non-reversible trigger if there is a respondent-level switch in true scores; an effect size, $\delta$, and a respondent level direction multiplier $\tau_i$ that indexes whether the change occurs in a positive or a negative direction. Using these components, the DGP follows

$$
y_{it} = Y_i + \delta \tau_i D_{it} + \epsilon_{it} \ \text{ with } \epsilon_{it} \sim \mathcal{N}(0, 1) \tag{DGP}
$$

Note that we allow actors to go through change at any time in the observation window of the study, meaning that $D_{it}$ indexes an actor's change status in their biographical history.

Since our research interest concerns individual trajectories of change, we specify three parameters for these DGPs: (1) *rate of change* refers to the proportion of our sample that goes through change in their political orientations, (2) *strength of change* refers to the standardized mean change in the true score when someone changes---described as Cohen's $d$---and (3) *direction of change* refers to the proportion of changers who change in positive and negative directions. The combination of these three parameters provides a complete picture of population movement in political beliefs.[^4]

[^4]: We define additional parameters that we fix in the case studies presented below: $N$ refers to the number of actors in the fake-data simulation, $T$ refers to the number of waves we have in the panel study, *balance* refers to the observed marginals in the real datasets, *reliability* refers to the item-reliability score regulating the extent to which a true score is realized in the survey context with given level of item reliability [@alwin2007], and *resolution* refers to the number of categories to simplify the latent variable, which is restricted to the binary case in this article.

Using these parameters, we can construct various generative models and compare the simulated trajectories to the observed trajectories to examine how closely certain parameter combinations approximate our empirical observations. In the next three sections, we will apply this procedure in different contexts. In all cases, we restrict our attention to latent variables with binary outcomes to simplify our analyses, but all results are generalizable to cases with higher resolution.[^repo]

[^repo]: The replication files for the article are stored in https://github.com/tkeskinturk/simulating_change

# Study 1: Three-Wave Panel Context

In the three-wave panel context with binary outcomes, the response trajectories are easy to inspect, given that there are only $2^3 = 8$ possible response patterns for each person:

$$
\bigg( \{0,0,0\}, \{0,0,1\}, \{0,1,0\}, \{0,1,1\}, \{1,0,0\}, \{1,0,1\}, \{1,1,0\}, \{1,1,1\} \bigg)
$$

To examine the distribution of these trajectories in the observed data, we can count the number of respondents who go through trajectory $j$. For instance, the marijuana legalization example from Figure 1 has the following contingency table of counts:

$$
\begin{bmatrix}
0, 0, 0 \\
0, 0, 1 \\
0, 1, 0 \\
0, 1, 1 \\
1, 0, 0 \\
1, 0, 1 \\
1, 1, 0 \\
1, 1, 1
\end{bmatrix} 
=
\begin{bmatrix}
869 \\
153 \\
75 \\
115 \\
65 \\
71 \\
54 \\
746 \\
\end{bmatrix}
$$

Approximately 75% of the respondents remain stable, following trajectories of $\{0,0,0\}$ and $\{1,1,1\}$, while others appear to *change* in some respect. The question is the extent to which this contingency table can be approximated by an underlying generative model, where we know the true change parameters.

Let us provide one such example. We simulate a panel dataset, with 20% of the respondents changing their beliefs on marijuana legalization, operationalized as a latent construct with $\mu = 0$ and $\sigma = 1$, by 2 standard deviations in the "negative" direction---note that this is the opposite of the mean pattern observed in Figure 1. We fix all the parameters in the DGP to their observed values: 2,148 individuals observed over 3 periods, with a reliability score of 0.91 [see @hout2016 for the calculation of this value]. We know that the marginals were equal to 0.47 in the entire panel, so we constrain the binary outcomes such that the final distribution of $y$ equals to 0.47 as well. Here are the generated counts for all trajectories, with differences from the original data appended:

$$
\begin{bmatrix}
\\
0, 0, 0 \\
0, 0, 1 \\
0, 1, 0 \\
0, 1, 1 \\
1, 0, 0 \\
1, 0, 1 \\
1, 1, 0 \\
1, 1, 1
\end{bmatrix} 
=
\begin{bmatrix}
\text{Observed} \\
869 \\
153 \\
75 \\
115 \\
65 \\
71 \\
54 \\
746 \\
\end{bmatrix}
-
\begin{bmatrix}
\text{Simulated} \\
810 \\
75 \\
103 \\
57 \\
188 \\
66 \\
148 \\
701 \\
\end{bmatrix}
=
\begin{bmatrix}
\text{Absolute Difference} \\
59 \\
78 \\
28 \\
58 \\
123 \\
5 \\
94 \\
45 \\
\end{bmatrix}
$$

The generative model does a pretty poor job of reproducing the contingency table. 23% of the respondents ($N = 490$) are assigned to the wrong trajectory. This means that we should be able to improve our predictions significantly if we switch to the true data generation model. Unfortunately, we don't know it; the set of true change parameters is the exact thing we want to derive using observed trajectories.

How can we improve our prediction accuracy? Our grid-search procedure allows us to estimate plausible approximation values across a grid of potential generative models. Let us now apply this procedure to the case of marijuana legalization to see whether we can retrieve plausible DGPs.

## The Case of Marijuana Legalization

We first define 2,100 potential generative models using a combination of (1) rate of change, where we vary values from 0 to 1 in increments of 0.05, (2) strength of change, where we vary values from 0 to 2 in increments of 0.1, and (3) five directional possibilities---all negative, all positive, half negative and half positive, 75% negative and 25% positive, and 75% positive and 25% negative. This results in $21 \times 20 \times 5 =$ 2,100 cells, each representing one potential data-generation process that might have produced the observed data. We then generate corresponding panel datasets based on each model. Due to the stochastic nature of our fake-data simulations, we repeat this procedure 1,000 times for each cell, resulting in 2,100,000 datasets for which we evaluate the observed trajectories.

Once we have our fake-data simulations, we count the number of respondents assigned to each response trajectory, calculate the absolute errors relative to the observed data, and normalize these counts by dividing the error by the total number of respondents. We then average these percent errors across 1,000 simulations to obtain stable estimates. Figure 2 presents the results.

```{=latex}
\begin{figure}[!ht]
\begin{center}
```
```{r figure2, out.width = "100%"}
knitr::include_graphics("../figures/grass02.png")
```
```{=latex}
\end{center}
\caption*{Figure 2: Fake-Data Simulations for Marijuana Legalization}
\footnotesize{\textit{Notes:} The figure depicts the extent to which specific parameter combinations approximate the observed data. Each point represents our distance measure, calculated as the sum of absolute errors between observed and simulated contingency tables, where darker colors represent smaller distances. We used 5\%, 10\%, and 20\% distance cuts for color shades, along with a second filter using quantiles of 0.05, 0.10, and 0.20 at each level. The shading thus reflects both absolute and relative accuracy among the simulations. The single yellow point represents the parameter combination that best approximates the observed data in the General Social Survey.}
\end{figure}
```

There are several notable findings. First, error values range from 4.6% to 69.6%, with an interquartile range of 10.6% and 24%. This indicates that many cases fail to achieve a reasonable approximation, but we still end up with reliably close generative models. Second, the best value occurs when the rate of change equals to 25%, with a positive strength of 0.9 standard deviation change (SD). That said, note the black values clustered around the yellow cell in the bottom right panel, where cells range from 25% to 35% rate of change, with strength ranging, once again, from 0.7 to 1 SD. These patterns suggest that the most plausible data generation process occurs when roughly 2 to 4 out of 10 people positively change their approvals by roughly 1 SD. Finally, as expected, there are no plausible estimates where change occurs in the negative direction, providing a plausibility check for our proposed method.

These findings demonstrate the utility of our approach. By iterating over DGPs, we can identify plausible values that may have produced the observed change trajectories in the panel context.

## The Case of the Legality of Abortion

What about cases where the direction or even existence of change is less clear-cut? We now turn our attention to the legality of abortion debate, a contentious issue characterized by high disagreement and high political sorting in the American context [@baldassarri2008; @hout2022]. In 2006, 2008, and 2010 GSS panels, respondents agreed or disagreed with the statement: "it should be possible for a pregnant woman to obtain a legal abortion if the woman wants it for any reason." Figure 3 shows response trajectories for this question across three panels, indicating a clear stasis in the marginals.

```{=latex}
\begin{figure}[!ht]
\begin{center}
```
```{r figure3, out.width = "100%"}
knitr::include_graphics("../figures/abany01.png")
```
```{=latex}
\end{center}
\caption*{Figure 3: The Percent Support for Abortion}
\footnotesize{\textit{Notes:} The figure depicts the percent of American adults saying "it should be possible for a pregnant woman to obtain a legal abortion if the woman wants it for any reason" in the General Social Survey 2006, 2008, and 2010 panel studies. We included respondents who participated in all three waves of each panel study and weighted the results using GSS post-stratification weights. Note that panel attrition caused compositional changes in the estimates, and compared to the GSS time-series data, the estimates of support are marginally inflated.}
\end{figure}
```

However, the contingency table for trajectories shows more heterogeneity. Again, there is considerable stability---approximately 73% of respondents remain stable, following trajectories of $\{0,0,0\}$ and $\{1,1,1\}$, similar to the case of marijuana legalization. But there is more homogeneity in terms of directional change: while trajectories like $\{0,0,1\}$ and $\{0,1,1\}$, cases indicating an upward trend, were 2.27 times more likely in the marijuana legalization case, we now observe similar numbers in the abortion case---230 clear upward trends and 209 clear downward trends. This suggests that the directional heterogeneity of political change might have cancelled the trends in the aggregate, resulting in stable marginals in the observed frequencies.

$$
\begin{bmatrix}
0, 0, 0 \\
0, 0, 1 \\
0, 1, 0 \\
0, 1, 1 \\
1, 0, 0 \\
1, 0, 1 \\
1, 1, 0 \\
1, 1, 1
\end{bmatrix} 
=
\begin{bmatrix}
972 \\
115 \\
101 \\
115 \\
118 \\
99 \\
91 \\
733 \\
\end{bmatrix}
$$

Figure 4 presents the results from our grid-search exercise, implemented across 2,100,000 datasets with 2,100 generative models. We set the reliability to .86 following @hout2016. This time, the estimates are much closer to the observed frequencies, with error values ranging from 3% to 64.5%, and an interquartile range of 5.1% and 23.5%. That said, there is strong variability in the plausible generative scenarios. Note how, in the top right corner, there is a plausibility bound ranging from a rate of change as low as 10% to as high as 100%, and strength of change ranging from as low as 0.3 SD to as high as 2 SD, indicating a variety of scenarios. In fact, the top 10 generative models out of 2,100 DGPs have error rates within a margin of 0.04%, meaning that it is difficult to distinguish the predictive abilities of these models.

```{=latex}
\begin{figure}[!ht]
\begin{center}
```
```{r figure4, out.width = "100%"}
knitr::include_graphics("../figures/abany02.png")
```
```{=latex}
\end{center}
\caption*{Figure 4: Fake-Data Simulations for Abortion}
\footnotesize{\textit{Notes:} The figure depicts the extent to which specific parameter combinations approximate the observed data. Each point represents our distance measure, calculated as the sum of absolute errors between observed and simulated contingency tables, where darker colors represent smaller distances. We used 5\%, 10\%, and 20\% distance cuts for color shades, along with a second filter using quantiles of 0.05, 0.10, and 0.20 at each level. The shading thus reflects both absolute and relative accuracy among the simulations. The single yellow point represents the parameter combination that best approximates the observed data in the General Social Survey.}
\end{figure}
```

This exercise is an appropriate illustration of the limitations of panel studies. Based on the observed data, we cannot determine whether the observed marginals result from a small number of people making substantial belief changes or a large number of people making minor political changes. This, ultimately, boils down to the issue of observation frequency: with three-wave panel data and reasonable doubts about the accuracy of every single survey response due to reliability and resolution concerns, it is challenging to adjudicate the underlying generative models. In the next section, we relax this assumption and extend our method to cases with more than three waves.

# Study 2: Longer Panels and Linear Trajectories

Once we relax the three-wave assumption, the immediate problem is that the contingency tables will be sparse and unwieldy. Think, for instance, a panel study with 1,000 people conducted over 10 waves. Assuming binary outcomes, there are $2^{10} =$ 1,024 potential trajectories---more than the number of individuals in the study. That said, having 10 waves instead of 3 is more desirable for one simple reason: while we had to be economical in our method and count every response very seriously, we now have the opportunity to regularize a person's response trajectories, resulting in more reasonable *trends* for each individual that are less vulnerable to reliability issues.

There are multiple approaches for achieving this with panel data.[^5] We use a fixed-effects design and fit individual linear slopes of the outcome on a normalized time trajectory [@rüttenauer2020]. That is, we simply do a separate regression for each individual and use their observed slope. We apply this method to the case of immigration spending.

[^5]: Two obvious alternatives are multilevel mixed-effects models with varying intercepts and slopes [@lersch2023], which allow for shrinkage in response trajectories using partial pooling, and latent growth curve models [@proust-lima2012], which enable us to postulate latent trajectories at the individual level. Both approaches are, of course, plausible. However, the shrinkage factor in the multilevel models is far too consequential in the simulation approach when the underlying generative model has too few respondents changing, as it would place excessive emphasis on the population mean. Similarly, latent modeling with simulated grid cells results in too many researcher degrees of freedom. Therefore, we settled on the fixed effects individual model as a reasonable approach to linear trend modeling.

## The Case of Immigration Spending

The data for this exercise comes from the *Political Psychology Data from a 26-Wave Yearlong Longitudinal Study* [@brandt2021], conducted biweekly in 2019-2020 for a year. In one question in the survey, respondents were asked whether they agree with the policy of increased federal spending on immigration control. We binarized the responses to this question such that those agreeing with the statement received a 1, and those who said the spending should be decreased or kept the same received a 0. We then dropped respondents with missing data and trimmed the periods to pre-COVID windows, resulting in 213 individuals observed 17 times over 34 weeks. Although the sample for this case study is obviously not representative of any national population, it provides an intensive set of repeated observations at the individual level.

We follow the same grid-search procedure, extending our approach to cases involving linear slopes. Figure 1 provides a schematic example of our method. First, we fit linear models to the observed data, normalizing time to be between 0 and 1, and retrieve individual slopes of change. The top-left panel in Figure 1 documents the distribution of these slopes across 213 individuals. Next, we build a simulated dataset and perform the same procedure to extract individual slopes. The bottom-left panel in Figure 1 shows one such instantiation, where we build a dataset with a fixed rate of change of 25%, a strength of change of 2 SD, and a 50%--50% split in directional change, with individual slopes for 213 simulated respondents. In the final step, we compare these two distributions using a two-sample Kolmogorov-Smirnov (KS) statistic to quantify the distance between the observed data and the simulated data from one generated model. The right panel in Figure 1 documents the empirical CDFs of these two distributions, which we use in our calculations of the KS statistics.

```{=latex}
\begin{figure}[!ht]
\begin{center}
```
```{r figure5, out.width = "100%"}
knitr::include_graphics("../figures/immigration01.png")
```
```{=latex}
\end{center}
\caption*{Figure 5: The Distribution of Linear Trajectories}
\footnotesize{\textit{Notes:} The figure shows linear trajectories from the first 17-waves of the \textit{Political Psychology Data from a 26-Wave Yearlong Longitudinal Study} (26W), covering the pre-COVID period of the study. The top-left panel depicts the distribution of individual slopes from varying fixed effects models, where a set of respondents agreed---or not---with the statement, "federal spending to control immigration should be increased." We included respondents who participated in all 17 waves (N = 213). The bottom-left panel depicts the distribution of individual slopes from a simulated dataset, with a fixed rate of change of 25\%, a strength of change of 2 SD, and a 50\%--50\% split in directional change. The right panel shows the empirical cumulative distribution functions from each of these distributions.}
\end{figure}
```

Figure 6 presents the results for this procedure, implemented across, once again, 2,100,000 datasets with 2,100 generative models. As can be seen in the figure, there is a clear cluster of results in a specific region of the grid. KS statistics range from 4.9% to 47.6%, with an interquartile range of 11% and 18.2%. The plausibility bounds indicate that between 50% to 80% of respondents may have made a change of roughly 0.5 SD, resulting in a mean decline in support for increased immigration control spending. Consistent with our baseline expectation of a mean decline, the plausible scenarios involve true changers making their changes in a negative direction.

```{=latex}
\begin{figure}[!ht]
\begin{center}
```
```{r figure6, out.width = "100%"}
knitr::include_graphics("../figures/immigration02.png")
```
```{=latex}
\end{center}
\caption*{Figure 6: Fake-Data Simulations for Immigration}
\footnotesize{\textit{Notes:} The figure depicts the extent to which specific parameter combinations approximate the observed data. Each point represents our distance measure, calculated as the Kilmogorov-Smirnov statistics of the empirical CDFs of observed and simulated slopes, where darker colors represent smaller distances. We used 5\%, 10\%, and 20\% distance cuts for color shades, along with a second filter using quantiles of 0.05, 0.10, and 0.20 at each level. The shading thus reflects both absolute and relative accuracy among the simulations. The single yellow point represents the parameter combination that best approximates the observed data in the 26W Study.}
\end{figure}
```

With this exercise, we extended our grid-search procedure to cases with more than three waves of panel data, but the same procedure is applicable to the three-wave case as well, provided that the analyst believes it is appropriate to use linear slopes estimated from three observations. The underlying logic, however, is the same: as long as we have comparable trajectories in the simulated data relative to our observed empirical data, we can scale up the process and perform algorithmic searches.[^6]

[^6]: Of course, one such procedure may involve a lazy algorithm that goes through a targeted search, where each iteration involves more focused and narrow bands of parameters. This is vulnerable to get stuck in local optima, however.

# Study 3: Detection of True Changers

So far, we have presented simulation evidence for recovering plausible generative models (DGPs) that may have produced the observed empirical data. But what about *classifying individuals* into those who change and those who do not? Suppose that, using the grid-search procedure, we happen to determine the true generative model for our empirical observations. Can we also recover an accurate classification of changers and non-changers at the individual level? That is, given a vector of predictions, can we measure our success in identifying the true changers [@tharwat2020]?

To address this question, we conduct another simulation exercise. Similar to earlier simulations, we generate $N$ datasets, this time using parameters fixed to the best-case values identified through our grid-search procedure. In each of these datasets, we estimate linear regressions at the individual level with time scaled from 0 to 1---effectively allowing us to extract a model-predicted measure of change for each person. We use a straightforward heuristic to decide whether a simulated person changed: if the linear slope is closer to the strength-of-change parameter than it is to 0, the respondent gets classified as a *changer*, otherwise, they are classified as a *non-changer*. We then cross-tabulate these predictions with the real change indicator used during the generation of the data, constructing a "confusion matrix" that compares true change classifications with observed change classifications.

In the binary, three-wave case, these classifications are straightforward. Only the 001, 011, 100, and 110 cells can be classified as changers because only they can have non-zero slopes---in fact, all have slopes of 1 or -1. In cases where the most plausible DGP is one directional, like for the marijuana question, only two of those cells would be classified as changers. The situation is more complicated for a more continuous case, like that of immigration in the 26-wave data. Many different slopes are possible. In those cases, classifying people as closer to "changers" or "non-changers" relies on choosing a cutoff, as we do here.

In either case, this binary classification allows us to assess how "accurate" we are in a single simulation run. Of course, there are multiple ways to conceptualize accuracy, but for our purposes, we focus on two: *sensitivity* and *specificity*. The former measures the probability that a true changer will get classified as a changer; the latter is the probability that a true non-changer will be classified as a non-changer. These are the formulas for each:

$$
\begin{aligned}
\text{Sensitivity} &= \frac{\text{True Positive}}{\text{True Positive + False Negative}} & \text{Specificity} &= \frac{\text{True Negative}}{\text{True Negative + False Positive}}
\end{aligned}
$$

We extract the best-case scenarios from our three empirical case studies---marijuana legalization, the legality of abortion, and immigration spending---and perform the classification task across 1,000 fake-data simulations. We then calculate the sensitivity and specificity scores in each run.

A concrete example will help clarify what these numbers mean. The mean sensitivity for the abortion question is .205. This means that only 21% of those who actually changed would be correctly classified as changers, even with full knowledge of the data generating process. This is due to several factors, including the shortness of the panel, imperfect reliability---though .86 is quite good---and the low level of response resolution, which is binary.

For sensitivity, the number is higher at .828. This means that 83% of the non-changers are correctly classified. This is largely due to the marginal distribution of change.

The distributions of sensitivity and specificity for all three analyses are presented in Figure 7. It is clear that our general ability to identify true changers is quite low, whereas the identification of true non-changers is rather better. Across simulations, sensitivity ranges between 17% and 54%, while specificity scores range between 74% and 95%. That said, there are also subtle differences among the case studies. Sensitivity scores are consistently lower in the legality of abortion case, reflecting our earlier discussion about how the plausible set of scenarios was much higher in that case. At the same time, the spread of simulation scores is rather high in the case of immigration spending for both sensitivity and specificity, suggesting that linear trajectories in longer panels might be rather susceptible to randomness in our analyses.

```{=latex}
\begin{figure}[!ht]
\begin{center}
```
```{r figure7, out.width = "100%"}
knitr::include_graphics("../figures/detection.png")
```
```{=latex}
\end{center}
\caption*{Figure 7: Sensitivity and Specificity Across Case Studies}
\footnotesize{\textit{Notes:} The figure depicts the sensitivity---true positives divided by the sum of true positives and false positives---and specificity---true negatives divided by the sum of true negatives and false negatives---scores across three studies.}
\end{figure}
```

These results suggest that, while the grid-search procedure allows us to identify plausible generative models that may have generated the empirical patterns, detecting true change at the level of individuals occurs under only very narrow conditions. The fact that linear trends with longer panels performed better, on average, than the three-wave cases indicates that having longer panels might be essential to dampen the negative effects of response reliability on our estimates. This finding emphasizes the importance of collecting panel data over more extended periods to increase the robustness of our proposed DGPs, and the accuracy of identifying true changers in the population.

# Discussion and Conclusions

In this article, we examined the identification of belief change in panel context. Using simulations and empirical case studies, we showed that a grid-search procedure that iterates over generative models can help us identify plausible data generation processes that may have produced our empirical observations, though the identification of individual cases occurs under narrow conditions. We argued that the *group-to-person generalizability problem* is an impediment to a robust understanding of political belief change, and scholars should understand the extent to which their target estimates can generalize to the population of their study. As such, we proposed a simulation approach that takes response generation seriously, which may help us move forward in addressing this problem.

This article has several implications for the study of belief change across the life-course. First, our findings suggest that recent debates about the existence or non-existence of belief change in adult populations [@kiley2020; @lersch2023] should tackle the heterogeneous nature of change and stability in individual preferences since group-level measures may hamper our understanding of individual mechanisms. Instead, scholars need to specify the generative models that aggregate average estimates imply. Analyses that examine population patterns over time allow us to adjudicate whether compositional changes or individual changes drive cumulative political change [@bartels2014; @vaisey2016]. Similarly, we need robust and reliable panel studies to decompose whether individual changes occur across all individuals---implying period changes---or in specific groups of people---implying political heterogeneity.

Second, the ambivalence in response trajectories and the observation that linear trend models may perform better with longer panels indicate that classical solutions to the measurement problem---ranging from scale construction to structural equation models and principal component analyses [@judd1980; @ansolabehere2008; @alwin1991]---should have an important place in panel designs, as even moderate reliability scores can significantly hinder our ability to specify plausible generative processes. We highlighted that a fundamental challenge in the study of belief change is distinguishing whether an observed change is due to true change or measurement error, and reliability and resolution are of utmost importance to resolve this problem.

One theoretical assumption of this article was to rely on a latent variable model, and avoid questions about contradictory considerations in human cognition that mitigate stable and coherent opinion formation [@converse1964; @zaller1992a]. This issue, of course, introduces another layer to the problem of belief change and measurement error. We relied on a very strong assumption that an individual has a central tendency when it comes to political issues, rather than being functionally "random" in their responses to survey questions. However, a careful examination of adult belief change should be grounded in a more theoretically-sound study of belief formation.

Third, our finding that individual classifications of change status are largely ineffective means that future research needs to be more sensitive to the detection of trends in individual survey responses over time. As mentioned earlier, varying modeling strategies, ranging from multilevel mixed-effects models that rely on partial pooling to various forms of latent mixture models, could improve the reliability of responses at the individual level. We believe that the use of these modeling strategies, along with longer panel designs, could provide important corrections to the study of belief change.

Finally, these results call for granular panel surveys that go beyond the classical 2-wave and 3-wave designs, as reliability concerns significantly limit the plausible conclusions we can derive from these shorter panels. In a 2-wave setup, where there is only one change, we can never properly establish whether a change we observe is due to measurement error or it is indeed a genuine shift, making extended panel designs necessary. The differences among our case studies, along with the sensitivity and specificity tests, suggest that panel studies with shorter designs are largely not suited for trend modeling. Since we care about generative models rather than one observed sample, strategies that enable us to reliably estimate target quantities are warranted and essential to make progress on longstanding questions about belief change.

# References

### 2024-06-28 scripts/1_grass.R

@@ -1,0 +1,263 @@
+
+# - promises and pitfalls of panel change ------------------------------------ #
+# - case study 1 # grass ----------------------------------------------------- #
+
+### note: this script
+###       (a) takes `grass` as an examplary GSS item,
+###       (b) simulates potential DGPs approximating the observed parameters,
+###       (c) shows the relative fit of the DGPs to observed data.
+
+rm(list = ls()) # clean-up
+pacman::p_load(
+  gssr,
+  ggh4x,
+  purrr,
+  furrr,
+  hrbrthemes,
+  srvyr,
+  tidyverse)
+theme_set(theme_ipsum_rc())
+
+# --- helpers
+my_oka <-
+  c("#eebd64",
+    "#8b3b36", 
+    "#a2b4b7")
+`%nin%` = Negate(`%in%`)
+
+# --- fire power
+source("./scripts/helper_functions/build_DGP.R")
+source("./scripts/helper_functions/classification_metrics.R")
+
+# PART 1: GSS DATA ----------------------------------------------------------- #
+
+data("gss_panel06_long")
+data("gss_panel08_long")
+data("gss_panel10_long")
+
+# --- get data
+gss <-
+  ## gather
+  bind_rows(
+    gss_panel06_long |>
+      select(pid = firstid, year, wave, wtpannr123, grass) |>
+      haven::zap_labels() |>
+      mutate(pid = as.integer(paste0(
+        2006, as.integer(pid)
+      ))) |> 
+      mutate(run = "2006-2010"),
+    gss_panel08_long |>
+      select(pid = firstid, year, wave, wtpannr123, grass) |>
+      haven::zap_labels() |>
+      mutate(pid = as.integer(paste0(
+        2008, as.integer(pid)
+      ))) |> 
+      mutate(run = "2008-2012"),
+    gss_panel10_long |>
+      select(pid = firstid, year, wave, wtpannr123, grass) |>
+      haven::zap_labels() |>
+      mutate(pid = as.integer(paste0(
+        2010, as.integer(pid)
+      ))) |> 
+      mutate(run = "2010-2014")
+  ) |>
+  ## organize
+  drop_na() |>
+  mutate(n = n(), .by = pid) |> filter(n == 3) |> select(-n) |>
+  mutate(grass = ifelse(grass == 2, 0, 1))
+
+# --- draw trajectories
+png(
+  "figures/grass01.png",
+  w = 7,
+  h = 5,
+  units = "in",
+  res = 500
+)
+gss |>
+  ## weight
[…]

### 2024-06-28 scripts/2_abortion.R

@@ -1,0 +1,263 @@
+
+# - promises and pitfalls of panel change ------------------------------------ #
+# - case study 2 # abortion -------------------------------------------------- #
+
+### note: this script
+###       (a) takes `abany` as an examplary GSS item,
+###       (b) simulates potential DGPs approximating the observed parameters,
+###       (c) shows the relative fit of the DGPs to observed data.
+
+rm(list = ls()) # clean-up
+pacman::p_load(
+  gssr,
+  ggh4x,
+  purrr,
+  furrr,
+  hrbrthemes,
+  srvyr,
+  tidyverse)
+theme_set(theme_ipsum_rc())
+
+# --- helpers
+my_oka <-
+  c("#eebd64",
+    "#8b3b36", 
+    "#a2b4b7")
+`%nin%` = Negate(`%in%`)
+
+# --- fire power
+source("./scripts/helper_functions/build_DGP.R")
+source("./scripts/helper_functions/classification_metrics.R")
+
+# PART 1: GSS DATA ----------------------------------------------------------- #
+
+data("gss_panel06_long")
+data("gss_panel08_long")
+data("gss_panel10_long")
+
+# --- get data
+gss <-
+  ## gather
+  bind_rows(
+    gss_panel06_long |>
+      select(pid = firstid, year, wave, wtpannr123, abany) |>
+      haven::zap_labels() |>
+      mutate(pid = as.integer(paste0(
+        2006, as.integer(pid)
+      ))) |> 
+      mutate(run = "2006-2010"),
+    gss_panel08_long |>
+      select(pid = firstid, year, wave, wtpannr123, abany) |>
+      haven::zap_labels() |>
+      mutate(pid = as.integer(paste0(
+        2008, as.integer(pid)
+      ))) |> 
+      mutate(run = "2008-2012"),
+    gss_panel10_long |>
+      select(pid = firstid, year, wave, wtpannr123, abany) |>
+      haven::zap_labels() |>
+      mutate(pid = as.integer(paste0(
+        2010, as.integer(pid)
+      ))) |> 
+      mutate(run = "2010-2014")
+  ) |>
+  ## organize
+  drop_na() |>
+  mutate(n = n(), .by = pid) |> filter(n == 3) |> select(-n) |>
+  mutate(abany = ifelse(abany == 2, 0, 1))
+
+# --- draw trajectories
+png(
+  "figures/abany01.png",
+  w = 7,
+  h = 5,
+  units = "in",
+  res = 500
+)
+gss |>
+  ## weight
[…]

### 2024-06-28 scripts/3_immig.R

@@ -1,0 +1,297 @@
+
+# - promises and pitfalls of panel change ------------------------------------ #
+# - case study 3 # immigration ----------------------------------------------- #
+
+### note: this script
+###       (a) takes `imm` as an examplary panel data item,
+###       (b) simulates potential DGP distributions using fixed-effects models,
+###       (c) shows the relative fit of the DGPs to observed data.
+
+rm(list = ls()) # clean-up
+pacman::p_load(
+  ggh4x,
+  purrr,
+  furrr,
+  patchwork,
+  hrbrthemes,
+  tidyverse)
+theme_set(theme_ipsum_rc())
+
+# --- helpers
+my_oka <-
+  c("#eebd64",
+    "#8b3b36", 
+    "#a2b4b7")
+`%nin%` = Negate(`%in%`)
+
+# --- fire power
+source("./scripts/helper_functions/build_DGP.R")
+source("./scripts/helper_functions/classification_metrics.R")
+
+# PART 1: 26-WAVE DATA ------------------------------------------------------- #
+
+# --- read in
+d <- read_csv("./data/yllanon.csv")
+
+# --- cleaning up
+d <- d |>
+  ## organize
+  mutate(pid = cur_group_id(), .by = "id") |>
+  ## chop off post-COVID periods
+  filter(wave < 18) |>
+  ## simplify
+  select(pid, t = wave, y = imm) |>
+  ## clean-up the outcome
+  filter(y %nin% c(8, 9)) |> drop_na() |>
+  ## fully balanced
+  mutate(n = n(), .by = "pid") |> filter(n == 17) |>
+  ## scale the variables
+  mutate(y = ifelse(y > 4, 1, 0)) |>
+  mutate(y_std = (y - mean(y)) / sd(y)) |>
+  mutate(t = scales::rescale(t)) |>
+  ## select and wrap up organization
+  select(pid, t, y, y_std) |> arrange(pid, t)
+
+# --- reference
+d_slopes <- d |>
+  nest(.by = "pid") |>
+  mutate(m  = map(
+    .x = data, 
+    .f = ~ lm(y_std ~ t, data = .))) |>
+  mutate(lm = map_dbl(
+    .x = m, 
+    .f = ~ round(coef(.)[[2]], 3)))  |>
+  select(pid, lm)
+
+# PART 2: ILLUSTRATION ------------------------------------------------------- #
+
+set.seed(112358)
+d_example <- generate_data(
[…]

### 2024-06-28 scripts/4_detection.R

@@ -1,0 +1,443 @@
+
+# - promises and pitfalls of panel change ------------------------------------ #
+# - change detection --------------------------------------------------------- #
+
+### note: this script
+###       (a) takes `grass`, `abany`, and `imm` as variable examples,
+###       (b) simulates potential DGPs approximating the observed parameters,
+###       (c) assesses whether we have leverage to detect individuals.
+
+rm(list = ls()) # clean-up
+pacman::p_load(
+  gssr,
+  ggh4x,
+  purrr,
+  furrr,
+  hrbrthemes,
+  srvyr,
+  patchwork,
+  tidyverse)
+theme_set(theme_ipsum_rc())
+
+# --- helpers
+my_oka <-
+  c("#eebd64",
+    "#8b3b36", 
+    "#a2b4b7")
+`%nin%` = Negate(`%in%`)
+
+# --- fire power
+source("./scripts/helper_functions/build_DGP.R")
+source("./scripts/helper_functions/classification_metrics.R")
+
+# --- prep
+plan(multisession)
+set.seed(112358)
+
+# PART 1: DATA --------------------------------------------------------------- #
+
+data("gss_panel06_long")
+data("gss_panel08_long")
+data("gss_panel10_long")
+
+# --- grass
+grass <-
+  ## gather
+  bind_rows(
+    gss_panel06_long |>
+      select(pid = firstid, year, wave, wtpannr123, grass) |>
+      haven::zap_labels() |>
+      mutate(pid = as.integer(paste0(
+        2006, as.integer(pid)
+      ))) |> 
+      mutate(run = "2006-2010"),
+    gss_panel08_long |>
+      select(pid = firstid, year, wave, wtpannr123, grass) |>
+      haven::zap_labels() |>
+      mutate(pid = as.integer(paste0(
+        2008, as.integer(pid)
+      ))) |> 
+      mutate(run = "2008-2012"),
+    gss_panel10_long |>
+      select(pid = firstid, year, wave, wtpannr123, grass) |>
+      haven::zap_labels() |>
+      mutate(pid = as.integer(paste0(
+        2010, as.integer(pid)
+      ))) |> 
+      mutate(run = "2010-2014")
+  ) |>
+  ## organize
+  drop_na() |>
+  mutate(n = n(), .by = pid) |> filter(n == 3) |> select(-n) |>
+  mutate(grass = ifelse(grass == 2, 0, 1))
+
+# --- abany
+abany <-
+  ## gather
+  bind_rows(
+    gss_panel06_long |>
[…]

### 2024-06-28 scripts/helper_functions/build_DGP.R

@@ -1,0 +1,252 @@
+
+# - promises and pitfalls of panel change ------------------------------------ #
+# - helper: build DGP -------------------------------------------------------- #
+
+### note: this script 
+###       (a) generates panel data from specified parameters values,
+###       (b) estimates user-specified models of change,
+###       (c) stores (1) patterns, (2) model results OR (3) data, if specified.
+
+require(tidyverse)
+
+# PART 1: FUNCTION ----------------------------------------------------------- #
+
+generate_data <-
+  
+  function(
+    
+    # ----------------------------------------------------- #
+    # FUNCTION ARGUMENTS AND WARNINGS                       #
+    # ----------------------------------------------------- #
+    
+    ## the number of Ns
+    n = 1000,
+    ## the number of Ts
+    t = 3,
+    ## the rate of change in the population
+    rate = 0.25,
+    ## % balance in changes
+    balance_dir = 0.5,
+    ## % balance in `y` = 1
+    balance_res = 0.5,
+    ## strength of change in latent scores
+    strength = 1,
+    ## reliability for latent scores
+    reliable = 0.8,
+    ## return options for components
+    export = TRUE, patterns = TRUE, slopes = TRUE) {
+    
+    if(n <= 0)
+      return(stop(
+        "Error: We need at least *some* people. Check your `n` call."))
+    if(t <= 1)
+      return(stop(
+        "Error: We need at least 2 time periods to generate panel data. Try again."))
+    if(rate < 0)
+      return(stop(
+        "Error: Rate of change must be 0 or a positive number."))
+    if(balance_dir <  0 | balance_dir >  1)
+      return(stop(
+        "Error: Directionality balance must be a number between 0 and 1."))
+    if(balance_res <= 0 | balance_res >= 1)
+      return(stop(
+        "Error: Outcome balance must be between 0 and 1 (not inclusive)."))
+    if(strength < 0)
+      return(stop(
+        "Error: Change strength can't be negative (remember: it's 'strength')"))
+    if(reliable < 0)
+      return(stop(
+        "Error: Reliability can't be negative. Try again."))
+    if(export == FALSE & patterns == FALSE & slopes == FALSE)
+      return(stop(
+        "You said no data, no patterns, or no model information. What do you want?"))
+    
+    # ----------------------------------------------------- #
+    # BUILDING BLOCKS                                       #
+    # ----------------------------------------------------- #
[…]

### 2024-06-28 scripts/helper_functions/classification_metrics.R

@@ -1,0 +1,65 @@
+
+# - promises and pitfalls of panel change ------------------------------------ #
+# - helper: classification metrics ------------------------------------------- #
+
+### note: this script
+###       (a) takes a confusion matrix of true/simulated counts as its input,
+###       (b) calculates kappa, sensitivity, and specificity.
+
+require(tidyverse)
+
+# FUNCTION ------------------------------------------------------------------- #
+
+confused_outcomes <-
+  
+  function(confusion_matrix) {
+    
+    # --- calculate kappa
+    
+    # accuracy
+    p_0 =
+      (confusion_matrix[4, 3] 
+       + 
+       confusion_matrix[1, 3]
+      ) / 
+      sum(confusion_matrix$n)
+    
+    # expected classification
+    p_e =
+      (
+        (
+        (confusion_matrix[1, 3] + confusion_matrix[2, 3])
+        *
+        (confusion_matrix[1, 3] + confusion_matrix[3, 3])
+        )
+      +
+        (
+        (confusion_matrix[3, 3] + confusion_matrix[4, 3])
+        *
+        (confusion_matrix[2, 3] + confusion_matrix[4, 3])
+        )
+      ) /
+      (sum(confusion_matrix$n) ^ 2)
+    
+    # kappa score
+    kappa = (p_0 - p_e) / (1 - p_e)
+    
+    # --- calculate sensitivity
+    sensitivity =
+      confusion_matrix[4, 3] /
+      (confusion_matrix[4, 3] + confusion_matrix[3, 3])
+    
+    # --- calculate specificity
+    specificity =
+      confusion_matrix[1, 3] /
+      (confusion_matrix[1, 3] + confusion_matrix[2, 3])
+    
+    # --- export
+    output <- list(kappa = kappa$n,
+                   sensitivity = sensitivity$n,
+                   specificity = specificity$n)
+    
+    return(output)
+  }
+
+# ---------------------------------------------------------------------------- #

### 2024-11-23 paper/manuscript.Rmd
initial revisions for the R&R.

@@
 First, the inability to identify the individuals who truly change gives us a false sense of homogeneity.
-When we report an average change of, say, 0.10 standard deviations, we do not know whether this reflects uniform change across all individuals, or significant changes in some while others remain stable.
+When we report an average change of, say, 0.10 standard deviations, we do not know whether this reflects uniform change in latent positions across all individuals, or significant changes in some while others remain stable.
 This is a big substantive difference, yet often overlooked in practice.
-Second, if we do not know the changers in the first place, our comparisons between groups---either across some pre-determined characteristic, or across treatment status---will average true processes with false negatives and false positives.
+Second, if we cannot identify the "changers" in the first place, our comparisons between groups---either across some pre-determined characteristic, or across treatment status---will average true processes with false negatives and false positives.
 This is obviously fine if the error occurs randomly, but it is challenging to verify this if individual-level accuracy is compromised.
@@
 
-We can now clearly state our research question: given a longitudinal panel study where we observe some individuals across time, how can we determine the underlying generative process---that is, how many people changed, how much they changed, and who changed---given a level of latent score reliability, response resolution, and a longitudinal survey design?
+We can now clearly state our research question: given a longitudinal panel study where we observe some individuals across time, how can we determine the underlying generative process---that is, how many people changed, how much their latent positions changed, and who changed---given a level of latent score reliability, response resolution, and a longitudinal survey design?
 

### 2024-11-23 paper/revision_memo.Rmd
initial revisions for the R&R.

---
title: |
  | \LARGE The Promises and Pitfalls of Using Panel Data to Understand
  | \LARGE Individual Belief Change
  |
  | \large Revision Memo
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  pdf_document:
    number_sections: true
bibliography: ./helpers/references.bib
csl: ./helpers/asa.csl
fontsize: 11pt
fontfamily: mathpazo
geometry: margin=1in
figsintext: yes
header-includes:
- \usepackage{array}
- \usepackage{longtable}
- \usepackage{booktabs}
- \usepackage{float}
- \usepackage{caption}
- \usepackage{amsmath}
- \captionsetup{justification=raggedright,
                singlelinecheck=false}
---

```{r setup, include=FALSE}
library(tidyverse); library(knitr); library(kableExtra)
knitr::opts_chunk$set(echo = F, 
                      message = F, 
                      warning = F)
knitr::opts_chunk$set(fig.align = "center")
```

We would like to thank the reviewers and the editors for their excellent and thorough comments. We especially appreciated the reviewers’ clear suggestions that allowed us to revise major portions of the first manuscript. In the following pages, we outline the changes we made in the paper.

# Substantive Issues

## The Take-Home Message of the Article

Both the reviewers and the editors urged us to clarify the take-home message of our article.

Particularly, R1 noted that:

> Readers may be uncertain of how to use the authors’ findings.

Similarly, R3 noted that:

> I remained puzzled if this [measurement error] is an issue the proposed model solves or the authors are just conscious of its existence.

We appreciated this feedback and made substantial rewrites in the body of our manuscript to clarify what the goal of the analyses is and why researchers should implement our method.

In a nutshell, we would like to note that we do not claim to have solved the measurement error problem. Rather, our aim is much more modest: we want to provide applied researchers a tool that will allow them to see what plausible DGPs may have generated their data. This might be thought of as something akin to a power analysis for theoretical claims. In the context of this special issue, this means that we urge researchers to think more carefully about how different compositional claims---a combination of rate of change, strength of change, and the mix of directional change---may be responsible for the same data, such that our claims about individual preferences might have different implications depending on different DGPs.

As we detail below, we now provide an R package that researchers can seamlessly implement in their own work to perform this analysis, which we hope will clarify our message.

## The Discussion of the Group-to-Person Generalizability Problem

Reviewers suggested that an extended evaluation of the group-to-person generalizability problem is warranted in the manuscript to highlight the importance of the measurement problem.

Particularly, R1 noted that:

> Examples of recent or important theories of individual-level change might highlight the tension between the pervasiveness of this kind of theorizing with the measurement challenge outlined in the paper.

Similarly, R3 noted that:

> I would encourage the authors to formalize the issue already in the introduction and highlight it with examples.

We extended our discussions with these suggestions. Particularly, we now walk through an example of how average coefficients represent a combination of our three quantities of interest.

## The Item Composition

R1 notes that

> Do many---or many politically important---items look like the marijuana legalization item or the abortion item?

Similarly, the editor notes that

> The panel data were collected over a relatively short time span. Does this matter? It may possibly make it harder to detect belief change.

We agree with these concerns. We did two changes to improve upon these points:

- We added a new item using the *British Elections Study*'s panel study, generalized trust, an item collected between 2019 and 2024 across 6 years and 5 waves, in a potentially disruptive time period due to COVID-19. As a more general note, we believe that the tool we propose should apply in all cases, but we agree that a more thorough set of examples are helpful.
- Additionally, we now present a simulation study that applies our protocol in different DGPs and assesses the conditions under which we can make better guesses about the true DGPs (see below for more details). We think this new section provides important answers about the questions of generalizability, panel span, varying DGP compositions, and related issues.

# Issues on Analytic Strategy and Implementation

## The Clarity of the Grid Search Discussions

R3 notes that

> "the five bullet point explanation of the Grid-Search Procedure is inadequate."

We implemented three solutions for this issue. First, we now introduce our grid-search algorithm as an R package, `gridsearch`, which will accompany to this article. As we detail in the paper, we revised our procedure and substantially increased its efficiency, which means that it will be easy for people to implement our procedure in their own projects. Second, we now include a full replication package for the analyses presented in this project as a footnote in the article. We were planning to do so along with publication, but for a paper of this kind, it is beneficial for the reviewers to see the full code, so we attached a view-only (anonymized) link in the manuscript. Finally, we made substantive rewrites to the manuscript to make our protocol as clear and concise as possible. We hope that the R package, the replication repo, and these rewrites make our strategy more clear.

## Simulations and Validations

As R2 noted, our grid search protocol will perform better in some cases than others---e.g., having few people changing their beliefs in the panel context might make it hard to find a plausible DGP under low reliability, or having a mixture of positive and negative changes might complicate our search approach. To address these issues, we conducted two simulation studies, as we described in a new section in the manuscript. These simulation studies allowed us to validate our procedure, while at the same time showing the conditions under which we may expect high uncertainty from the data. These studies strengthened our paper greatly and we thank R2 for their great suggestion.

## The Detection of True Changers in the Sample

R2 notes that

> "One thing I was concerned about in the end is that the third question---who changed---seems not have been responded to adequately."

We agree. That said, we believe that the reason for this is *substantive* rather than *methodological*. As stated above, our goal is to provide researchers a general tool to assess plausible DGPs and the conditions under which they can make claims about (1) compositional distributions such as rate of change and strength of change and (2) their own ability to detect "true changers." We are less interested in providing particular solutions to these problems, as they are context-specific.

This is particularly important when it comes to R3's comments on the estimator choice:

> "The solutions I am familiar with to the heterogeneity of effects problem involve mixture models that allow the different mixtures to highlight different effects in population sub-groups [...] if the proposed model works better than these, I would have expected to see explicit comparisons or at least direct contrasts. Such models are only passingly mentioned in the conclusion. A fixed effects model explain these away entirely while the two modeling approaches discarded allow for their modeling."

We agree. This issue is not a concern for the grid search algorithm, where our central objective is to generate a distance score between real data and simulated data, but it is indeed a concern when it comes to providing classification scores for whether we can recover *true changers*. As R3 stated, the use of varying individual fixed effects is one choice among many, and there are many alternatives.

While we agree with this point, we have two reasons for refraining from these extensions. First, these models proved problematic when it comes to scaling. Remember that we perform $N$ simulations for detecting true changers, and map the true changers we know from the DGP to the observed ones we recovered from our estimation. In the context of population-wide growth curve models, partial pooling causes a strong shrinkage in the estimates that, in turn, increases the researcher degrees of freedom as to how we can classify individuals---this is not a concern in the fixed-effects case, where the estimates have a basic decision rule with a benchmark using the strength parameter in the DGP. This is easier to circumvent in the context of mixture models, where the aim is to classify in the first place. That said, we found that the choice of model parameters affects these classifications in panels with low number of $T$s, which is why we decided to stick to the fixed-effects method.

We now have an extended discussion about these issues in the `Discussion` section and we provide specific suggestions for further research in this area. We thank R3 for their keen observation.

# Issues on Presentation and Results

## The Size of Changes

R1 notes:

> I believe the authors are referring to change in latent positions, but this should be made explicit.

That is true. We always separate the latent positions and the observed survey positions. However, we agree that the discussion of the change scores was not sufficiently clear in the manuscript. We thus changed the relevant sections to emphasize this difference. We thank the R1 for this warning.

## Reliability and Resolution

R1 notes:

> Varying parameters relating to reliability and resolution in simulations might be useful, given the prominence of these things in the conclusion section.

We have two comments regarding this suggestion:

(1) *Reliability*: We agree with R1 that an exploration of reliability is warranted. To provide a more complete picture of how "successful" we can be as a function of different reliability scores, we conducted simulation analyses. The manuscript now includes a section on how real reliability and assumed reliability in the grid search interact to produce different error values.
(2) *Resolution*: One important weakness of our current manuscript is that we restrict our attention to the binary cases. The generalization of our procedure to cases with more than two response categories, however, requires a substantial addition to our protocol---an extension that we are currently working. We included an extensive discussion about this issue in the `Discussion` section, urging future research about this issue. We thank R1 for their suggestion.

## The Grid Search Plots

The editor notes:

> I would have also liked a bit more explanation on how to read the fake data simulation graphs.

In response to this, we revised our plots and extended our discussions about how to interpret them.

---

Thank you all for this intellectual exchange, and we greatly appreciate your deep engagement with our article. We hope that the revisions we made adequately responded to your concerns.

Regards,

The Authors

### 2024-11-23 scripts/1_grass.R

@@ -1,263 +1,0 @@
-
-# - promises and pitfalls of panel change ------------------------------------ #
-# - case study 1 # grass ----------------------------------------------------- #
-
-### note: this script
-###       (a) takes `grass` as an examplary GSS item,
-###       (b) simulates potential DGPs approximating the observed parameters,
-###       (c) shows the relative fit of the DGPs to observed data.
-
-rm(list = ls()) # clean-up
-pacman::p_load(
-  gssr,
-  ggh4x,
-  purrr,
-  furrr,
-  hrbrthemes,
-  srvyr,
-  tidyverse)
-theme_set(theme_ipsum_rc())
-
-# --- helpers
-my_oka <-
-  c("#eebd64",
-    "#8b3b36", 
-    "#a2b4b7")
-`%nin%` = Negate(`%in%`)
-
-# --- fire power
-source("./scripts/helper_functions/build_DGP.R")
-source("./scripts/helper_functions/classification_metrics.R")
-
-# PART 1: GSS DATA ----------------------------------------------------------- #
-
-data("gss_panel06_long")
-data("gss_panel08_long")
-data("gss_panel10_long")
-
-# --- get data
-gss <-
-  ## gather
-  bind_rows(
-    gss_panel06_long |>
-      select(pid = firstid, year, wave, wtpannr123, grass) |>
-      haven::zap_labels() |>
-      mutate(pid = as.integer(paste0(
-        2006, as.integer(pid)
-      ))) |> 
-      mutate(run = "2006-2010"),
-    gss_panel08_long |>
-      select(pid = firstid, year, wave, wtpannr123, grass) |>
-      haven::zap_labels() |>
-      mutate(pid = as.integer(paste0(
-        2008, as.integer(pid)
-      ))) |> 
-      mutate(run = "2008-2012"),
-    gss_panel10_long |>
-      select(pid = firstid, year, wave, wtpannr123, grass) |>
-      haven::zap_labels() |>
-      mutate(pid = as.integer(paste0(
-        2010, as.integer(pid)
-      ))) |> 
-      mutate(run = "2010-2014")
-  ) |>
-  ## organize
-  drop_na() |>
-  mutate(n = n(), .by = pid) |> filter(n == 3) |> select(-n) |>
-  mutate(grass = ifelse(grass == 2, 0, 1))
-
-# --- draw trajectories
-png(
-  "figures/grass01.png",
-  w = 7,
-  h = 5,
-  units = "in",
-  res = 500
-)
-gss |>
-  ## weight
[…]

### 2024-11-23 scripts/2_abortion.R

@@ -1,263 +1,0 @@
-
-# - promises and pitfalls of panel change ------------------------------------ #
-# - case study 2 # abortion -------------------------------------------------- #
-
-### note: this script
-###       (a) takes `abany` as an examplary GSS item,
-###       (b) simulates potential DGPs approximating the observed parameters,
-###       (c) shows the relative fit of the DGPs to observed data.
-
-rm(list = ls()) # clean-up
-pacman::p_load(
-  gssr,
-  ggh4x,
-  purrr,
-  furrr,
-  hrbrthemes,
-  srvyr,
-  tidyverse)
-theme_set(theme_ipsum_rc())
-
-# --- helpers
-my_oka <-
-  c("#eebd64",
-    "#8b3b36", 
-    "#a2b4b7")
-`%nin%` = Negate(`%in%`)
-
-# --- fire power
-source("./scripts/helper_functions/build_DGP.R")
-source("./scripts/helper_functions/classification_metrics.R")
-
-# PART 1: GSS DATA ----------------------------------------------------------- #
-
-data("gss_panel06_long")
-data("gss_panel08_long")
-data("gss_panel10_long")
-
-# --- get data
-gss <-
-  ## gather
-  bind_rows(
-    gss_panel06_long |>
-      select(pid = firstid, year, wave, wtpannr123, abany) |>
-      haven::zap_labels() |>
-      mutate(pid = as.integer(paste0(
-        2006, as.integer(pid)
-      ))) |> 
-      mutate(run = "2006-2010"),
-    gss_panel08_long |>
-      select(pid = firstid, year, wave, wtpannr123, abany) |>
-      haven::zap_labels() |>
-      mutate(pid = as.integer(paste0(
-        2008, as.integer(pid)
-      ))) |> 
-      mutate(run = "2008-2012"),
-    gss_panel10_long |>
-      select(pid = firstid, year, wave, wtpannr123, abany) |>
-      haven::zap_labels() |>
-      mutate(pid = as.integer(paste0(
-        2010, as.integer(pid)
-      ))) |> 
-      mutate(run = "2010-2014")
-  ) |>
-  ## organize
-  drop_na() |>
-  mutate(n = n(), .by = pid) |> filter(n == 3) |> select(-n) |>
-  mutate(abany = ifelse(abany == 2, 0, 1))
-
-# --- draw trajectories
-png(
-  "figures/abany01.png",
-  w = 7,
-  h = 5,
-  units = "in",
-  res = 500
-)
-gss |>
-  ## weight
[…]

### 2024-11-23 scripts/3_grass.R

@@ -1,0 +1,141 @@
+
+# - promises and pitfalls of panel change ------------------------------------ #
+# - case 1 # grass ----------------------------------------------------------- #
+
+### note: this script
+###       (a) takes `grass` as an examplary GSS item,
+###       (b) simulates potential DGPs approximating the observed parameters,
+###       (c) shows the relative fit of the DGPs to observed data.
+
+rm(list = ls()) # clean-up
+
+# install.packages("devtools")
+# devtools::install_github("tkeskinturk/gridsearch")
+
+pacman::p_load(
+  gssr,
+  purrr,
+  furrr,
+  srvyr,
+  hrbrthemes,
+  gridsearch,
+  tidyverse)
+theme_set(theme_ipsum_rc())
+
+# --- helpers
+my_oka <-
+  c("#eebd64",
+    "#8b3b36", 
+    "#a2b4b7")
+`%nin%` = Negate(`%in%`)
+
+# PART 1: GSS DATA ----------------------------------------------------------- #
+
+data("gss_panel06_long")
+data("gss_panel08_long")
+data("gss_panel10_long")
+
+# --- get data
+gss <-
+  bind_rows(
+    gss_panel06_long |>
+      select(pid = firstid, year, wave, wtpannr123, grass) |>
+      haven::zap_labels() |>
+      mutate(pid = as.integer(paste0(
+        2006, as.integer(pid)
+      ))) |>
+      mutate(run = "2006-2010"),
+    gss_panel08_long |>
+      select(pid = firstid, year, wave, wtpannr123, grass) |>
+      haven::zap_labels() |>
+      mutate(pid = as.integer(paste0(
+        2008, as.integer(pid)
+      ))) |>
+      mutate(run = "2008-2012"),
+    gss_panel10_long |>
+      select(pid = firstid, year, wave, wtpannr123, grass) |>
+      haven::zap_labels() |>
+      mutate(pid = as.integer(paste0(
+        2010, as.integer(pid)
+      ))) |>
+      mutate(run = "2010-2014")
+  )
+
+# --- clean-up
+gss <- gss |>
+  mutate(grass = as.character(grass)) |>
+  mutate(grass = case_when(grass == "Legal" ~ 1,
+                           grass == "Not Legal" ~ 0,
+                           .default = NA_real_)) |>
+  drop_na() |>
[…]

### 2024-11-23 scripts/3_immig.R

@@ -1,297 +1,0 @@
-
-# - promises and pitfalls of panel change ------------------------------------ #
-# - case study 3 # immigration ----------------------------------------------- #
-
-### note: this script
-###       (a) takes `imm` as an examplary panel data item,
-###       (b) simulates potential DGP distributions using fixed-effects models,
-###       (c) shows the relative fit of the DGPs to observed data.
-
-rm(list = ls()) # clean-up
-pacman::p_load(
-  ggh4x,
-  purrr,
-  furrr,
-  patchwork,
-  hrbrthemes,
-  tidyverse)
-theme_set(theme_ipsum_rc())
-
-# --- helpers
-my_oka <-
-  c("#eebd64",
-    "#8b3b36", 
-    "#a2b4b7")
-`%nin%` = Negate(`%in%`)
-
-# --- fire power
-source("./scripts/helper_functions/build_DGP.R")
-source("./scripts/helper_functions/classification_metrics.R")
-
-# PART 1: 26-WAVE DATA ------------------------------------------------------- #
-
-# --- read in
-d <- read_csv("./data/yllanon.csv")
-
-# --- cleaning up
-d <- d |>
-  ## organize
-  mutate(pid = cur_group_id(), .by = "id") |>
-  ## chop off post-COVID periods
-  filter(wave < 18) |>
-  ## simplify
-  select(pid, t = wave, y = imm) |>
-  ## clean-up the outcome
-  filter(y %nin% c(8, 9)) |> drop_na() |>
-  ## fully balanced
-  mutate(n = n(), .by = "pid") |> filter(n == 17) |>
-  ## scale the variables
-  mutate(y = ifelse(y > 4, 1, 0)) |>
-  mutate(y_std = (y - mean(y)) / sd(y)) |>
-  mutate(t = scales::rescale(t)) |>
-  ## select and wrap up organization
-  select(pid, t, y, y_std) |> arrange(pid, t)
-
-# --- reference
-d_slopes <- d |>
-  nest(.by = "pid") |>
-  mutate(m  = map(
-    .x = data, 
-    .f = ~ lm(y_std ~ t, data = .))) |>
-  mutate(lm = map_dbl(
-    .x = m, 
-    .f = ~ round(coef(.)[[2]], 3)))  |>
-  select(pid, lm)
-
-# PART 2: ILLUSTRATION ------------------------------------------------------- #
-
-set.seed(112358)
-d_example <- generate_data(
[…]

### 2024-11-23 scripts/4_abortion.R

@@ -1,0 +1,141 @@
+
+# - promises and pitfalls of panel change ------------------------------------ #
+# - case 2 # abortion -------------------------------------------------------- #
+
+### note: this script
+###       (a) takes `abany` as an examplary GSS item,
+###       (b) simulates potential DGPs approximating the observed parameters,
+###       (c) shows the relative fit of the DGPs to observed data.
+
+rm(list = ls()) # clean-up
+
+# install.packages("devtools")
+# devtools::install_github("tkeskinturk/gridsearch")
+
+pacman::p_load(
+  gssr,
+  purrr,
+  furrr,
+  srvyr,
+  hrbrthemes,
+  gridsearch,
+  tidyverse)
+theme_set(theme_ipsum_rc())
+
+# --- helpers
+my_oka <-
+  c("#eebd64",
+    "#8b3b36", 
+    "#a2b4b7")
+`%nin%` = Negate(`%in%`)
+
+# PART 1: GSS DATA ----------------------------------------------------------- #
+
+data("gss_panel06_long")
+data("gss_panel08_long")
+data("gss_panel10_long")
+
+# --- get data
+gss <-
+  bind_rows(
+    gss_panel06_long |>
+      select(pid = firstid, year, wave, wtpannr123, abany) |>
+      haven::zap_labels() |>
+      mutate(pid = as.integer(paste0(
+        2006, as.integer(pid)
+      ))) |> 
+      mutate(run = "2006-2010"),
+    gss_panel08_long |>
+      select(pid = firstid, year, wave, wtpannr123, abany) |>
+      haven::zap_labels() |>
+      mutate(pid = as.integer(paste0(
+        2008, as.integer(pid)
+      ))) |> 
+      mutate(run = "2008-2012"),
+    gss_panel10_long |>
+      select(pid = firstid, year, wave, wtpannr123, abany) |>
+      haven::zap_labels() |>
+      mutate(pid = as.integer(paste0(
+        2010, as.integer(pid)
+      ))) |> 
+      mutate(run = "2010-2014")
+  ) 
+
+# --- clean-up
+gss <- gss |>
+  mutate(abany = as.character(abany)) |>
+  mutate(abany = case_when(abany == "Yes" ~ 1,
+                           abany == "No" ~ 0,
+                           .default = NA_real_)) |>
+  drop_na() |>
[…]

### 2024-11-23 scripts/4_detection.R

@@ -1,443 +1,0 @@
-
-# - promises and pitfalls of panel change ------------------------------------ #
-# - change detection --------------------------------------------------------- #
-
-### note: this script
-###       (a) takes `grass`, `abany`, and `imm` as variable examples,
-###       (b) simulates potential DGPs approximating the observed parameters,
-###       (c) assesses whether we have leverage to detect individuals.
-
-rm(list = ls()) # clean-up
-pacman::p_load(
-  gssr,
-  ggh4x,
-  purrr,
-  furrr,
-  hrbrthemes,
-  srvyr,
-  patchwork,
-  tidyverse)
-theme_set(theme_ipsum_rc())
-
-# --- helpers
-my_oka <-
-  c("#eebd64",
-    "#8b3b36", 
-    "#a2b4b7")
-`%nin%` = Negate(`%in%`)
-
-# --- fire power
-source("./scripts/helper_functions/build_DGP.R")
-source("./scripts/helper_functions/classification_metrics.R")
-
-# --- prep
-plan(multisession)
-set.seed(112358)
-
-# PART 1: DATA --------------------------------------------------------------- #
-
-data("gss_panel06_long")
-data("gss_panel08_long")
-data("gss_panel10_long")
-
-# --- grass
-grass <-
-  ## gather
-  bind_rows(
-    gss_panel06_long |>
-      select(pid = firstid, year, wave, wtpannr123, grass) |>
-      haven::zap_labels() |>
-      mutate(pid = as.integer(paste0(
-        2006, as.integer(pid)
-      ))) |> 
-      mutate(run = "2006-2010"),
-    gss_panel08_long |>
-      select(pid = firstid, year, wave, wtpannr123, grass) |>
-      haven::zap_labels() |>
-      mutate(pid = as.integer(paste0(
-        2008, as.integer(pid)
-      ))) |> 
-      mutate(run = "2008-2012"),
-    gss_panel10_long |>
-      select(pid = firstid, year, wave, wtpannr123, grass) |>
-      haven::zap_labels() |>
-      mutate(pid = as.integer(paste0(
-        2010, as.integer(pid)
-      ))) |> 
-      mutate(run = "2010-2014")
-  ) |>
-  ## organize
-  drop_na() |>
-  mutate(n = n(), .by = pid) |> filter(n == 3) |> select(-n) |>
-  mutate(grass = ifelse(grass == 2, 0, 1))
-
-# --- abany
-abany <-
-  ## gather
-  bind_rows(
-    gss_panel06_long |>
[…]

### 2024-11-23 scripts/5_trust.R

@@ -1,0 +1,130 @@
+
+# - promises and pitfalls of panel change ------------------------------------ #
+# - case 3 # trust ----------------------------------------------------------- #
+
+### note: this script
+###       (a) takes `trust` as an examplary BES item,
+###       (b) simulates potential DGPs approximating the observed parameters,
+###       (c) shows the relative fit of the DGPs to observed data.
+
+rm(list = ls()) # clean-up
+
+# install.packages("devtools")
+# devtools::install_github("tkeskinturk/gridsearch")
+
+pacman::p_load(
+  purrr,
+  furrr,
+  srvyr,
+  hrbrthemes,
+  gridsearch,
+  tidyverse)
+theme_set(theme_ipsum_rc())
+
+# --- helpers
+my_oka <-
+  c("#eebd64",
+    "#8b3b36", 
+    "#a2b4b7")
+`%nin%` = Negate(`%in%`)
+
+# PART 1: BES DATA ----------------------------------------------------------- #
+
+# --- get data
+data <-
+  haven::read_dta("./data/britishelection.dta") |>
+  select(
+    pid = id,
+    trust_1 = genTrustW17,
+    trust_2 = genTrustW20,
+    trust_3 = genTrustW21,
+    trust_4 = genTrustW23,
+    trust_5 = genTrustW27
+  ) |>
+  haven::zap_labels() |>
+  mutate_at(
+    .vars = vars(matches("trust")),
+    .funs = ~ case_when(. == 1 ~ 1, . == 2 ~ 0, TRUE ~ NA_real_)
+  ) |>
+  drop_na() |>
+  pivot_longer(cols = -pid,
+               names_to = "var",
+               values_to = "y") |>
+  separate(var, into = c("var", "t")) |>
+  mutate(t = as.integer(t)) |> select(-var) |> 
+  mutate(
+    year =
+      case_when(
+        t == 1 ~ 2019,
+        t == 2 ~ 2020,
+        t == 3 ~ 2021,
+        t == 4 ~ 2022,
+        t == 5 ~ 2024
+      )
+  )
+
+# --- draw trajectories
+png(
+  "figures/trust01.png",
+  w = 7,
+  h = 5,
+  units = "in",
+  res = 500
+)
+glm(y ~ factor(year), data = data, family = "binomial") |>
+  ggeffects::ggemmeans(terms = "year") |>
+  as_tibble() |>
+  ## draw
+  ggplot(aes(
+    x = x,
[…]

### 2024-11-23 scripts/6_immigration.R

@@ -1,0 +1,179 @@
+
+# - promises and pitfalls of panel change ------------------------------------ #
+# - case 4 # immigration ----------------------------------------------------- #
+
+### note: this script
+###       (a) takes `imm` as an examplary panel data item,
+###       (b) simulates potential DGP distributions using fixed-effects models,
+###       (c) shows the relative fit of the DGPs to observed data.
+
+rm(list = ls()) # clean-up
+
+# install.packages("devtools")
+# devtools::install_github("tkeskinturk/gridsearch")
+
+pacman::p_load(
+  purrr,
+  furrr,
+  srvyr,
+  hrbrthemes,
+  gridsearch,
+  tidyverse,
+  patchwork)
+theme_set(theme_ipsum_rc())
+
+# --- helpers
+my_oka <-
+  c("#eebd64",
+    "#8b3b36", 
+    "#a2b4b7")
+`%nin%` = Negate(`%in%`)
+
+# PART 1: 26-WAVE DATA ------------------------------------------------------- #
+
+# --- read in
+data <- read_csv("./data/yllanon.csv")
+
+# --- cleaning up
+data <- data |>
+  ## organize
+  mutate(pid = cur_group_id(), .by = "id") |>
+  ## chop off post-COVID periods
+  filter(wave < 18) |>
+  ## simplify
+  select(pid, t = wave, y = imm) |>
+  ## clean-up the outcome
+  filter(y %nin% c(8, 9)) |> drop_na() |>
+  ## fully balanced
+  mutate(n = n(), .by = "pid") |> filter(n == 17) |>
+  ## scale the variables
+  mutate(y = ifelse(y > 4, 1, 0)) |>
+  mutate(y_std = (y - mean(y)) / sd(y)) |>
+  mutate(t = scales::rescale(t)) |>
+  ## select and wrap up organization
+  select(pid, t, y, y_std) |> arrange(pid, t)
+
+# --- reference
+d_slopes <- data |>
+  nest(.by = "pid") |>
+  mutate(m  = map(
+    .x = data, 
+    .f = ~ lm(y_std ~ t, data = .))) |>
+  mutate(lm = map_dbl(
+    .x = m, 
+    .f = ~ round(coef(.)[[2]], 3)))  |>
+  select(pid, lm)
+
+# PART 2: ILLUSTRATION ------------------------------------------------------- #
+
+set.seed(112358)
+d_example <- buildDGP(
[…]

### 2024-11-23 scripts/7_detection.R

@@ -1,0 +1,147 @@
+
+# - promises and pitfalls of panel change ------------------------------------ #
+# - change detection --------------------------------------------------------- #
+
+### note: this script
+###       (a) takes `grass`, `abany`, and `imm` as variable examples,
+###       (b) simulates potential DGPs approximating the observed parameters,
+###       (c) assesses whether we have leverage to detect individuals.
+
+rm(list = ls()) # clean-up
+
+# devtools (install if necessary!)
+# devtools::install_github("tkeskinturk/gridsearch")
+
+pacman::p_load(
+  purrr,
+  furrr,
+  hrbrthemes,
+  gridsearch,
+  tidyverse,
+  patchwork)
+theme_set(theme_ipsum_rc())
+
+# --- helpers
+my_oka <-
+  c("#eebd64",
+    "#8b3b36", 
+    "#a2b4b7",
+    "#4c6337")
+`%nin%` = Negate(`%in%`)
+
+# PART 1: DATA --------------------------------------------------------------- #
+
+grass <- readRDS("./data/output/grass.rds")
+abany <- readRDS("./data/output/abany.rds")
+trust <- readRDS("./data/output/trust.rds")
+immig <- readRDS("./data/output/immigration.rds")
+
+grass <- grass |> arrange(error) |> slice(1)
+abany <- abany |> arrange(error) |> slice(1)
+trust <- trust |> arrange(error) |> slice(1)
+immig <- immig |> arrange(error) |> slice(1)
+
+# --- grass
+grass_sim <-
+  simulateChangers(
+    n = grass |> pull(n),
+    t = 3,
+    rate = grass |> pull(rate),
+    balance_dir = 1,
+    balance_res = 0.47,
+    strength = grass |> pull(strength),
+    reliable = 0.91
+  )
+
+# --- abany
+abany_sim <-
+  simulateChangers(
+    n = abany |> pull(n),
+    t = 3,
+    rate = abany |> pull(rate), ## for help
+    balance_dir = 1,
+    balance_res = 0.45,
+    strength = abany |> pull(strength),
+    reliable = 0.86
+  )
+
+# --- trust
+trust_sim <-
+  simulateChangers(
+    n = trust |> pull(n),
+    t = 5,
+    rate = trust |> pull(rate),
+    balance_dir = 0.5,
+    balance_res = 0.43,
+    strength = trust |> pull(strength),
+    reliable = 0.88
+  )
+
+# --- imm
+immig_sim <-
+  simulateChangers(
[…]

### 2024-11-23 scripts/helper_functions/build_DGP.R

@@ -1,252 +1,0 @@
-
-# - promises and pitfalls of panel change ------------------------------------ #
-# - helper: build DGP -------------------------------------------------------- #
-
-### note: this script 
-###       (a) generates panel data from specified parameters values,
-###       (b) estimates user-specified models of change,
-###       (c) stores (1) patterns, (2) model results OR (3) data, if specified.
-
-require(tidyverse)
-
-# PART 1: FUNCTION ----------------------------------------------------------- #
-
-generate_data <-
-  
-  function(
-    
-    # ----------------------------------------------------- #
-    # FUNCTION ARGUMENTS AND WARNINGS                       #
-    # ----------------------------------------------------- #
-    
-    ## the number of Ns
-    n = 1000,
-    ## the number of Ts
-    t = 3,
-    ## the rate of change in the population
-    rate = 0.25,
-    ## % balance in changes
-    balance_dir = 0.5,
-    ## % balance in `y` = 1
-    balance_res = 0.5,
-    ## strength of change in latent scores
-    strength = 1,
-    ## reliability for latent scores
-    reliable = 0.8,
-    ## return options for components
-    export = TRUE, patterns = TRUE, slopes = TRUE) {
-    
-    if(n <= 0)
-      return(stop(
-        "Error: We need at least *some* people. Check your `n` call."))
-    if(t <= 1)
-      return(stop(
-        "Error: We need at least 2 time periods to generate panel data. Try again."))
-    if(rate < 0)
-      return(stop(
-        "Error: Rate of change must be 0 or a positive number."))
-    if(balance_dir <  0 | balance_dir >  1)
-      return(stop(
-        "Error: Directionality balance must be a number between 0 and 1."))
-    if(balance_res <= 0 | balance_res >= 1)
-      return(stop(
-        "Error: Outcome balance must be between 0 and 1 (not inclusive)."))
-    if(strength < 0)
-      return(stop(
-        "Error: Change strength can't be negative (remember: it's 'strength')"))
-    if(reliable < 0)
-      return(stop(
-        "Error: Reliability can't be negative. Try again."))
-    if(export == FALSE & patterns == FALSE & slopes == FALSE)
-      return(stop(
-        "You said no data, no patterns, or no model information. What do you want?"))
-    
-    # ----------------------------------------------------- #
-    # BUILDING BLOCKS                                       #
-    # ----------------------------------------------------- #
[…]

### 2024-11-23 scripts/helper_functions/classification_metrics.R

@@ -1,65 +1,0 @@
-
-# - promises and pitfalls of panel change ------------------------------------ #
-# - helper: classification metrics ------------------------------------------- #
-
-### note: this script
-###       (a) takes a confusion matrix of true/simulated counts as its input,
-###       (b) calculates kappa, sensitivity, and specificity.
-
-require(tidyverse)
-
-# FUNCTION ------------------------------------------------------------------- #
-
-confused_outcomes <-
-  
-  function(confusion_matrix) {
-    
-    # --- calculate kappa
-    
-    # accuracy
-    p_0 =
-      (confusion_matrix[4, 3] 
-       + 
-       confusion_matrix[1, 3]
-      ) / 
-      sum(confusion_matrix$n)
-    
-    # expected classification
-    p_e =
-      (
-        (
-        (confusion_matrix[1, 3] + confusion_matrix[2, 3])
-        *
-        (confusion_matrix[1, 3] + confusion_matrix[3, 3])
-        )
-      +
-        (
-        (confusion_matrix[3, 3] + confusion_matrix[4, 3])
-        *
-        (confusion_matrix[2, 3] + confusion_matrix[4, 3])
-        )
-      ) /
-      (sum(confusion_matrix$n) ^ 2)
-    
-    # kappa score
-    kappa = (p_0 - p_e) / (1 - p_e)
-    
-    # --- calculate sensitivity
-    sensitivity =
-      confusion_matrix[4, 3] /
-      (confusion_matrix[4, 3] + confusion_matrix[3, 3])
-    
-    # --- calculate specificity
-    specificity =
-      confusion_matrix[1, 3] /
-      (confusion_matrix[1, 3] + confusion_matrix[2, 3])
-    
-    # --- export
-    output <- list(kappa = kappa$n,
-                   sensitivity = sensitivity$n,
-                   specificity = specificity$n)
-    
-    return(output)
-  }
-
-# ---------------------------------------------------------------------------- #

### 2024-12-24 paper/manuscript.Rmd
drafting for the front-end.

@@
 title: |
-  | The Promises and Pitfalls of Using Panel Data to Understand
-  | Individual Belief Change
+  | The Promises and Pitfalls of Using Panel Data to Understand Individual Belief Change[^thanks]
 author:
-  - Stephen Vaisey
-  - Turgut Keskintürk
+  - Turgut Keskintürk[^TK]
+  - Pablo Bello[^PB]
+  - Stephen Vaisey[^SV]
 date: "`r format(Sys.time(), '%B %Y')`"
@@
 csl: ./helpers/asa.csl
-abstract: "In this article, we investigate whether studies on political belief change can identify change trajectories at the individual level.
-Using simulations and empirical case studies, we propose a grid-search framework that allows researchers to evaluate the extent to which their target estimates generalize to individuals in their study population.
+abstract: "We investigate whether studies on political belief change can identify change trajectories at the individual level.
+Using simulations and case studies, we propose a grid-search framework that allows researchers to evaluate the extent to which their target estimates generalize to individuals in their study population.
 We use simulated datasets to estimate plausible values for how many people changed, how much they changed, and who changed, based on observed survey response trajectories.
-Our results suggest that researchers should think carefully about the conditions under which they can make claims about belief change at the individual level."
+Our results suggest that researchers should think carefully about the conditions under which they may make claims about belief change at the individual level.
+We provide an R package to help researchers implement this procedure in their own work."
 fontsize: 11pt
@@
 - \usepackage{array}
-- \usepackage{longtable}
-- \usepackage{booktabs}
 - \usepackage{float}
@@
 ---
+
+[^thanks]: We thank the participants of the *Worldview Lab at Kenan Institute for Ethics* at Duke University for their insightful and helpful comments on this project.
+[^TK]: Department of Sociology, Duke University, `turgut.keskinturk@duke.edu`.
+[^PB]: Department of Sociology, Duke University, `pablo.bello@duke.edu`.
+[^SV]: Department of Sociology, Duke University, `stephen.vaisey@duke.edu`.
 
@@
 
-Studies on political change at the individual level---change in one's attitudes, values, or preferences---face a challenging problem: while our methods typically provide estimates for change and stability at the group level, our goal is often to understand these processes at the individual level.
+Studies on political change at the individual level---change in one's beliefs and preferences---face a challenging problem: while our methods typically provide estimates for change and stability at the group level, our goal is often to understand these processes at the individual level.
 
@@
 While the previous research quantified change with various methods---ranging from assessing the time order of longitudinal observations [@kiley2020; @vaisey2020] to multilevel decomposition of over-time variance [@lersch2023]---these studies have fallen short of disentangling whether they find a small amount of change in a large number of people, or a large amount of change in a small number of people.
+In other words, while social science researchers have become widely sophisticated in evaluating the average levels of change over time in a population, they have mostly overlooked the fact that these average changes are produced by a combination of various compositional processes.
 
@@
 For example, the General Social Survey (GSS) shows that roughly 35% of the American adults in 2006 believed that "marijuana should be legalized." In less than 10 years, this support had increased to nearly 60%.
-As seen in Figure 1, this change occurred both *cross-sectionally*, where average support increased with time, and *longitudinally*, where at least some segments of the American population changed their minds on this policy.
+As seen in Figure 1, this change occurred both *cross-sectionally*, where average support increased with time, and *longitudinally*, where at least some segments of the American population changed their minds on this policy over these periods.
 
@@
 
-However, these two processes, cross-sectional change and longitudinal change, attract competing theories of political change, and group-level estimates may not properly adjudicate the underlying generative mechanisms.
-Following the same example, the cross-sectional change in beliefs about marijuana legalization may have occurred via two mechanisms: people may have changed their minds following a changing _zeitgeist_, or old generations who disapproved this policy might have been replaced by new generations approving it.
-Similarly, the longitudinal change could result from various mechanisms: large segments of the population may have shifted their opinions positively in small amounts, a small segment of the population may have made large changes, or a mix of changes---including both approvals and disapprovals---may have altered the overall balance.
+However, these two processes, cross-sectional change and longitudinal change, attract competing theories of political change, and group-level estimates may not properly adjudicate the underlying data generation processes.
+Following the same example, the cross-sectional change in beliefs about marijuana legalization may have occurred via two processes: people may have changed their minds following a changing _zeitgeist_, or old generations who disapproved this policy might have been replaced by new generations approving it.
+Similarly, the longitudinal change could result from various processes: large segments of the population may have shifted their opinions positively in small amounts, a small segment of the population may have made large changes, or a mix of changes---including both approvals and disapprovals---may have altered the overall balance.
 
-While the cross-sectional processes received strong scholarly attention in the literature [@vaisey2016; @restrepoochoa2024; @bartels2014], questions of the second sort remained largely unexplored.
+While the cross-sectional processes have received strong scholarly attention in the literature [@vaisey2016; @restrepoochoa2024; @bartels2014], questions of the second sort remained largely unexplored.
 However, given that theories of political ideology involve theoretical expectations about individual-level outcomes, the fact that we lack information about whether our models capture person-level empirical processes is detrimental to understanding adult belief change in the political context.
@@
 Using simulations and empirical case studies, we show that it is indeed possible to provide plausible scenarios as to which generative models are more likely to have produced empirical observations, even though a reliable identification of *who* changed occurs only under very narrow conditions.
-In doing this, we provide a simple grid-search procedure to aid researchers in evaluating the extent to which their target estimates may be generalized to the population of their study.
-We explore how competing generative models can lead to varying assumptions, emphasizing the importance for researchers to carefully consider the conditions under which claims about personal-level change can be made.
-Accordingly, we claim that when the meaning of "political change" is elusive and depends too much on how we define change in the first place, our empirical strategies could potentially mislead rather than inform, especially if we overlook the underlying generative mechanisms.
+In doing this, we build on the "Approximate Bayesian Computation" (APC) literature [@beaumont2010] and provide a simple grid-search procedure to help social researchers in evaluating the extent to which their target estimates might be generalized to the population of their studies.
+Using this procedure, we show that it is possible to approximate to the underlying data generation processes for *rate of change*, *strength of change*, and *direction of change* in a panel data.
+We provide an R package, `gridsearch`, to help researchers in implementing this procedure in their own work.
 
-We start with an overview of the "latent variable model" of beliefs and the problems of reliability and resolution in survey response measurement [@alwin2007].
-We then introduce our grid-search procedure in the context of three-wave panel data with binary outcomes, examining cases from the GSS on marijuana legalization and abortion.
-In the next section, we relax the period assumption and investigate how linear trajectories at the individual level can assist us in understanding underlying generative mechanisms in contexts with more than three waves.
-In the final empirical section, we analyze whether these generative models help us in classification tasks.
-We conclude by discussing the promises and pitfalls of panel data for understanding individual belief change.
+We emphasize a striking heterogeneity in these underlying processes even for a seemingly straightforward case like the marijuana legalization in the GSS, showing that theoretically opposite claims---e.g., large swaths of the population increasing their support for the policy, as opposed to a small group of people completely switching their positions---may very well account for the same data.
+This finding is particularly important in the longitudinal panel context, which, for most applications, represents the gold standard in evaluating change in observational social research.
+As such, we claim that when the meaning of "political change" is elusive, and depends too much on how we define change in the first place, our current empirical strategies could potentially mislead rather than inform, especially if we overlook the underlying generative mechanisms.
+
+We start with a general overview of the "latent variable model" of beliefs, the problems of reliability and resolution in survey measurement, and the compositional sources of aggregate change.
+We then introduce our grid-search procedure in the context of three-wave panel data with binary outcomes.
+Next, we relax this assumption and investigate how linear trajectories at the individual level can assist us in contexts with more than three waves.
+In the final section, we analyze whether this procedure helps us in classification tasks, particularly in evaluating *who* changed.
+We conclude by discussing the promises and pitfalls of panel data for understanding individual political change.
 
@@
 
-Alwin's [-@alwin2007] latent-variable approach to survey response is based on the assumption that there is a random variable $Y$, which captures a person $i$'s "true" position in a continuous latent construct.
-This random variable might stand in for any disposition, including one's political attitudes, values, and preferences, which allows us to define one's political position---at the level of one's "personal culture" [@lizardo2017]---as their *true score* $Y_i$.[^1]
+The latent-variable approach to survey response is based on the assumption that there is a random variable $Y$, which captures a person $i$'s "true" position in a continuous latent distribution [@alwin2007].
+This random variable might stand in for any disposition, including one's political beliefs and preferences, which allows us to define one's political position as their *true score* $Y_i$.[^truescore]
 
-[^1]: Naturally, the assumption that *there is indeed a true score* might miss cases where individuals hold contradictory ideas [@swidler1986] and are ambivalent about their true positions [@zaller1992; @zaller1992a].
-In this article, we follow the latent variable model and assume that there *is* a true position for each individual.
+[^truescore]: Naturally, the assumption that *there is indeed a true score* might miss cases where individuals hold contradictory ideas [@swidler1986] and are ambivalent about their true positions [@zaller1992; @zaller1992a].
+In this article, we follow the latent variable model and assume that there *is* a true position for each individual; hence, bracketing the cases where this ambivalence needs to be incorporated into the survey response measurement model.
 
@@
 This error might result from various factors that introduce biases to how we extract information from individuals, given that problems like question construction and interviewer effects might contaminate how individuals reveal their true preferences [@alwin1991].
-The inherent stochasticity in this measurement process implies that each draw from $Y_i$ might at least slightly differ from the true score, and this error term is difficult to circumvent with one measure, which is often the case in public opinion surveys.[^2] In this sense, there is often a degree of systematic gap between $Y_i$ and $y_i$.
+The inherent stochasticity in this measurement process implies that each draw from $Y_i$ might at least slightly differ from the true score, and this error term is difficult to circumvent with one measure, which is often the case in public opinion surveys.[^scales] In this sense, there is often a degree of systematic gap between $Y_i$ and $y_i$.
 
-[^2]: The idea that we should have repeated measures of $Y_i$ is the main intuition behind scale construction [@ansolabehere2008], with the explicit expectation that multiple measures can help us reduce measurement error.
+[^scales]: The idea that we should have repeated measures of $Y_i$ is the main intuition behind scale construction [@ansolabehere2008], with the explicit expectation that multiple measures can help us reduce measurement error.
 
-Hence, the main problem is mapping one's true scores to their observed scores, and there are at least two ways we can fail in this context.
-First, our observed measures might have low *reliability*, such that a hypothetical set of repeated draws may significantly differ from one's true score, due to, e.g., poor question construction or individual idiosyncrasies.
-Second, survey response categories may be imperfect: the *resolution* of response categories may fail to distinguish consequential differences in people's true scores.
-In either case, the observed score may not reliably indicate one's true score.
-
-## Separating True Change and Measurement Error in Panel Data
-
-In the panel context, where we observe the same individuals across time, the measurement problem causes substantive questions: if true scores are imperfectly realized and we only have access to one's observed responses, how can we know if a person *actually changed* between time $t$ and $t + 1$, or whether the observed change is merely due to measurement error?
+In the panel context, where we observe the same people multiple times, this measurement problem has substantive implications: if true scores are imperfectly realized and we only have surface access to a person's observed survey responses, how can we know if they *actually* changed between time $t$ and $t + 1$, or whether the observed change is merely due to measurement error?
 
 Scholars in public opinion research have attempted to address this issue in multiple ways, using structural equation models that explicitly quantify the measurement error component [@achen1975], strategies that integrate longitudinal surveys to exploit the information in the ordering of responses [@kiley2020], and various multilevel mixed-effects models that shrink responses based on population estimates to derive reliable trends [@lersch2023; @hout2016].
-To some extent, all these strategies were fruitful to offset the problems associated with measurement error.[^3] That said, while these empirical strategies were largely successful in accounting for the amount of change *in the population*, they were rather silent about *who* changed in our observations.
+To some extent, all these strategies were fruitful to offset the problems associated with measurement error.[^stability] That said, while these empirical strategies were largely successful in accounting for the amount of change *in the population*, they were rather silent about changes *at the level of individuals*.
 
-[^3]: However, note that the use of repeated observations in the panel context comes with additional assumptions: it is rather straightforward to assume that multiple measures at time $t$ can be indicative of an underlying construct---e.g., we can average them to have an index---but the use of different time periods needs a *stability* assumption, that is, the reliability estimates may only be computed if one assumes that people do not really change between waves, which might confuse true change with measurement error [see, e.g., @hout2016].
+[^stability]: However, note that the use of repeated observations in the panel context comes with additional assumptions: it is rather straightforward to assume that multiple measures at time $t$ can be indicative of an underlying construct---e.g., we can average them to have an index---but the use of different time periods needs a *stability* assumption, that is, the reliability estimates may only be computed if one assumes that people do not really change between waves, which might confuse true change with measurement error [see, e.g., @hout2016].
 
-Why is this important?
-After all, if we can produce reasonable estimates at the population level, this must surely be enough.
-We believe that there are at least two reasons why we should care about this problem.
-First, the inability to identify the individuals who truly change gives us a false sense of homogeneity.
-When we report an average change of, say, 0.10 standard deviations, we do not know whether this reflects uniform change in latent positions across all individuals, or significant changes in some while others remain stable.
-This is a big substantive difference, yet often overlooked in practice.
-Second, if we cannot identify the "changers" in the first place, our comparisons between groups---either across some pre-determined characteristic, or across treatment status---will average true processes with false negatives and false positives.
-This is obviously fine if the error occurs randomly, but it is challenging to verify this if individual-level accuracy is compromised.
-Hence, glossing over this problem discourages us to think about what exactly we are averaging.
+## The Compositional Sources of Aggregate Change
 
-## The Question of Individual Belief Change
+Suppose we have a panel data, where individuals $i = 1, ..., N$ are observed over two time periods $t_0$ and $t_1$.
+Let $y_{it}$ represent an observed survey response.
+To see how much observed change there is at the individual level, we can simply define a change quantity, $\delta y_{i, t_0 \to t_1} = y_{it_1} - y_{it_0}$, the average of which, $\bar{y}$, represents the mean change in the dataset from $t_0$ to $t_1$.
+
+This quantity is a natural candidate for describing aggregate belief change, but, under the hood, it summarizes three compositional processes in one metric.
+Let us define these processes intuitively.
+The first component is $R$, representing the proportion of individuals in our dataset with $\delta y_{i, t_0 \to t_1} \neq 0$, i.e., the proportion of individuals who changed in this period.
+Call this *the rate of change*.
+The second component is $S$, representing the "average" magnitude of change among those who changed.[^heterogeneity] Call this *the strength of change*.
+The third component is $D$, representing the net directionality of change, i.e., the difference in proportions of positive and negative changers in our observations.
+Call this *the direction of change*.
+If we combine these three components, we can represent $\bar{y}$ simply as
+
+$$
+\bar{y} = R \times S \times D
+$$
+
+[^heterogeneity]: Of course, we might further decompose this average strength to a distribution of heterogeneous change trajectories across individuals.
+For analytical and expository clarity, we bracket this issue from the scope of this article.
+
+This simple decomposition should let us appreciate that the average change score may be generated via multiple data generation processes, each having strongly different theoretical implications.
+
+For instance, if we want to evaluate whether a society became more pro-redistribution over time, we need to note that an increase in pro-redistributive preferences can occur via all or most people reporting higher preference for redistribution, some people switching their positions altogether while some others remain completely stable, or everyone changing in one or the other direction, and the net score we observe is simply the difference in their overall trajectories.
+The same reasoning applies to the GSS's marijuana legalization policy that we have discussed before.
+Most discussions in political belief change---confidence in legislative groups or financial institutions, beliefs about the economy, or ethnic attitudes---have the exact same question about their data generation processes.
+Simply put, the findings about change at the population level may simply result from a variety of conflicting, though observationally equivalent, data generation processes.
+
+We can now complicate this composition question by noting, once again, the survey measurement problem of mapping one's true scores to their observed scores, which often fails in at least two ways: (1) our observed measures might have low *reliability*, such that a hypothetical set of repeated draws may significantly differ from one's true score, due to, e.g., poor question construction or individual idiosyncrasies, and (2) question response categories may be imperfect: the *resolution* of response categories may fail to distinguish consequential differences in true scores.
+Therefore, the average change captured in $\bar{y}$ may represent a biased or imperfect estimate of change to begin with.
 
@@
 
-We pose these questions in a fixed-effects panel context [@allison2009; @vaisey2017], with varying individual trajectories across time [@rüttenauer2020].
-This achieves two things.
-First, it allows us to shift the focus from *distributional change* to *individual change*, given that our theoretical interest is to understand how people change their political beliefs over the life-course.
-Second, it allows us to model response trajectories, rather than individual changes in response to a shock.
-Instead, our focus is "classification"---the extent to which we can discriminate those who change from those who do not change in a specific time window.
+We pose these questions in a fixed-effects longitudinal panel data context [@allison2009; @vaisey2017], with varying individual trajectories across time periods [@rüttenauer2020].
+This allows us to shift the focus from *aggregate change* to *individual change*, given that our theoretical interest is to understand how people change their political beliefs over the life-course.
 
-## A Grid-Search Procedure for Identifying Individual Belief Change
+## A Grid-Search Procedure
 
-We propose a grid-search procedure to answer our research questions. The idea is straightforward:
+We build on the Approximate Bayesian Computation (APC) literature [@beaumont2010] and propose a grid-search procedure to answer our research questions.
+The APC approach involves simulating parameter values given the data and evaluating the distribution of these parameters that potentially generated our observations.
+We build a rejection sampler algorithm that
 
-(1) Generate simulated datasets with several parameters of interest---rate of change, strength of change, and the direction of change, so that we know the real data-generation process (DGP);
-(2) Calculate the distribution of response trajectories in both the observed and simulated data;
-(3) Compare the simulated trajectories to the observed trajectories, calculating a distance measure to quantify the extent to which specific simulated values reproduce the observed values;
-(4) Replicate this analysis $N$ times for all possible parameter combinations;
-(5) Inspect the parameter combinations that most closely approximate to the observed values.
-
-This is simply what Gelman and Hill [-@gelman2007] call a "fake-data simulation," and it is powerful for at least two reasons.
-First, it allows us to produce reasonable estimates as to what plausible generative models might have produced the real data, ensuring that we can make educated guesses about the extent to which our parameter estimates might generalize to our populations.
-Second, it is generative, meaning that we can ask specific questions about response trajectories and modeling assumptions.
-
-Our fake-data simulations involve several DGP components: a random variable $Y_{it}$ for true scores, with varying realizations $y_{it}$, representing a person $i$'s beliefs at time $t$; an indicator for change, $D_{it}$, operationalized as a non-reversible trigger if there is a respondent-level switch in true scores; an effect size, $\delta$, and a respondent level direction multiplier $\tau_i$ that indexes whether the change occurs in a positive or a negative direction.
-Using these components, the DGP follows
+1) Simulates data with several parameters of interest for our DGPs---rate of change, strength of change, direction of change, and reliability of response measurement using
 
@@
 
-Note that we allow actors to go through change at any time in the observation window of the study, meaning that $D_{it}$ indexes an actor's change status in their biographical history.
+    where we have a random variable $Y_{i}$ for true scores, with varying realizations $y_{it}$, representing a person $i$'s beliefs at time $t$; an indicator for change, $D_{it}$, operationalized as a non-reversible trigger if there is a respondent-level change in true scores; an effect size, $\delta$, and a respondent level direction multiplier $\tau_i$ that indexes whether the change is positive or negative.[^changing]
 
-Since our research interest concerns individual trajectories of change, we specify three parameters for these DGPs: (1) *rate of change* refers to the proportion of our sample that goes through change in their political orientations, (2) *strength of change* refers to the standardized mean change in the true score when someone changes---described as Cohen's $d$---and (3) *direction of change* refers to the proportion of changers who change in positive and negative directions.
-The combination of these three parameters provides a complete picture of population movement in political beliefs.[^4]
+2) Calculates the distribution of response trajectories in both the observed and simulated data, by either tabulating response patterns across time or estimating slope coefficients;
 
-[^4]: We define additional parameters that we fix in the case studies presented below: $N$ refers to the number of actors in the fake-data simulation, $T$ refers to the number of waves we have in the panel study, *balance* refers to the observed marginals in the real datasets, *reliability* refers to the item-reliability score regulating the extent to which a true score is realized in the survey context with given level of item reliability [@alwin2007], and *resolution* refers to the number of categories to simplify the latent variable, which is restricted to the binary case in this article.
+3) Compares the simulated trajectories to the observed trajectories, calculating a "distance" score to quantify the extent to which specific simulated values reproduce the observed values;
 
-Using these parameters, we can construct various generative models and compare the simulated trajectories to the observed trajectories to examine how closely certain parameter combinations approximate our empirical observations.
-In the next three sections, we will apply this procedure in different contexts.
-In all cases, we restrict our attention to latent variables with binary outcomes to simplify our analyses, but all results are generalizable to cases with higher resolution.[^repo]
+[^changing]: Note that we "allow" actors to go through change at any time in the observation window of the study, meaning that $D_{it}$ is variable over $t$ precisely because actors can change any time in their biographical history.
 
-[^repo]: The replication files for the article are stored in https://github.com/tkeskinturk/simulating_change
+4) Rejects the simulated data if the distance measure is above a certain pre-specified "caliper" value, and accepts it otherwise;
+
+5) Replicates this analysis $N$ times by drawing samples from the specified priors;
+
+6) Inspects the parameter combinations that most closely approximate to the observed values.
+
+The grid-search approach we propose is powerful for at least two reasons.
+First, it allows us to produce reasonable estimates as to what plausible generative models might have produced the real data, ensuring that we can make educated inferences about the extent to which our parameter estimates might generalize to our populations.
+Second, it is generative, meaning that we can ask specific questions about response trajectories and modeling assumptions.[^parameters]
+
+[^parameters]: We define additional parameters that we fix in the case studies presented below: $N$ refers to the number of actors in the fake-data simulation, $T$ refers to the number of waves we have in the panel study, *balance* refers to the observed marginals in the real datasets, allowing us to "cut" the continuous distribution to 0s and 1s, and *resolution* refers to the number of categories to simplify the latent variable, which is restricted to the binary case in this article.
+
+In the next four sections, we apply this procedure in different contexts.
+In all cases, we restrict our attention to latent variables with binary outcomes to simplify our analyses.
+The replication files for the article are stored in https://osf.io/AAAAA.
+The R package for the use of this approach, `gridsearch`, can be accessed via the replication repo or https://tkeskinturk.github.io/gridsearch/.
+
+\newpage
 
@@
 
-# Study 2: Longer Panels and Linear Trajectories
+## The Case of Generalized Trust
+
+# Study 2: Simulation Studies and Performance Assessment
+
+# Study 3: Longer Panels and Linear Trajectories
 
@@
 
-# Study 3: Detection of True Changers
+# Study 4: Detection of True Changers
 
@@
 
+\clearpage
+
 # Discussion and Conclusions
@@
 Using simulations and empirical case studies, we showed that a grid-search procedure that iterates over generative models can help us identify plausible data generation processes that may have produced our empirical observations, though the identification of individual cases occurs under narrow conditions.
-We argued that the *group-to-person generalizability problem* is an impediment to a robust understanding of political belief change, and scholars should understand the extent to which their target estimates can generalize to the population of their study.
+We argued that the *group-to-person generalizability problem* is an impediment to a robust understanding of political belief change, and scholars should understand the extent to which their target estimates may generalize to the population of their study.
 As such, we proposed a simulation approach that takes response generation seriously, which may help us move forward in addressing this problem.
 
-This article has several implications for the study of belief change across the life-course.
+This article has several implications for the study of political belief change across the life-course.
 First, our findings suggest that recent debates about the existence or non-existence of belief change in adult populations [@kiley2020; @lersch2023] should tackle the heterogeneous nature of change and stability in individual preferences since group-level measures may hamper our understanding of individual mechanisms.

### 2024-12-24 paper/revision_memo.Rmd
drafting for the front-end.

@@
 - We added a new item using the *British Elections Study*'s panel study, generalized trust, an item collected between 2019 and 2024 across 6 years and 5 waves, in a potentially disruptive time period due to COVID-19.
-As a more general note, we believe that the tool we propose should apply in all cases, but we agree that a more thorough set of examples are helpful.
+As a more general note, we believe that the tool we propose should apply in all cases, but we agree that a more thorough set of examples is helpful.
 - Additionally, we now present a simulation study that applies our protocol in different DGPs and assesses the conditions under which we can make better guesses about the true DGPs (see below for more details).
@@
 
-## The Clarity of the Grid Search Discussions
+## The Grid-Search Procedure
 
@@
 
-We implemented three solutions for this issue.
-First, we now introduce our grid-search algorithm as an R package, `gridsearch`, which will accompany to this article.
-As we detail in the paper, we revised our procedure and substantially increased its efficiency, which means that it will be easy for people to implement our procedure in their own projects.
-Second, we now include a full replication package for the analyses presented in this project as a footnote in the article.
-We were planning to do so along with publication, but for a paper of this kind, it is beneficial for the reviewers to see the full code, so we attached a view-only (anonymized) link in the manuscript.
-Finally, we made substantive rewrites to the manuscript to make our protocol as clear and concise as possible.
+We implemented three solutions and a larger revision for this problem:
+
+1. We updated the mechanics of our approach by using Approximate Bayesian Computation [@beaumont2010], which increased the efficiency and generalizability of our procedure.
+2.
+We now introduce our grid-search algorithm as an accompanying R package, `gridsearch`.
+As we detailed in the paper, we revised our procedure and substantially increased its efficiency, which means that it will be easy for people to implement our method in their own projects.
+3. We included a full (anonymized) replication package for the analyses presented in this project.
+4. We made substantial rewrites to the manuscript to increase the clarity of our protocol.
+
 We hope that the R package, the replication repo, and these rewrites make our strategy more clear.
@@
 The Authors
+
+# References

### 2024-12-24 paper/manuscript.Rmd
ugh, what is even APC?

@@
 Using simulations and empirical case studies, we show that it is indeed possible to provide plausible scenarios as to which generative models are more likely to have produced empirical observations, even though a reliable identification of *who* changed occurs only under very narrow conditions.
-In doing this, we build on the "Approximate Bayesian Computation" (APC) literature [@beaumont2010] and provide a simple grid-search procedure to help social researchers in evaluating the extent to which their target estimates might be generalized to the population of their studies.
+In doing this, we build on the "Approximate Bayesian Computation" (ABC) literature [@beaumont2010] and provide a simple grid-search procedure to help social researchers in evaluating the extent to which their target estimates might be generalized to the population of their studies.
 Using this procedure, we show that it is possible to approximate to the underlying data generation processes for *rate of change*, *strength of change*, and *direction of change* in a panel data.
@@
 
-We build on the Approximate Bayesian Computation (APC) literature [@beaumont2010] and propose a grid-search procedure to answer our research questions.
-The APC approach involves simulating parameter values given the data and evaluating the distribution of these parameters that potentially generated our observations.
+We build on the Approximate Bayesian Computation (ABC) literature [@beaumont2010] and propose a grid-search procedure to answer our research questions.
+The ABC approach involves simulating parameter values given the data and evaluating the distribution of these parameters that potentially generated our observations.
 We build a rejection sampler algorithm that

### 2025-01-09 paper/manuscript.Rmd
reworking with ABC.

@@
 [^PB]: Department of Sociology, Duke University, `pablo.bello@duke.edu`.
-[^SV]: Department of Sociology, Duke University, `stephen.vaisey@duke.edu`.
+[^SV]: Department of Sociology and Political Science, Duke University, `stephen.vaisey@duke.edu`.
 
@@
 
-```{=latex}
-\begin{figure}[!ht]
-\begin{center}
-```
-```{r figure7, out.width = "100%"}
-knitr::include_graphics("../figures/detection.png")
-```
-```{=latex}
-\end{center}
-\caption*{Figure 7: Sensitivity and Specificity Across Case Studies}
-\footnotesize{\textit{Notes:} The figure depicts the sensitivity---true positives divided by the sum of true positives and false positives---and specificity---true negatives divided by the sum of true negatives and false negatives---scores across three studies.}
-\end{figure}
-```
-
 These results suggest that, while the grid-search procedure allows us to identify plausible generative models that may have generated the empirical patterns, detecting true change at the level of individuals occurs under only very narrow conditions.

### 2025-01-09 scripts/1_ksdist.R

@@ -1,0 +1,168 @@
+
+# - promises and pitfalls of panel change ------------------------------------ #
+# - ks distance example ------------------------------------------------------ #
+
+### note: this script
+###       (a) generates an exemplary panel data from a DGP A,
+###       (b) simulates an exemplary panel data from a DGP B,
+###       (c) showcases the KS distance approach using their slope distribution.
+
+rm(list = ls()) # clean-up
+
+# install.packages("devtools")
+# devtools::install_github("tkeskinturk/gridsearch")
+
+pacman::p_load(
+  purrr,
+  gridsearch,
+  tidyverse,
+  hrbrthemes,
+  GGally,
+  patchwork)
+theme_set(theme_ipsum_rc())
+
+# --- helpers
+my_oka <-
+  c("#eebd64",
+    "#8b3b36", 
+    "#a2b4b7")
+`%nin%` = Negate(`%in%`)
+
+set.seed(112358)
+
+# --- glm function
+glm_fit <- function(data) {
+    fit <- suppressWarnings(
+      stats::glm(y ~ scales::rescale(t), 
+                 data = data, family = stats::binomial()))
+    prob1 <- stats::plogis(fit$coefficients[1])
+    prob2 <- stats::plogis(fit$coefficients[1] + fit$coefficients[2])
+    estimate <- round(prob2 - prob1, 3)
+    return(estimate)
+  }
+
+# PART 1: DGPs --------------------------------------------------------------- #
+
+### DGP A
+
+# --- read in
+data_A <- buildDGP(
+  n = 500,
+  t = 12,
+  rate = 0.5,
+  balance_dir = 1,
+  balance_res = .50,
+  strength =  1,
+  reliable = .8,
+  export = TRUE
+)$data
+
+# --- organize
+data_A <- data_A |>
+  as_tibble() |>
+  mutate(pid = 1:500) |>
+  pivot_longer(cols = -pid,
+               names_to = "cols",
+               values_to = "y") |>
+  mutate(cols = factor(cols, levels = paste0("V", 1:12))) |> 
+  group_by(cols) |> mutate(t = cur_group_id()) |> ungroup() |>
+  nest(.by = "pid") |>
+  mutate(slopes_A = purrr::map_dbl(.x = data, .f = ~ glm_fit(.x))) |>
+  select(pid, slopes_A)
[…]

### 2025-01-09 scripts/1_sims.R

@@ -1,213 +1,0 @@
-
-# - promises and pitfalls of panel change ------------------------------------ #
-# - script 1 # parameter space for `gridSearch` ------------------------------ #
-
-### note: this script
-###       (a) generates simulated DGPs with varying parameters,
-###       (b) applies the grid search algorithm to recover error distributions,
-###       (c) evaluates whether the algorithm can recover "true" DGP.
-
-rm(list = ls()) # clean-up
-
-# install.packages("devtools")
-# devtools::install_github("tkeskinturk/gridsearch")
-
-pacman::p_load(
-  purrr,
-  furrr,
-  hrbrthemes,
-  gridsearch,
-  ggforce,
-  tidyverse,
-  broom.helpers)
-theme_set(theme_ipsum_rc())
-
-# --- helpers
-my_oka <-
-  c("#eebd64",
-    "#8b3b36", 
-    "#a2b4b7")
-`%nin%` = Negate(`%in%`)
-
-# PART 1: DATA --------------------------------------------------------------- #
-
-set.seed(11235)
-
-# --- conditions
-d <-
-  expand_grid(
-    ## 500 individuals
-    p_n = 500,
-    ## varying panel waves (restrict to contingency)
-    p_t = c(3, 5, 7),
-    ## the rate of change in the population
-    p_rate = c(.1, .25, .5),
-    ## the direction of change (halfsies versus secular)
-    p_dir = c(.5, 1),
-    ## the fixed level of marginal distributions
-    p_res = c(.5, .75),
-    ## the strength of change in latent variable
-    p_strength = c(.5, 1),
-    ## the reliability score of the outcome measurement
-    p_reliable = .9,
-    ## the number of simulated datasets
-    sims = 1:30)
-
-# --- dataframes
-d <- d |>
-  dplyr::mutate(
-    data =
-      furrr::future_pmap(
-        ## mapping list
-        .l = list(p_n,
-                  p_t,
-                  p_rate,
-                  p_dir,
-                  p_res,
-                  p_strength,
-                  p_reliable),
-        ## refer to list based on index
-        .f = purrr::possibly(
-          ~ buildDGP(
-            # varying parameters
-            n = ..1,
-            t = ..2,
-            rate = ..3,
-            balance_dir = ..4,
-            balance_res = ..5,
-            strength = ..6,
-            reliable = ..7,
-            export = TRUE,
[…]

### 2025-01-09 scripts/2_grass.R

@@ -1,0 +1,172 @@
+
+# - promises and pitfalls of panel change ------------------------------------ #
+# - case 1 # grass ----------------------------------------------------------- #
+
+### note: this script
+###       (a) takes `grass` as an examplary GSS item,
+###       (b) simulates potential DGPs approximating the observed parameters,
+###       (c) shows the relative fit of the DGPs to observed data.
+
+rm(list = ls()) # clean-up
+
+# install.packages("devtools")
+# devtools::install_github("tkeskinturk/gridsearch")
+
+pacman::p_load(
+  srvyr,
+  gssr,
+  gridsearch,
+  tidyverse,
+  hrbrthemes,
+  GGally)
+theme_set(theme_ipsum_rc())
+
+# --- helpers
+my_oka <-
+  c("#eebd64",
+    "#8b3b36", 
+    "#a2b4b7")
+`%nin%` = Negate(`%in%`)
+
+# PART 1: GSS DATA ----------------------------------------------------------- #
+
+data("gss_panel06_long")
+data("gss_panel08_long")
+data("gss_panel10_long")
+
+# --- get data
+gss <-
+  bind_rows(
+    gss_panel06_long |>
+      select(pid = firstid, year, wave, wtpannr123, grass) |>
+      haven::zap_labels() |>
+      mutate(pid = as.integer(paste0(
+        2006, as.integer(pid)
+      ))) |>
+      mutate(run = "2006-2010"),
+    gss_panel08_long |>
+      select(pid = firstid, year, wave, wtpannr123, grass) |>
+      haven::zap_labels() |>
+      mutate(pid = as.integer(paste0(
+        2008, as.integer(pid)
+      ))) |>
+      mutate(run = "2008-2012"),
+    gss_panel10_long |>
+      select(pid = firstid, year, wave, wtpannr123, grass) |>
+      haven::zap_labels() |>
+      mutate(pid = as.integer(paste0(
+        2010, as.integer(pid)
+      ))) |>
+      mutate(run = "2010-2014")
+  )
+
+# --- clean-up
+gss <- gss |>
+  mutate(grass = as.character(grass)) |>
+  mutate(grass = case_when(grass == "Legal" ~ 1,
+                           grass == "Not Legal" ~ 0,
+                           .default = NA_real_)) |>
+  drop_na() |>
+  mutate(n = n(), .by = pid) |> 
+  filter(n == 3) |> 
[…]

### 2025-01-09 scripts/3_abortion.R

@@ -1,0 +1,172 @@
+
+# - promises and pitfalls of panel change ------------------------------------ #
+# - case 2 # abortion -------------------------------------------------------- #
+
+### note: this script
+###       (a) takes `abany` as an examplary GSS item,
+###       (b) simulates potential DGPs approximating the observed parameters,
+###       (c) shows the relative fit of the DGPs to observed data.
+
+rm(list = ls()) # clean-up
+
+# install.packages("devtools")
+# devtools::install_github("tkeskinturk/gridsearch")
+
+pacman::p_load(
+  srvyr,
+  gssr,
+  gridsearch,
+  tidyverse,
+  hrbrthemes,
+  GGally)
+theme_set(theme_ipsum_rc())
+
+# --- helpers
+my_oka <-
+  c("#eebd64",
+    "#8b3b36", 
+    "#a2b4b7")
+`%nin%` = Negate(`%in%`)
+
+# PART 1: GSS DATA ----------------------------------------------------------- #
+
+data("gss_panel06_long")
+data("gss_panel08_long")
+data("gss_panel10_long")
+
+# --- get data
+gss <-
+  bind_rows(
+    gss_panel06_long |>
+      select(pid = firstid, year, wave, wtpannr123, abany) |>
+      haven::zap_labels() |>
+      mutate(pid = as.integer(paste0(
+        2006, as.integer(pid)
+      ))) |> 
+      mutate(run = "2006-2010"),
+    gss_panel08_long |>
+      select(pid = firstid, year, wave, wtpannr123, abany) |>
+      haven::zap_labels() |>
+      mutate(pid = as.integer(paste0(
+        2008, as.integer(pid)
+      ))) |> 
+      mutate(run = "2008-2012"),
+    gss_panel10_long |>
+      select(pid = firstid, year, wave, wtpannr123, abany) |>
+      haven::zap_labels() |>
+      mutate(pid = as.integer(paste0(
+        2010, as.integer(pid)
+      ))) |> 
+      mutate(run = "2010-2014")
+  ) 
+
+# --- clean-up
+gss <- gss |>
+  mutate(abany = as.character(abany)) |>
+  mutate(abany = case_when(abany == "Yes" ~ 1,
+                           abany == "No" ~ 0,
+                           .default = NA_real_)) |>
+  drop_na() |>
+  mutate(n = n(), .by = pid) |> 
+  filter(n == 3) |> 
[…]

### 2025-01-09 scripts/3_grass.R

@@ -1,141 +1,0 @@
-
-# - promises and pitfalls of panel change ------------------------------------ #
-# - case 1 # grass ----------------------------------------------------------- #
-
-### note: this script
-###       (a) takes `grass` as an examplary GSS item,
-###       (b) simulates potential DGPs approximating the observed parameters,
-###       (c) shows the relative fit of the DGPs to observed data.
-
-rm(list = ls()) # clean-up
-
-# install.packages("devtools")
-# devtools::install_github("tkeskinturk/gridsearch")
-
-pacman::p_load(
-  gssr,
-  purrr,
-  furrr,
-  srvyr,
-  hrbrthemes,
-  gridsearch,
-  tidyverse)
-theme_set(theme_ipsum_rc())
-
-# --- helpers
-my_oka <-
-  c("#eebd64",
-    "#8b3b36", 
-    "#a2b4b7")
-`%nin%` = Negate(`%in%`)
-
-# PART 1: GSS DATA ----------------------------------------------------------- #
-
-data("gss_panel06_long")
-data("gss_panel08_long")
-data("gss_panel10_long")
-
-# --- get data
-gss <-
-  bind_rows(
-    gss_panel06_long |>
-      select(pid = firstid, year, wave, wtpannr123, grass) |>
-      haven::zap_labels() |>
-      mutate(pid = as.integer(paste0(
-        2006, as.integer(pid)
-      ))) |>
-      mutate(run = "2006-2010"),
-    gss_panel08_long |>
-      select(pid = firstid, year, wave, wtpannr123, grass) |>
-      haven::zap_labels() |>
-      mutate(pid = as.integer(paste0(
-        2008, as.integer(pid)
-      ))) |>
-      mutate(run = "2008-2012"),
-    gss_panel10_long |>
-      select(pid = firstid, year, wave, wtpannr123, grass) |>
-      haven::zap_labels() |>
-      mutate(pid = as.integer(paste0(
-        2010, as.integer(pid)
-      ))) |>
-      mutate(run = "2010-2014")
-  )
-
-# --- clean-up
-gss <- gss |>
-  mutate(grass = as.character(grass)) |>
-  mutate(grass = case_when(grass == "Legal" ~ 1,
-                           grass == "Not Legal" ~ 0,
-                           .default = NA_real_)) |>
-  drop_na() |>
[…]

### 2025-01-09 scripts/4_abortion.R

@@ -1,141 +1,0 @@
-
-# - promises and pitfalls of panel change ------------------------------------ #
-# - case 2 # abortion -------------------------------------------------------- #
-
-### note: this script
-###       (a) takes `abany` as an examplary GSS item,
-###       (b) simulates potential DGPs approximating the observed parameters,
-###       (c) shows the relative fit of the DGPs to observed data.
-
-rm(list = ls()) # clean-up
-
-# install.packages("devtools")
-# devtools::install_github("tkeskinturk/gridsearch")
-
-pacman::p_load(
-  gssr,
-  purrr,
-  furrr,
-  srvyr,
-  hrbrthemes,
-  gridsearch,
-  tidyverse)
-theme_set(theme_ipsum_rc())
-
-# --- helpers
-my_oka <-
-  c("#eebd64",
-    "#8b3b36", 
-    "#a2b4b7")
-`%nin%` = Negate(`%in%`)
-
-# PART 1: GSS DATA ----------------------------------------------------------- #
-
-data("gss_panel06_long")
-data("gss_panel08_long")
-data("gss_panel10_long")
-
-# --- get data
-gss <-
-  bind_rows(
-    gss_panel06_long |>
-      select(pid = firstid, year, wave, wtpannr123, abany) |>
-      haven::zap_labels() |>
-      mutate(pid = as.integer(paste0(
-        2006, as.integer(pid)
-      ))) |> 
-      mutate(run = "2006-2010"),
-    gss_panel08_long |>
-      select(pid = firstid, year, wave, wtpannr123, abany) |>
-      haven::zap_labels() |>
-      mutate(pid = as.integer(paste0(
-        2008, as.integer(pid)
-      ))) |> 
-      mutate(run = "2008-2012"),
-    gss_panel10_long |>
-      select(pid = firstid, year, wave, wtpannr123, abany) |>
-      haven::zap_labels() |>
-      mutate(pid = as.integer(paste0(
-        2010, as.integer(pid)
-      ))) |> 
-      mutate(run = "2010-2014")
-  ) 
-
-# --- clean-up
-gss <- gss |>
-  mutate(abany = as.character(abany)) |>
-  mutate(abany = case_when(abany == "Yes" ~ 1,
-                           abany == "No" ~ 0,
-                           .default = NA_real_)) |>
-  drop_na() |>
[…]

### 2025-01-09 scripts/4_trust.R

@@ -1,0 +1,159 @@
+
+# - promises and pitfalls of panel change ------------------------------------ #
+# - case 3 # trust ----------------------------------------------------------- #
+
+### note: this script
+###       (a) takes `trust` as an examplary BES item,
+###       (b) simulates potential DGPs approximating the observed parameters,
+###       (c) shows the relative fit of the DGPs to observed data.
+
+rm(list = ls()) # clean-up
+
+# install.packages("devtools")
+# devtools::install_github("tkeskinturk/gridsearch")
+
+pacman::p_load(
+  srvyr,
+  gridsearch,
+  tidyverse,
+  hrbrthemes,
+  GGally)
+theme_set(theme_ipsum_rc())
+
+# --- helpers
+my_oka <-
+  c("#eebd64",
+    "#8b3b36", 
+    "#a2b4b7")
+`%nin%` = Negate(`%in%`)
+
+# PART 1: BES DATA ----------------------------------------------------------- #
+
+# --- get data
+data <-
+  haven::read_dta("./data/source_files/bes.dta") |>
+  select(
+    pid = id,
+    trust_1 = genTrustW17,
+    trust_2 = genTrustW20,
+    trust_3 = genTrustW21,
+    trust_4 = genTrustW23,
+    trust_5 = genTrustW27
+  ) |>
+  haven::zap_labels() |>
+  mutate_at(
+    .vars = vars(matches("trust")),
+    .funs = ~ case_when(. == 1 ~ 1, . == 2 ~ 0, TRUE ~ NA_real_)
+  ) |>
+  drop_na() |>
+  pivot_longer(cols = -pid,
+               names_to = "var",
+               values_to = "y") |>
+  separate(var, into = c("var", "t")) |>
+  mutate(t = as.integer(t)) |> select(-var) |> 
+  mutate(
+    year =
+      case_when(
+        t == 1 ~ 2019,
+        t == 2 ~ 2020,
+        t == 3 ~ 2021,
+        t == 4 ~ 2022,
+        t == 5 ~ 2024
+      )
+  )
+
+# --- draw trajectories
+png(
+  "figures/trust01.png",
+  w = 7,
+  h = 5,
+  units = "in",
+  res = 500
+)
+glm(y ~ factor(year), data = data, family = "binomial") |>
+  ggeffects::ggemmeans(terms = "year") |>
+  as_tibble() |>
+  ## draw
+  ggplot(aes(
+    x = x,
+    y = predicted
+  )) +
+  ## statistics
[…]

### 2025-01-09 scripts/5_immig.R

@@ -1,0 +1,141 @@
+
+# - promises and pitfalls of panel change ------------------------------------ #
+# - case 4 # immigration ----------------------------------------------------- #
+
+### note: this script
+###       (a) takes `imm` as an examplary panel data item,
+###       (b) simulates potential DGPs approximating the observed parameters,
+###       (c) shows the relative fit of the DGPs to observed data.
+
+rm(list = ls()) # clean-up
+
+# install.packages("devtools")
+# devtools::install_github("tkeskinturk/gridsearch")
+
+pacman::p_load(
+  srvyr,
+  gridsearch,
+  tidyverse,
+  hrbrthemes,
+  GGally)
+theme_set(theme_ipsum_rc())
+
+# --- helpers
+my_oka <-
+  c("#eebd64",
+    "#8b3b36", 
+    "#a2b4b7")
+`%nin%` = Negate(`%in%`)
+
+# PART 1: 26-WAVE DATA ------------------------------------------------------- #
+
+# --- get data
+data <- 
+  read_csv("./data/source_files/yllanon.csv") |> 
+  ## organize
+  mutate(pid = cur_group_id(), .by = "id") |>
+  ## chop off post-COVID periods
+  filter(wave < 18) |>
+  ## simplify
+  select(pid, t = wave, y = imm) |>
+  ## clean-up the outcome
+  filter(y %nin% c(8, 9)) |> drop_na() |>
+  ## fully balanced
+  mutate(n = n(), .by = "pid") |> filter(n == 17) |>
+  ## scale the variables
+  mutate(y = ifelse(y > 4, 1, 0)) |>
+  ## select and wrap up organization
+  select(pid, t, y) |> arrange(pid, t)
+
+# --- draw trajectories
+png(
+  "figures/immigration01.png",
+  w = 7,
+  h = 5,
+  units = "in",
+  res = 500
+)
+data |>
+  group_by(t) |>
+  summarize(y = mean(y)) |>
+  ## draw
+  ggplot(aes(x = t, y = y)) +
+  ## statistics
+  geom_line(linetype = "dashed", linewidth = 0.2) +
+  geom_point(size = 2) +
+  ## labels
+  labs(x = "Time", y = "Immigration Spending Preference") +
+  scale_x_continuous(breaks = scales::pretty_breaks(n = 17)) +
+  scale_y_continuous(
+    limits = c(0.35, 0.55),
+    labels = scales::label_percent()) +
+  ## structure
[…]

### 2025-01-09 scripts/5_trust.R

@@ -1,130 +1,0 @@
-
-# - promises and pitfalls of panel change ------------------------------------ #
-# - case 3 # trust ----------------------------------------------------------- #
-
-### note: this script
-###       (a) takes `trust` as an examplary BES item,
-###       (b) simulates potential DGPs approximating the observed parameters,
-###       (c) shows the relative fit of the DGPs to observed data.
-
-rm(list = ls()) # clean-up
-
-# install.packages("devtools")
-# devtools::install_github("tkeskinturk/gridsearch")
-
-pacman::p_load(
-  purrr,
-  furrr,
-  srvyr,
-  hrbrthemes,
-  gridsearch,
-  tidyverse)
-theme_set(theme_ipsum_rc())
-
-# --- helpers
-my_oka <-
-  c("#eebd64",
-    "#8b3b36", 
-    "#a2b4b7")
-`%nin%` = Negate(`%in%`)
-
-# PART 1: BES DATA ----------------------------------------------------------- #
-
-# --- get data
-data <-
-  haven::read_dta("./data/britishelection.dta") |>
-  select(
-    pid = id,
-    trust_1 = genTrustW17,
-    trust_2 = genTrustW20,
-    trust_3 = genTrustW21,
-    trust_4 = genTrustW23,
-    trust_5 = genTrustW27
-  ) |>
-  haven::zap_labels() |>
-  mutate_at(
-    .vars = vars(matches("trust")),
-    .funs = ~ case_when(. == 1 ~ 1, . == 2 ~ 0, TRUE ~ NA_real_)
-  ) |>
-  drop_na() |>
-  pivot_longer(cols = -pid,
-               names_to = "var",
-               values_to = "y") |>
-  separate(var, into = c("var", "t")) |>
-  mutate(t = as.integer(t)) |> select(-var) |> 
-  mutate(
-    year =
-      case_when(
-        t == 1 ~ 2019,
-        t == 2 ~ 2020,
-        t == 3 ~ 2021,
-        t == 4 ~ 2022,
-        t == 5 ~ 2024
-      )
-  )
-
-# --- draw trajectories
-png(
-  "figures/trust01.png",
-  w = 7,
-  h = 5,
-  units = "in",
-  res = 500
-)
-glm(y ~ factor(year), data = data, family = "binomial") |>
-  ggeffects::ggemmeans(terms = "year") |>
-  as_tibble() |>
-  ## draw
-  ggplot(aes(
-    x = x,
[…]

### 2025-01-09 scripts/6_immigration.R

@@ -1,179 +1,0 @@
-
-# - promises and pitfalls of panel change ------------------------------------ #
-# - case 4 # immigration ----------------------------------------------------- #
-
-### note: this script
-###       (a) takes `imm` as an examplary panel data item,
-###       (b) simulates potential DGP distributions using fixed-effects models,
-###       (c) shows the relative fit of the DGPs to observed data.
-
-rm(list = ls()) # clean-up
-
-# install.packages("devtools")
-# devtools::install_github("tkeskinturk/gridsearch")
-
-pacman::p_load(
-  purrr,
-  furrr,
-  srvyr,
-  hrbrthemes,
-  gridsearch,
-  tidyverse,
-  patchwork)
-theme_set(theme_ipsum_rc())
-
-# --- helpers
-my_oka <-
-  c("#eebd64",
-    "#8b3b36", 
-    "#a2b4b7")
-`%nin%` = Negate(`%in%`)
-
-# PART 1: 26-WAVE DATA ------------------------------------------------------- #
-
-# --- read in
-data <- read_csv("./data/yllanon.csv")
-
-# --- cleaning up
-data <- data |>
-  ## organize
-  mutate(pid = cur_group_id(), .by = "id") |>
-  ## chop off post-COVID periods
-  filter(wave < 18) |>
-  ## simplify
-  select(pid, t = wave, y = imm) |>
-  ## clean-up the outcome
-  filter(y %nin% c(8, 9)) |> drop_na() |>
-  ## fully balanced
-  mutate(n = n(), .by = "pid") |> filter(n == 17) |>
-  ## scale the variables
-  mutate(y = ifelse(y > 4, 1, 0)) |>
-  mutate(y_std = (y - mean(y)) / sd(y)) |>
-  mutate(t = scales::rescale(t)) |>
-  ## select and wrap up organization
-  select(pid, t, y, y_std) |> arrange(pid, t)
-
-# --- reference
-d_slopes <- data |>
-  nest(.by = "pid") |>
-  mutate(m  = map(
-    .x = data, 
-    .f = ~ lm(y_std ~ t, data = .))) |>
-  mutate(lm = map_dbl(
-    .x = m, 
-    .f = ~ round(coef(.)[[2]], 3)))  |>
-  select(pid, lm)
-
-# PART 2: ILLUSTRATION ------------------------------------------------------- #
-
-set.seed(112358)
-d_example <- buildDGP(
[…]

### 2025-01-09 scripts/7_detection.R

@@ -1,147 +1,0 @@
-
-# - promises and pitfalls of panel change ------------------------------------ #
-# - change detection --------------------------------------------------------- #
-
-### note: this script
-###       (a) takes `grass`, `abany`, and `imm` as variable examples,
-###       (b) simulates potential DGPs approximating the observed parameters,
-###       (c) assesses whether we have leverage to detect individuals.
-
-rm(list = ls()) # clean-up
-
-# devtools (install if necessary!)
-# devtools::install_github("tkeskinturk/gridsearch")
-
-pacman::p_load(
-  purrr,
-  furrr,
-  hrbrthemes,
-  gridsearch,
-  tidyverse,
-  patchwork)
-theme_set(theme_ipsum_rc())
-
-# --- helpers
-my_oka <-
-  c("#eebd64",
-    "#8b3b36", 
-    "#a2b4b7",
-    "#4c6337")
-`%nin%` = Negate(`%in%`)
-
-# PART 1: DATA --------------------------------------------------------------- #
-
-grass <- readRDS("./data/output/grass.rds")
-abany <- readRDS("./data/output/abany.rds")
-trust <- readRDS("./data/output/trust.rds")
-immig <- readRDS("./data/output/immigration.rds")
-
-grass <- grass |> arrange(error) |> slice(1)
-abany <- abany |> arrange(error) |> slice(1)
-trust <- trust |> arrange(error) |> slice(1)
-immig <- immig |> arrange(error) |> slice(1)
-
-# --- grass
-grass_sim <-
-  simulateChangers(
-    n = grass |> pull(n),
-    t = 3,
-    rate = grass |> pull(rate),
-    balance_dir = 1,
-    balance_res = 0.47,
-    strength = grass |> pull(strength),
-    reliable = 0.91
-  )
-
-# --- abany
-abany_sim <-
-  simulateChangers(
-    n = abany |> pull(n),
-    t = 3,
-    rate = abany |> pull(rate), ## for help
-    balance_dir = 1,
-    balance_res = 0.45,
-    strength = abany |> pull(strength),
-    reliable = 0.86
-  )
-
-# --- trust
-trust_sim <-
-  simulateChangers(
-    n = trust |> pull(n),
-    t = 5,
-    rate = trust |> pull(rate),
-    balance_dir = 0.5,
-    balance_res = 0.43,
-    strength = trust |> pull(strength),
-    reliable = 0.88
-  )
-
-# --- imm
-immig_sim <-
-  simulateChangers(
[…]

### 2025-01-09 scripts/backups/confusionMat.R

@@ -1,0 +1,73 @@
+
+# - `gridsearch` ------------------------------------------------------------- #
+# - function: confusion matrix ----------------------------------------------- #
+
+### note: this function
+###       (a) takes a confusion matrix of true/simulated counts as its input,
+###       (b) calculates kappa, sensitivity, and specificity.
+
+# PART 1: FUNCTION ----------------------------------------------------------- #
+
+#' Calculate Accuracy Metrics
+#'
+#' This function takes a confusion matrix as its input and calculates sensitivity, specificity, and kappa scores for classification.
+#' @param confusion_matrix A confusion matrix.
+#'
+#' @return A data frame.
+#' @noRd
+#' @keywords internal
+#'
+
+confusionMatrix <-
+
+  function(confusion_matrix) {
+
+    # --- calculate kappa
+
+    # accuracy
+    p_0 =
+      (confusion_matrix[4, 3]
+       +
+       confusion_matrix[1, 3]
+      ) /
+      sum(confusion_matrix$n)
+
+    # expected classification
+    p_e =
+      (
+        (
+        (confusion_matrix[1, 3] + confusion_matrix[2, 3])
+        *
+        (confusion_matrix[1, 3] + confusion_matrix[3, 3])
+        )
+      +
+        (
+        (confusion_matrix[3, 3] + confusion_matrix[4, 3])
+        *
+        (confusion_matrix[2, 3] + confusion_matrix[4, 3])
+        )
+      ) /
+      (sum(confusion_matrix$n) ^ 2)
+
+    # kappa score
+    kappa = (p_0 - p_e) / (1 - p_e)
+
+    # --- calculate sensitivity
+    sensitivity =
+      confusion_matrix[4, 3] /
+      (confusion_matrix[4, 3] + confusion_matrix[3, 3])
+
+    # --- calculate specificity
+    specificity =
+      confusion_matrix[1, 3] /
+      (confusion_matrix[1, 3] + confusion_matrix[2, 3])
+
+    # --- export
+    output <- list(kappa = kappa$n,
+                   sensitivity = sensitivity$n,
+                   specificity = specificity$n)
+
+    return(output)
+  }
+
+# ---------------------------------------------------------------------------- #

### 2025-01-09 scripts/backups/simulations_reliability.R

@@ -1,0 +1,246 @@
+
+# - promises and pitfalls of panel change ------------------------------------ #
+# - script 2 # reliability space for `gridSearch` ---------------------------- #
+
+### note: this script
+###       (a) generates simulated DGPs with varying reliability scores,
+###       (b) applies the grid search algorithm to recover error distributions,
+###       (c) evaluates whether the algorithm can recover "true" DGP.
+
+rm(list = ls()) # clean-up
+
+# install.packages("devtools")
+# devtools::install_github("tkeskinturk/gridsearch")
+
+pacman::p_load(
+  purrr,
+  furrr,
+  hrbrthemes,
+  gridsearch,
+  ggforce,
+  tidyverse,
+  marginaleffects)
+theme_set(theme_ipsum_rc())
+
+# --- helpers
+my_oka <-
+  c("#eebd64",
+    "#8b3b36", 
+    "#a2b4b7")
+`%nin%` = Negate(`%in%`)
+
+# PART 1: DATA --------------------------------------------------------------- #
+
+set.seed(11235)
+
+# --- conditions
+d <-
+  expand_grid(
+    ## 500 individuals
+    p_n = 500,
+    ## varying panel waves (restrict to contingency)
+    p_t = 3,
+    ## the rate of change in the population
+    p_rate = c(.1, .25, .5),
+    ## the direction of change (halfsies versus secular)
+    p_dir = c(.5, 1),
+    ## the fixed level of marginal distributions
+    p_res = .5,
+    ## the strength of change in latent variable
+    p_strength = c(.5, 1),
+    ## the reliability score of the outcome measurement
+    p_reliable = c(.7, .9),
+    ## the number of simulated datasets
+    sims = 1:30)
+
+# --- dataframes
+d <- d |>
+  dplyr::mutate(
+    data =
+      furrr::future_pmap(
+        ## mapping list
+        .l = list(p_n,
+                  p_t,
+                  p_rate,
+                  p_dir,
+                  p_res,
+                  p_strength,
+                  p_reliable),
+        ## refer to list based on index
+        .f = purrr::possibly(
+          ~ buildDGP(
+            # varying parameters
+            n = ..1,
+            t = ..2,
+            rate = ..3,
+            balance_dir = ..4,
+            balance_res = ..5,
+            strength = ..6,
[…]

### 2025-01-11 paper/manuscript.Rmd
SV edits

@@
 
-Studies on political change at the individual level---change in one's beliefs and preferences---face a challenging problem: while our methods typically provide estimates for change and stability at the group level, our goal is often to understand these processes at the individual level.
+Studies on political change at the individual level---change in one's beliefs and preferences---face a challenging problem: while our methods typically provide estimates for change and stability at the population level, our goal is often to understand these processes at the individual level.
 
 This *group-to-person generalizability problem*, the claim that group-level findings may not generalize to each person [@mcmanus2023], has important implications for what we can and can't say about political change across the life-course.
-While the previous research quantified change with various methods---ranging from assessing the time order of longitudinal observations [@kiley2020; @vaisey2020] to multilevel decomposition of over-time variance [@lersch2023]---these studies have fallen short of disentangling whether they find a small amount of change in a large number of people, or a large amount of change in a small number of people.
-In other words, while social science researchers have become widely sophisticated in evaluating the average levels of change over time in a population, they have mostly overlooked the fact that these average changes are produced by a combination of various compositional processes.
+While the previous research quantified change with various methods---ranging from assessing the time order of longitudinal observations [@kiley2020; @vaisey2020] to multilevel decomposition of over-time variance [@lersch2023]---these studies have been unable to disentangle whether they find a small amount of change in a large number of people, or a large amount of change in a small number of people.
+In other words, while social science researchers have become sophisticated in evaluating the average levels of change over time in a population, they have mostly overlooked the fact that these average changes can be produced by different processes.
 
-This issue arises in a variety of ways in empirical practice, particularly when we want to theorize the mechanisms of political belief change.
+This issue arises in many ways in empirical research, particularly when we want to theorize the mechanisms of political belief change.
 For example, the General Social Survey (GSS) shows that roughly 35% of the American adults in 2006 believed that "marijuana should be legalized." In less than 10 years, this support had increased to nearly 60%.
@@
 While the cross-sectional processes have received strong scholarly attention in the literature [@vaisey2016; @restrepoochoa2024; @bartels2014], questions of the second sort remained largely unexplored.
-However, given that theories of political ideology involve theoretical expectations about individual-level outcomes, the fact that we lack information about whether our models capture person-level empirical processes is detrimental to understanding adult belief change in the political context.
+However, given that theories of political ideology involve theoretical expectations about individual-level outcomes, the fact that we lack information about whether our models capture person-level empirical processes is detrimental to understanding adult political belief change.
 Therefore, our article aims to address this gap by examining whether we can identify (a) *how many people changed*, (b) *how much they changed*, and (c) *who changed*, based on observed survey response trajectories of political beliefs in empirical data.
 
-Using simulations and empirical case studies, we show that it is indeed possible to provide plausible scenarios as to which generative models are more likely to have produced empirical observations, even though a reliable identification of *who* changed occurs only under very narrow conditions.
-In doing this, we build on the "Approximate Bayesian Computation" (ABC) literature [@beaumont2010] and provide a simple grid-search procedure to help social researchers in evaluating the extent to which their target estimates might be generalized to the population of their studies.
-Using this procedure, we show that it is possible to approximate to the underlying data generation processes for *rate of change*, *strength of change*, and *direction of change* in a panel data.
-We provide an R package, `gridsearch`, to help researchers in implementing this procedure in their own work.
+Using simulations and empirical case studies, we show that it is indeed possible to use panel data to provide plausible estimates of which generative models are more likely to have produced particular empirical observations.
+We also find, however, that a reliable identification of *who* changed occurs only under very narrow conditions.
+In this sense, the modeling individual change problem is very much like the age-period-cohort identification problem that arises in repeated cross-sectional data [CITATION TO FOSSE].
+In both cases, the problem cannot be "solved" by any particular methodological trick.
+Thus, our objective is not to provide a methodological panacea to eliminate the problem, but rather to provide a concrete approach that can help us understand the issue more clearly and manage it more effectively.
 
-We emphasize a striking heterogeneity in these underlying processes even for a seemingly straightforward case like the marijuana legalization in the GSS, showing that theoretically opposite claims---e.g., large swaths of the population increasing their support for the policy, as opposed to a small group of people completely switching their positions---may very well account for the same data.
-This finding is particularly important in the longitudinal panel context, which, for most applications, represents the gold standard in evaluating change in observational social research.
-As such, we claim that when the meaning of "political change" is elusive, and depends too much on how we define change in the first place, our current empirical strategies could potentially mislead rather than inform, especially if we overlook the underlying generative mechanisms.
+This approach builds on the "Approximate Bayesian Computation" (ABC) literature [@beaumont2010] and provides a simple grid-search procedure to help researchers evaluate the extent to which their target estimates might be generalized to a study population.
+Using this procedure, we show that it is possible to approximate to the underlying data generation processes for *rate of change*, *strength of change*, and *direction of change* in a panel dataset.
+We provide an R package, `gridsearch`, to help researchers implement this procedure in their own work.
 
-We start with a general overview of the "latent variable model" of beliefs, the problems of reliability and resolution in survey measurement, and the compositional sources of aggregate change.
+We emphasize the existence of striking heterogeneity in these underlying processes even for a seemingly straightforward case like the marijuana legalization in the GSS; that is, we show that theoretically opposite claims---e.g., large swaths of the population increasing their support for the policy, as opposed to a small group of people completely switching their positions---may very well account for the same data.
+This finding is particularly important in the longitudinal panel context, which, for most applications, represents the gold standard in evaluating change in observational research.
+For this reason, we argue that when the meaning of "political change" is not clearly specified, our current empirical strategies could potentially mislead rather than inform, especially if we overlook the underlying generative mechanisms.
+
+We begin with a general overview of the "latent variable model" of beliefs, the problems of reliability and resolution in survey measurement, and the compositional sources of aggregate change.
 We then introduce our grid-search procedure in the context of three-wave panel data with binary outcomes.
@@
 
-The latent-variable approach reveals a fundamental problem in survey research: one's true score $Y_i$ is almost always realized with a certain degree of measurement error.
-Call this realization the *observed score* $y_i$.
+The latent-variable approach acknowledges a fundamental problem in survey research: one's true score $Y_i$ is almost always realized with a certain degree of measurement error.
+Call this realization the *observed score*, $y_i$.
 We often assume that $y_i$ represents the expected value of hypothetical draws from $Y_i$, with an $\epsilon$ term reflecting the measurement error in this process---i.e., $y_i = Y_i + \epsilon_i$.
-This error might result from various factors that introduce biases to how we extract information from individuals, given that problems like question construction and interviewer effects might contaminate how individuals reveal their true preferences [@alwin1991].
-The inherent stochasticity in this measurement process implies that each draw from $Y_i$ might at least slightly differ from the true score, and this error term is difficult to circumvent with one measure, which is often the case in public opinion surveys.[^scales] In this sense, there is often a degree of systematic gap between $Y_i$ and $y_i$.
+This error might result from various factors that introduce biases in how we extract information from individuals, given that problems like question construction and interviewer effects might influence how individuals reveal their true preferences [@alwin1991].
+The inherent randomness in the measurement process implies that each draw from $Y_i$ might at least slightly differ from the true score, and this error term is difficult to circumvent with a one-item measure, which is often the case in large public opinion surveys.[^scales] In this sense, there is often a degree of systematic gap between $Y_i$ and $y_i$.
 
@@
 
-In the panel context, where we observe the same people multiple times, this measurement problem has substantive implications: if true scores are imperfectly realized and we only have surface access to a person's observed survey responses, how can we know if they *actually* changed between time $t$ and $t + 1$, or whether the observed change is merely due to measurement error?
+In the panel context, where we observe the same people multiple times, this measurement problem has substantive implications: if true scores are imperfectly realized and we only have access to a person's observed survey responses, how can we know if they *actually* changed between time $t$ and $t + 1$, or whether the observed change is merely due to measurement error?
 
 Scholars in public opinion research have attempted to address this issue in multiple ways, using structural equation models that explicitly quantify the measurement error component [@achen1975], strategies that integrate longitudinal surveys to exploit the information in the ordering of responses [@kiley2020], and various multilevel mixed-effects models that shrink responses based on population estimates to derive reliable trends [@lersch2023; @hout2016].
-To some extent, all these strategies were fruitful to offset the problems associated with measurement error.[^stability] That said, while these empirical strategies were largely successful in accounting for the amount of change *in the population*, they were rather silent about changes *at the level of individuals*.
+To some extent, all these strategies have been useful for offseting the problems associated with measurement error.[^stability] That said, while these empirical strategies have been largely successful in accounting for the amount of change *in the population*, they were rather silent about changes *at the level of individuals*.
 
@@
 Let $y_{it}$ represent an observed survey response.
-To see how much observed change there is at the individual level, we can simply define a change quantity, $\delta y_{i, t_0 \to t_1} = y_{it_1} - y_{it_0}$, the average of which, $\bar{y}$, represents the mean change in the dataset from $t_0$ to $t_1$.
+To see how much observed change there is at the individual level, we can simply define a change quantity, $\delta y_{i, t_0 \to t_1} = y_{it_1} - y_{it_0}$, the average of which, $\bar{\delta_y}$, represents the mean change in the dataset from $t_0$ to $t_1$.
 
@@
 Let us define these processes intuitively.
-The first component is $R$, representing the proportion of individuals in our dataset with $\delta y_{i, t_0 \to t_1} \neq 0$, i.e., the proportion of individuals who changed in this period.
-Call this *the rate of change*.
-The second component is $S$, representing the "average" magnitude of change among those who changed.[^heterogeneity] Call this *the strength of change*.
-The third component is $D$, representing the net directionality of change, i.e., the difference in proportions of positive and negative changers in our observations.
-Call this *the direction of change*.
+The first component is $R$, or the *rate of change*.
+This represents the proportion of individuals in our dataset with $\delta y_{i, t_0 \to t_1} \neq 0$, i.e., the proportion of individuals who changed in this period.
+The second component is $S$, the *strength of change*.
+This represents the "average" magnitude of change among those who changed.[^heterogeneity] The third component is $D$, the *direction of change*.
+This represents the net directionality of change, i.e., the difference in proportions of positive and negative changers in our observations.
 If we combine these three components, we can represent $\bar{y}$ simply as
@@
 $$
-\bar{y} = R \times S \times D
+\bar{\delta_y} = R \times S \times D
 $$
@@
 [^heterogeneity]: Of course, we might further decompose this average strength to a distribution of heterogeneous change trajectories across individuals.
-For analytical and expository clarity, we bracket this issue from the scope of this article.
+For analytical and expository clarity, we bracket this issue from in this article.
 
@@
 
-For instance, if we want to evaluate whether a society became more pro-redistribution over time, we need to note that an increase in pro-redistributive preferences can occur via all or most people reporting higher preference for redistribution, some people switching their positions altogether while some others remain completely stable, or everyone changing in one or the other direction, and the net score we observe is simply the difference in their overall trajectories.
-The same reasoning applies to the GSS's marijuana legalization policy that we have discussed before.
+For instance, if we want to evaluate how a society became more pro-redistribution over time, we need to note that an increase in pro-redistributive preferences could occur via all or most people reporting higher preference for redistribution, some people switching their positions altogether while some others remain completely stable, or everyone changing in one or the other direction, and the net score we observe is simply the difference in their overall trajectories.
+The same reasoning applies to the GSS's marijuana legalization policy we discussed above.
 Most discussions in political belief change---confidence in legislative groups or financial institutions, beliefs about the economy, or ethnic attitudes---have the exact same question about their data generation processes.
-Simply put, the findings about change at the population level may simply result from a variety of conflicting, though observationally equivalent, data generation processes.
+Simply put, the findings about individual change at the population level may simply result from a variety of incompatible, though observationally equivalent, data generation processes.[^APC]
 
-We can now complicate this composition question by noting, once again, the survey measurement problem of mapping one's true scores to their observed scores, which often fails in at least two ways: (1) our observed measures might have low *reliability*, such that a hypothetical set of repeated draws may significantly differ from one's true score, due to, e.g., poor question construction or individual idiosyncrasies, and (2) question response categories may be imperfect: the *resolution* of response categories may fail to distinguish consequential differences in true scores.
+[^APC]: Again, this is a very similar problem to the age-period-cohort identification problem we mentioned above.
+
+As if this weren't complicated enough, we must complicate the issue further by, once again, recognizing the measurement problem of mapping true scores to observed scores.
+This process often fails in at least two ways: (1) our observed measures might have low *reliability*, such that a hypothetical set of repeated draws may significantly differ from one's true score, and (2) question response categories may be imperfect: the *resolution* of response categories may fail to distinguish consequential differences in true scores.
 Therefore, the average change captured in $\bar{y}$ may represent a biased or imperfect estimate of change to begin with.
@@
 
-We pose these questions in a fixed-effects longitudinal panel data context [@allison2009; @vaisey2017], with varying individual trajectories across time periods [@rüttenauer2020].
+We pose this question in a fixed-effects longitudinal panel data context [@allison2009; @vaisey2017], with varying individual trajectories across time periods [@rüttenauer2020].
 This allows us to shift the focus from *aggregate change* to *individual change*, given that our theoretical interest is to understand how people change their political beliefs over the life-course.
@@
 
-We build on the Approximate Bayesian Computation (ABC) literature [@beaumont2010] and propose a grid-search procedure to answer our research questions.
+We build on the Approximate Bayesian Computation (ABC) literature [@beaumont2010] and propose a grid-search procedure to answer this research question.
 The ABC approach involves simulating parameter values given the data and evaluating the distribution of these parameters that potentially generated our observations.
@@
 
-3) Compares the simulated trajectories to the observed trajectories, calculating a "distance" score to quantify the extent to which specific simulated values reproduce the observed values;
+3) Compares the distribution of simulated trajectories to the distribution of observed trajectories, calculating a "distance" score to quantify the extent to which specific simulated values reproduce the observed values;
 

### 2025-01-12 paper/manuscript.Rmd
TK edits.

@@
 abstract: "We investigate whether studies on political belief change can identify change trajectories at the individual level.
-Using simulations and case studies, we propose a grid-search framework that allows researchers to evaluate the extent to which their target estimates generalize to individuals in their study population.
-We use simulated datasets to estimate plausible values for how many people changed, how much they changed, and who changed, based on observed survey response trajectories.
+Using simulations and case studies, we propose a grid-search framework that allows researchers to evaluate the extent to which their target estimates generalize to a study population.
+We use simulated datasets to estimate plausible values for how many people changed, how much they changed, and who changed, based on observed response trajectories.
 Our results suggest that researchers should think carefully about the conditions under which they may make claims about belief change at the individual level.
@@
 This *group-to-person generalizability problem*, the claim that group-level findings may not generalize to each person [@mcmanus2023], has important implications for what we can and can't say about political change across the life-course.
-While the previous research quantified change with various methods---ranging from assessing the time order of longitudinal observations [@kiley2020; @vaisey2020] to multilevel decomposition of over-time variance [@lersch2023]---these studies have been unable to disentangle whether they find a small amount of change in a large number of people, or a large amount of change in a small number of people.
+While the previous research quantified change with various methods---ranging from assessing the time order of longitudinal observations [@kiley2020; @vaisey2020] to multilevel decomposition of over-time variance [@lersch2023]---these studies have been unable to disentangle whether they find a small amount of change in a large number of people, a large amount of change in a small number of people, or a mix of positive and negative trajectories.
 In other words, while social science researchers have become sophisticated in evaluating the average levels of change over time in a population, they have mostly overlooked the fact that these average changes can be produced by different processes.
@@
 However, given that theories of political ideology involve theoretical expectations about individual-level outcomes, the fact that we lack information about whether our models capture person-level empirical processes is detrimental to understanding adult political belief change.
-Therefore, our article aims to address this gap by examining whether we can identify (a) *how many people changed*, (b) *how much they changed*, and (c) *who changed*, based on observed survey response trajectories of political beliefs in empirical data.
+Our article aims to address this gap by examining whether we can identify (a) *how many people changed*, (b) *how much they changed*, and (c) *who changed* using empirical data.
 
@@
 We also find, however, that a reliable identification of *who* changed occurs only under very narrow conditions.
-In this sense, the modeling individual change problem is very much like the age-period-cohort identification problem that arises in repeated cross-sectional data [CITATION TO FOSSE].
+Hence, the modeling individual change problem is very much like the age-period-cohort identification problem that arises in repeated cross-sectional data [@fosse2019].
 In both cases, the problem cannot be "solved" by any particular methodological trick.
@@
 This approach builds on the "Approximate Bayesian Computation" (ABC) literature [@beaumont2010] and provides a simple grid-search procedure to help researchers evaluate the extent to which their target estimates might be generalized to a study population.
-Using this procedure, we show that it is possible to approximate to the underlying data generation processes for *rate of change*, *strength of change*, and *direction of change* in a panel dataset.
-We provide an R package, `gridsearch`, to help researchers implement this procedure in their own work.
+Using this procedure, we show that it is possible to approximate to the set of underlying data generation processes for *rate of change*, *strength of change*, and *direction of change* in a panel dataset.
+We provide an R package, "gridsearch," to help researchers implement our approach in their own empirical applications.
 
-We emphasize the existence of striking heterogeneity in these underlying processes even for a seemingly straightforward case like the marijuana legalization in the GSS; that is, we show that theoretically opposite claims---e.g., large swaths of the population increasing their support for the policy, as opposed to a small group of people completely switching their positions---may very well account for the same data.
-This finding is particularly important in the longitudinal panel context, which, for most applications, represents the gold standard in evaluating change in observational research.
+We emphasize the existence of striking heterogeneity in these underlying generative processes even for a seemingly straightforward case like the marijuana legalization in the GSS; that is, we show that theoretically opposite claims---e.g., large swaths of the population increasing their support for the policy, as opposed to a small group of people completely switching their positions---may very well account for the same data.
+This finding is particularly significant in the longitudinal panel context, which, for most applications, represents the gold standard in evaluating change in observational research.
 For this reason, we argue that when the meaning of "political change" is not clearly specified, our current empirical strategies could potentially mislead rather than inform, especially if we overlook the underlying generative mechanisms.
@@
 We begin with a general overview of the "latent variable model" of beliefs, the problems of reliability and resolution in survey measurement, and the compositional sources of aggregate change.
-We then introduce our grid-search procedure in the context of three-wave panel data with binary outcomes.
-Next, we relax this assumption and investigate how linear trajectories at the individual level can assist us in contexts with more than three waves.
-In the final section, we analyze whether this procedure helps us in classification tasks, particularly in evaluating *who* changed.
+We then introduce our grid-search approach with an illustrating exercise, provide empirical examples, and present a simulation study to explore its implications.
+Next, we analyze whether this procedure helps us in classification tasks, particularly in evaluating *who* changed.
 We conclude by discussing the promises and pitfalls of panel data for understanding individual political change.
 
-# Measuring Change in Political Beliefs
+# Measuring Change in Political Beliefs and Preferences
 
@@
 This error might result from various factors that introduce biases in how we extract information from individuals, given that problems like question construction and interviewer effects might influence how individuals reveal their true preferences [@alwin1991].
-The inherent randomness in the measurement process implies that each draw from $Y_i$ might at least slightly differ from the true score, and this error term is difficult to circumvent with a one-item measure, which is often the case in large public opinion surveys.[^scales] In this sense, there is often a degree of systematic gap between $Y_i$ and $y_i$.
-
-[^scales]: The idea that we should have repeated measures of $Y_i$ is the main intuition behind scale construction [@ansolabehere2008], with the explicit expectation that multiple measures can help us reduce measurement error.
+The inherent randomness in the measurement process implies that each draw from $Y_i$ might at least slightly differ from the true score, and this error term is difficult to circumvent with a one-item measure, which is often the case in large public opinion surveys.
+Therefore, there is often a degree of systematic gap between $Y_i$ and $y_i$.
 
@@
 This represents the net directionality of change, i.e., the difference in proportions of positive and negative changers in our observations.
-If we combine these three components, we can represent $\bar{y}$ simply as
+If we combine these three components, we can represent $\bar{\delta_y}$ simply as
 
@@
 [^heterogeneity]: Of course, we might further decompose this average strength to a distribution of heterogeneous change trajectories across individuals.
-For analytical and expository clarity, we bracket this issue from in this article.
+For analytical and expository clarity, we bracket this issue in this article.
 
@@
 
-As if this weren't complicated enough, we must complicate the issue further by, once again, recognizing the measurement problem of mapping true scores to observed scores.
-This process often fails in at least two ways: (1) our observed measures might have low *reliability*, such that a hypothetical set of repeated draws may significantly differ from one's true score, and (2) question response categories may be imperfect: the *resolution* of response categories may fail to distinguish consequential differences in true scores.
-Therefore, the average change captured in $\bar{y}$ may represent a biased or imperfect estimate of change to begin with.
+If we were to believe that *each and every single survey response* represents an accurate mapping of true scores and observed scores, we could seamlessly solve this problem by simply looking at the empirical data.
+But we contend that the measurement problem does not allow us to do that.
+The process often fails in at least two ways: (1) our observed measures might have low *reliability*, such that a hypothetical set of repeated draws may significantly differ from one's true score, and (2) the question response categories may be imperfect: the *resolution* of responses may fail to distinguish consequential differences in true scores.
+Therefore, the average change captured in $\bar{\delta_y}$ may represent a biased or imperfect estimate of change to begin with.
 
@@
 
-## A Grid-Search Procedure
+## A Grid-Search Procedure to Adjudicate Data Generation Processes
 
@@
 The ABC approach involves simulating parameter values given the data and evaluating the distribution of these parameters that potentially generated our observations.
-We build a rejection sampler algorithm that
+We build an algorithm that
 
-1) Simulates data with several parameters of interest for our DGPs---rate of change, strength of change, direction of change, and reliability of response measurement using
+1) Simulates data with several parameters of interest for our DGPs---rate of change, strength of change, direction of change, and reliability of response measurement---using
 
@@
 
-2) Calculates the distribution of response trajectories in both the observed and simulated data, by either tabulating response patterns across time or estimating slope coefficients;
+2) Calculates the distribution of change scores in observed and simulated data, by estimating respondent-level change scores using survey response trajectories at the person level.
+To put it differently, it first fits, for a binary outcome,
 
-3) Compares the distribution of simulated trajectories to the distribution of observed trajectories, calculating a "distance" score to quantify the extent to which specific simulated values reproduce the observed values;
+   $$
+   \mathbb{P}(y = 1) = \beta_0 + \beta_1t, \ \ t \in [0, 1]
+   $$
+
+   where \( t \) is a normalized time variable, and estimates a predicted change score by calculating the difference between the prediction at \( t = 1 \) and the prediction at \( t = 0 \).
+
+3) Compares the distribution of simulated trajectories to the distribution of observed trajectories by calculating a "distance" score---specifically, Kolmogorov-Smirnov distance---to quantify the extent to which specific simulated values reproduce the observed values;
+
+4) Replicates this analysis $N$ times by drawing samples for our main parameters;
+
+5) Inspects the parameter combinations that most closely approximate to the observed values.
 
 [^changing]: Note that we "allow" actors to go through change at any time in the observation window of the study, meaning that $D_{it}$ is variable over $t$ precisely because actors can change any time in their biographical history.
-
-4) Rejects the simulated data if the distance measure is above a certain pre-specified "caliper" value, and accepts it otherwise;
-
-5) Replicates this analysis $N$ times by drawing samples from the specified priors;
-
-6) Inspects the parameter combinations that most closely approximate to the observed values.
 
@@
 
-[^parameters]: We define additional parameters that we fix in the case studies presented below: $N$ refers to the number of actors in the fake-data simulation, $T$ refers to the number of waves we have in the panel study, *balance* refers to the observed marginals in the real datasets, allowing us to "cut" the continuous distribution to 0s and 1s, and *resolution* refers to the number of categories to simplify the latent variable, which is restricted to the binary case in this article.
+[^parameters]: We define additional parameters that we fix in the case studies presented below: $N$ refers to the number of actors in the simulation, $T$ refers to the number of waves we have in the panel study, *balance* refers to the observed marginals in the real datasets, allowing us to "cut" the continuous distribution to 0s and 1s, and *resolution* refers to the number of categories to simplify the latent variable, which is restricted to the binary case in this article.
 
-In the next four sections, we apply this procedure in different contexts.
-In all cases, we restrict our attention to latent variables with binary outcomes to simplify our analyses.
-The replication files for the article are stored in https://osf.io/AAAAA.
-The R package for the use of this approach, `gridsearch`, can be accessed via the replication repo or https://tkeskinturk.github.io/gridsearch/.
+## Approximating Observed Data via Simulations
 
-\newpage
-
-# Study 1: Three-Wave Panel Context
-
-In the three-wave panel context with binary outcomes, the response trajectories are easy to inspect, given that there are only $2^3 = 8$ possible response patterns for each person:
-
-$$
-\bigg( \{0,0,0\}, \{0,0,1\}, \{0,1,0\}, \{0,1,1\}, \{1,0,0\}, \{1,0,1\}, \{1,1,0\}, \{1,1,1\} \bigg)
-$$
-
-To examine the distribution of these trajectories in the observed data, we can count the number of respondents who go through trajectory $j$.
-For instance, the marijuana legalization example from Figure 1 has the following contingency table of counts:
-
-$$
-\begin{bmatrix}
-0, 0, 0 \\
-0, 0, 1 \\
-0, 1, 0 \\
-0, 1, 1 \\
-1, 0, 0 \\
-1, 0, 1 \\
-1, 1, 0 \\
-1, 1, 1
-\end{bmatrix} 
-=
-\begin{bmatrix}
-869 \\
-153 \\
-75 \\
-115 \\
-65 \\
-71 \\
-54 \\
-746 \\
-\end{bmatrix}
-$$
-
-Approximately 75% of the respondents remain stable, following trajectories of $\{0,0,0\}$ and $\{1,1,1\}$, while others appear to *change* in some respect.
-The question is the extent to which this contingency table can be approximated by an underlying generative model, where we know the true change parameters.
-
-Let us provide one such example.
-We simulate a panel dataset, with 20% of the respondents changing their beliefs on marijuana legalization, operationalized as a latent construct with $\mu = 0$ and $\sigma = 1$, by 2 standard deviations in the "negative" direction---note that this is the opposite of the mean pattern observed in Figure 1.
-We fix all the parameters in the DGP to their observed values: 2,148 individuals observed over 3 periods, with a reliability score of 0.91 [see @hout2016 for the calculation of this value].
-We know that the marginals were equal to 0.47 in the entire panel, so we constrain the binary outcomes such that the final distribution of $y$ equals to 0.47 as well.
-Here are the generated counts for all trajectories, with differences from the original data appended:
-
-$$
-\begin{bmatrix}
-\\
-0, 0, 0 \\
-0, 0, 1 \\
-0, 1, 0 \\
-0, 1, 1 \\
-1, 0, 0 \\
-1, 0, 1 \\
-1, 1, 0 \\
-1, 1, 1
-\end{bmatrix} 
-=
-\begin{bmatrix}
-\text{Observed} \\
-869 \\
-153 \\
-75 \\
-115 \\
-65 \\
-71 \\
-54 \\
-746 \\
-\end{bmatrix}
--
-\begin{bmatrix}
-\text{Simulated} \\
-810 \\
-75 \\
-103 \\
-57 \\
-188 \\
-66 \\
-148 \\
-701 \\
-\end{bmatrix}
-=
-\begin{bmatrix}
-\text{Absolute Difference} \\
-59 \\
-78 \\
-28 \\
-58 \\
-123 \\
-5 \\
-94 \\
-45 \\
-\end{bmatrix}
-$$
-
-The generative model does a pretty poor job of reproducing the contingency table.
-23% of the respondents ($N = 490$) are assigned to the wrong trajectory.
-This means that we should be able to improve our predictions significantly if we switch to the true data generation model.
-Unfortunately, we don't know it; the set of true change parameters is the exact thing we want to derive using observed trajectories.
-
-How can we improve our prediction accuracy?
-Our grid-search procedure allows us to estimate plausible approximation values across a grid of potential generative models.
-Let us now apply this procedure to the case of marijuana legalization to see whether we can retrieve plausible DGPs.
-
-## The Case of Marijuana Legalization
-
-We first define 2,100 potential generative models using a combination of (1) rate of change, where we vary values from 0 to 1 in increments of 0.05, (2) strength of change, where we vary values from 0 to 2 in increments of 0.1, and (3) five directional possibilities---all negative, all positive, half negative and half positive, 75% negative and 25% positive, and 75% positive and 25% negative.
-This results in $21 \times 20 \times 5 =$ 2,100 cells, each representing one potential data-generation process that might have produced the observed data.
-We then generate corresponding panel datasets based on each model.
-Due to the stochastic nature of our fake-data simulations, we repeat this procedure 1,000 times for each cell, resulting in 2,100,000 datasets for which we evaluate the observed trajectories.
-
-Once we have our fake-data simulations, we count the number of respondents assigned to each response trajectory, calculate the absolute errors relative to the observed data, and normalize these counts by dividing the error by the total number of respondents.
-We then average these percent errors across 1,000 simulations to obtain stable estimates.
-Figure 2 presents the results.
+Let us flesh out our approach by walking through an illustrative example of how we approximate to the observed empirical responses.
+Suppose that we *know* the underlying DGP of a dataset of interest.
+In this dataset, called DGP 1, 500 people are observed over a period of 12 waves.
+We know that 50% of these respondents make certain changes to their latent positions, and these changes are all in the positive direction with a strength of 1 SD.
+However, we also know that the item reliability is 80%, which means that, across time, we sometimes miss respondents' true scores.
+The top-left panel of Figure 2 shows the distribution of change scores from this dataset.
 
@@
 ```{r figure2, out.width = "100%"}
-knitr::include_graphics("../figures/grass02.png")
+knitr::include_graphics("../figures/ks_distance.png")
 ```
@@
 \end{center}
-\caption*{Figure 2: Fake-Data Simulations for Marijuana Legalization}
-\footnotesize{\textit{Notes:} The figure depicts the extent to which specific parameter combinations approximate the observed data.
-Each point represents our distance measure, calculated as the sum of absolute errors between observed and simulated contingency tables, where darker colors represent smaller distances.
-We used 5\%, 10\%, and 20\% distance cuts for color shades, along with a second filter using quantiles of 0.05, 0.10, and 0.20 at each level.
-The shading thus reflects both absolute and relative accuracy among the simulations.
-The single yellow point represents the parameter combination that best approximates the observed data in the General Social Survey.}
+\caption*{Figure 2: Calculating Distance for Observed and Simulated Trajectories}
+\footnotesize{\textit{Notes:} The figure presents an illustrative example of two known DGPs: (1) the top-left panel shows the distribution of change scores from DGP 1, (2) the bottom-left panel shows the distribution of change scores from DGP 2, and (3) the right panel shows the empirical cumulative distribution functions from each of these distributions.}
 \end{figure}
@@
 
-There are several notable findings.
-First, error values range from 4.6% to 69.6%, with an interquartile range of 10.6% and 24%.
-This indicates that many cases fail to achieve a reasonable approximation, but we still end up with reliably close generative models.
-Second, the best value occurs when the rate of change equals to 25%, with a positive strength of 0.9 standard deviation change (SD).
-That said, note the black values clustered around the yellow cell in the bottom right panel, where cells range from 25% to 35% rate of change, with strength ranging, once again, from 0.7 to 1 SD.
-These patterns suggest that the most plausible data generation process occurs when roughly 2 to 4 out of 10 people positively change their approvals by roughly 1 SD.
-Finally, as expected, there are no plausible estimates where change occurs in the negative direction, providing a plausibility check for our proposed method.
+Assume that this dataset is our *observed* dataset.
+We will now see whether an alternative dataset, called DGP 2, can approximate to its observed parameters.
+From DGP 1, we know that there are 500 individuals, observed across 12 waves.
+We also know the marginal distribution of our response variable.
+We can treat these parameter values fixed for our new simulation.
+However, we do not know several other parameters: we do not know the *real* rate of change, nor do we know the *real* strength, direction, or reliability.
+Let us assume, for this specific iteration, a rate of 50% (correct), a strength of 1 SD (correct), a reliability score of 80% (correct), and an equally split direction, half changing positively and half changing negatively (incorrect).
+The bottom-left panel in Figure 2 shows the distribution of change scores from this particular DGP.
 
-These findings demonstrate the utility of our approach. By iterating over DGPs, we can identify plausible values that may have produced the observed change trajectories in the panel context.
+How adequately does DGP 2 represent DGP 1?
+In the right panel of Figure 2, we show the empirical cumulative distribution functions (ECDFs) for our two distributions, which indicates that DGP 2 does a very poor job of reproducing DGP 1, with a Kolmogorov-Smirnov distance of 0.16.
+This means that we should be able to improve our predictions significantly if we start tweaking some parameters for DGP 2.
+Unfortunately, we of course do not know which parameters to tweak; the set of true change parameters is the exact thing we want to derive using observed trajectories.
+This leads us to consider a wide variety of parameter values that minimize the Kolmogorov-Smirnov distance scores between the observed trajectories and the said simulated trajectories.
 
-## The Case of the Legality of Abortion
+In the following sections, we apply this procedure in different contexts.
+In all cases, we restrict our attention to latent variables with binary outcomes to simplify our analyses.
+The replication files for the article are stored in https://osf.io/AAAAA.
+The R package for the use of this approach, "gridsearch," can be accessed via the replication repo or https://anonymized.github.io/gridsearch/.
 
-What about cases where the direction or even existence of change is less clear-cut?
-We now turn our attention to the legality of abortion debate, a contentious issue characterized by high disagreement and high political sorting in the American context [@baldassarri2008; @hout2022].
-In 2006, 2008, and 2010 GSS panels, respondents agreed or disagreed with the statement: "it should be possible for a pregnant woman to obtain a legal abortion if the woman wants it for any reason." Figure 3 shows response trajectories for this question across three panels, indicating a clear stasis in the marginals.
+# Empirical Case Studies
+
+## Exploring the Case of Marijuana Legalization
+
+Let us now return to the marijuana legalization case in the GSS and apply our grid-search procedure to see whether we can detect a set of plausible DGPs for our empirical data.
+
+We first fit 100,000 simulations by drawing samples from four different parameters: *rate of change*, fixed between 0 and 1, *strength of change*, fixed between 0 SD and 2 SD, *balance of change*, fixed between 0 and 1, representing all negative change at 0 and all positive change at 1, and a *reliability score*, fixed between 60% and 100%.
+We use uniform priors for each parameter.
+We then calculate the Kolmogorov-Smirnov (KS) distance for each draw and keep the 2% quantile of the DGPs that minimize the KS distance, ending up with 2,000 potential generative models.
+
+Figure 3 shows the distribution of, and bivariate relationships among, these grid-search parameters, suggesting that the posterior distributions of strength and rate parameters are highly variable---and strongly correlated---while the posterior distributions for balance and reliability much more concentrated.
+Looking at the distributions, the interquartile ranges for strength and rate are 0.63 and 0.42, respectively, while these values are only 0.17 for balance and 0.07 for reliability.
 
@@
 ```{r figure3, out.width = "100%"}
+knitr::include_graphics("../figures/grass02.png")
+```
+```{=latex}
+\end{center}
+\caption*{Figure 3: Pair Plots for Marijuana Legalization}
+\footnotesize{\textit{Notes:} The figure depicts the distribution of, and bivariate relationships among, grid-search parameter values from the best-fitting DGPs for the case of marijuana legalization (N = 2,000 DGPs from 100,000 simulations).}
+\end{figure}
+```
+
+These patterns make sense when we examine the conditional relationships among our parameters, presented in Figure 4.
+We see that the algorithm can easily suggest, at least for this particular case, a specific direction estimate---a median balance of 0.82 with an IQR of 0.73 and 0.90---but it converges to almost analytically determined bands in high-balance panels: we detect either small changes in a large number of people, or large changes in a small number of people.
+To put this result in perspective, note that the KS distance among these accepted draws are virtually indistinguishable from one another: for instance, the "best" DGP with 40% rate and 0.90 SD strength is within less than 0.001 KS value of a DGP with 20% rate and 1.80 SD strength.
+
+```{=latex}
+\begin{figure}[!ht]
+\begin{center}
+```
+```{r figure4, out.width = "100%"}
+knitr::include_graphics("../figures/grass03.png")
+```
+```{=latex}
+\end{center}
+\caption*{Figure 4: Best-Fitting DGPs for Marijuana Legalization}
+\footnotesize{\textit{Notes:} The figure depicts the best fitting 2,000 DGPs (from 100,000 simulations) for the empirical case study.
+Each point represents an accepted draw.
+We rounded the balance parameters to show conditional relationships.}
+\end{figure}
+```
+
+There is one strong implication of this exercise: there are competing generative models that may easily explain the same aggregate patterns, and there is no methodological trick to solve this issue.
+If we take the measurement error problem to its natural conclusions, the same data can be explained by incompatible theoretical mechanisms, and our conclusions may be wildly off-the-mark.
+
+## Exploring the Case of the Legality of Abortion
+
+What about cases where the direction or even existence of change is less clear-cut?
+We now turn our attention to the legality of abortion debate, a contentious issue characterized by high disagreement and high political sorting in the American context [@baldassarri2008; @hout2022].
+In 2006, 2008, and 2010 GSS panels, respondents agreed or disagreed with the statement: "it should be possible for a pregnant woman to obtain a legal abortion if the woman wants it for any reason." Figure 5 shows response trajectories for this question across three panels, indicating a clear stasis in the marginals.
+
+```{=latex}
+\begin{figure}[!ht]
+\begin{center}
+```
+```{r figure5, out.width = "100%"}
 knitr::include_graphics("../figures/abany01.png")
@@
 \end{center}
-\caption*{Figure 3: The Percent Support for Abortion}
+\caption*{Figure 5: The Percent Support for the Legality of Abortion}
 \footnotesize{\textit{Notes:} The figure depicts the percent of American adults saying "it should be possible for a pregnant woman to obtain a legal abortion if the woman wants it for any reason" in the General Social Survey 2006, 2008, and 2010 panel studies.
@@
 
-However, the contingency table for trajectories shows more heterogeneity.
-Again, there is considerable stability---approximately 73% of respondents remain stable, following trajectories of $\{0,0,0\}$ and $\{1,1,1\}$, similar to the case of marijuana legalization.
-But there is more homogeneity in terms of directional change: while trajectories like $\{0,0,1\}$ and $\{0,1,1\}$, cases indicating an upward trend, were 2.27 times more likely in the marijuana legalization case, we now observe similar numbers in the abortion case---230 clear upward trends and 209 clear downward trends.
-This suggests that the directional heterogeneity of political change might have cancelled the trends in the aggregate, resulting in stable marginals in the observed frequencies.
-
-$$
-\begin{bmatrix}
-0, 0, 0 \\
-0, 0, 1 \\
-0, 1, 0 \\
-0, 1, 1 \\
-1, 0, 0 \\
-1, 0, 1 \\
-1, 1, 0 \\
-1, 1, 1
-\end{bmatrix} 
-=
-\begin{bmatrix}
-972 \\
-115 \\
-101 \\
-115 \\
-118 \\
-99 \\
-91 \\
-733 \\
-\end{bmatrix}
-$$
-
-Figure 4 presents the results from our grid-search exercise, implemented across 2,100,000 datasets with 2,100 generative models.
-We set the reliability to .86 following @hout2016.
-This time, the estimates are much closer to the observed frequencies, with error values ranging from 3% to 64.5%, and an interquartile range of 5.1% and 23.5%.
-That said, there is strong variability in the plausible generative scenarios.
-Note how, in the top right corner, there is a plausibility bound ranging from a rate of change as low as 10% to as high as 100%, and strength of change ranging from as low as 0.3 SD to as high as 2 SD, indicating a variety of scenarios.
-In fact, the top 10 generative models out of 2,100 DGPs have error rates within a margin of 0.04%, meaning that it is difficult to distinguish the predictive abilities of these models.
+We apply our grid-search algorithm once more, using the same specifications we applied in the case of marijuana legalization: 100,000 draws with $R \in [0, 1]$, $S \in [0, 2]$, and $D \in [0, 1]$, with reliability score ranging between 60% and 100%, each having uniform priors.
+Once again, we kept 2% of the DGPs.
+Figure 6 shows the pair plots, while Figure 7 shows the conditional estimates.
 
@@
 ```
-```{r figure4, out.width = "100%"}
+```{r figure6, out.width = "100%"}
 knitr::include_graphics("../figures/abany02.png")
@@
 \end{center}
-\caption*{Figure 4: Fake-Data Simulations for Abortion}
-\footnotesize{\textit{Notes:} The figure depicts the extent to which specific parameter combinations approximate the observed data.
-Each point represents our distance measure, calculated as the sum of absolute errors between observed and simulated contingency tables, where darker colors represent smaller distances.
-We used 5\%, 10\%, and 20\% distance cuts for color shades, along with a second filter using quantiles of 0.05, 0.10, and 0.20 at each level.
-The shading thus reflects both absolute and relative accuracy among the simulations.
-The single yellow point represents the parameter combination that best approximates the observed data in the General Social Survey.}
+\caption*{Figure 6: Pair Plots for the Legality of Abortion}
+\footnotesize{\textit{Notes:} The figure depicts the distribution of, and bivariate relationships among, grid-search parameter values from the best-fitting DGPs for the case of the legality of abortion (N = 2,000 DGPs from 100,000 simulations).}
 \end{figure}
@@
 
+```{=latex}
+\begin{figure}[!ht]
+\begin{center}
+```
+```{r figure7, out.width = "100%"}
+knitr::include_graphics("../figures/abany03.png")
+```
+```{=latex}
+\end{center}
+\caption*{Figure 7: Best-Fitting DGPs for the Legality of Abortion}
+\footnotesize{\textit{Notes:} The figure depicts the best fitting 2,000 DGPs (from 100,000 simulations) for the empirical case study.
+Each point represents an accepted draw.
+We rounded the balance parameters to show conditional relationships.}
+\end{figure}
+```
+
+We find strong variability in the plausible generative scenarios.
+Note how, in all cases, the accepted draws are squished towards 0---either via 0 rate of change or via 0 strength of change---indicating a clear preference towards no change in the underlying processes.
+That said, the plausibility bounds also become strongly noisy when the estimated balance is even: accepting scenarios ranging from no change to the now-familiar high change in a low number of people and low change in a high number of people.
+The highest KS distance scores among all the accepted samples is less than 0.01, suggesting that all draws fit reasonably similarly across all scenarios.
+
 This exercise is an appropriate illustration of the limitations of panel studies.
 Based on the observed data, we cannot determine whether the observed marginals result from a small number of people making substantial belief changes or a large number of people making minor political changes.
-This, ultimately, boils down to the issue of observation frequency: with three-wave panel data and reasonable doubts about the accuracy of every single survey response due to reliability and resolution concerns, it is challenging to adjudicate the underlying generative models.
+This may, ultimately, boil down to the issue of observation frequency: with three-wave panel data and reasonable doubts about the accuracy of every single survey response due to reliability and resolution concerns, it is challenging to adjudicate the underlying generative models.
 In the next section, we relax this assumption and extend our method to cases with more than three waves.
 
-## The Case of Generalized Trust
-
-# Study 2: Simulation Studies and Performance Assessment
-
-# Study 3: Longer Panels and Linear Trajectories
-
-Once we relax the three-wave assumption, the immediate problem is that the contingency tables will be sparse and unwieldy.
-Think, for instance, a panel study with 1,000 people conducted over 10 waves.
-Assuming binary outcomes, there are $2^{10} =$ 1,024 potential trajectories---more than the number of individuals in the study.
-That said, having 10 waves instead of 3 is more desirable for one simple reason: while we had to be economical in our method and count every response very seriously, we now have the opportunity to regularize a person's response trajectories, resulting in more reasonable *trends* for each individual that are less vulnerable to reliability issues.
-
-There are multiple approaches for achieving this with panel data.[^5] We use a fixed-effects design and fit individual linear slopes of the outcome on a normalized time trajectory [@rüttenauer2020].
-That is, we simply do a separate regression for each individual and use their observed slope.
-We apply this method to the case of immigration spending.
-
-[^5]: Two obvious alternatives are multilevel mixed-effects models with varying intercepts and slopes [@lersch2023], which allow for shrinkage in response trajectories using partial pooling, and latent growth curve models [@proust-lima2012], which enable us to postulate latent trajectories at the individual level.
-Both approaches are, of course, plausible.
-However, the shrinkage factor in the multilevel models is far too consequential in the simulation approach when the underlying generative model has too few respondents changing, as it would place excessive emphasis on the population mean.
-Similarly, latent modeling with simulated grid cells results in too many researcher degrees of freedom.
-Therefore, we settled on the fixed effects individual model as a reasonable approach to linear trend modeling.
-
-## The Case of Immigration Spending
+## Exploring the Case of Immigration Spending
 
@@
 Although the sample for this case study is obviously not representative of any national population, it provides an intensive set of repeated observations at the individual level.
-
-We follow the same grid-search procedure, extending our approach to cases involving linear slopes.
-Figure 1 provides a schematic example of our method.
-First, we fit linear models to the observed data, normalizing time to be between 0 and 1, and retrieve individual slopes of change.
-The top-left panel in Figure 1 documents the distribution of these slopes across 213 individuals.
-Next, we build a simulated dataset and perform the same procedure to extract individual slopes.
-The bottom-left panel in Figure 1 shows one such instantiation, where we build a dataset with a fixed rate of change of 25%, a strength of change of 2 SD, and a 50%--50% split in directional change, with individual slopes for 213 simulated respondents.
-In the final step, we compare these two distributions using a two-sample Kolmogorov-Smirnov (KS) statistic to quantify the distance between the observed data and the simulated data from one generated model.
-The right panel in Figure 1 documents the empirical CDFs of these two distributions, which we use in our calculations of the KS statistics.
+Figure 8 shows the trajectory of average responses across these 17 waves within the study window.
 
@@
 ```
-```{r figure5, out.width = "100%"}
+```{r figure8, out.width = "100%"}
 knitr::include_graphics("../figures/immigration01.png")
@@
 \end{center}
-\caption*{Figure 5: The Distribution of Linear Trajectories}
-\footnotesize{\textit{Notes:} The figure shows linear trajectories from the first 17-waves of the \textit{Political Psychology Data from a 26-Wave Yearlong Longitudinal Study} (26W), covering the pre-COVID period of the study.
-The top-left panel depicts the distribution of individual slopes from varying fixed effects models, where a set of respondents agreed---or not---with the statement, "federal spending to control immigration should be increased." We included respondents who participated in all 17 waves (N = 213).
-The bottom-left panel depicts the distribution of individual slopes from a simulated dataset, with a fixed rate of change of 25\%, a strength of change of 2 SD, and a 50\%--50\% split in directional change.
-The right panel shows the empirical cumulative distribution functions from each of these distributions.}
+\caption*{Figure 8: The Percent Support for an Increase in the Control of Immigration Spending}
+\footnotesize{\textit{Notes:} The figure depicts the percent support for the statement "the federal spending to control immigration should be increased" in the Political Psychology Data from a 26-Wave Yearlong Longitudinal Study.}
 \end{figure}
@@
 
-Figure 6 presents the results for this procedure, implemented across, once again, 2,100,000 datasets with 2,100 generative models.
-As can be seen in the figure, there is a clear cluster of results in a specific region of the grid.
-KS statistics range from 4.9% to 47.6%, with an interquartile range of 11% and 18.2%.
-The plausibility bounds indicate that between 50% to 80% of respondents may have made a change of roughly 0.5 SD, resulting in a mean decline in support for increased immigration control spending.
-Consistent with our baseline expectation of a mean decline, the plausible scenarios involve true changers making their changes in a negative direction.
+We follow the same grid-search procedure, ending up with significantly higher KS distance scores.
+While the median KS score was 0.015 for marijuana legalization and 0.006 for the legality of abortion, we now end up with a median KS distance of 0.063---a significantly worse fit to the change scores.
+In fact, even the "best" DGP retrieved from the algorithm had a KS distance of 0.02, suggesting that the algorithm simply cannot converge to a reasonably close fit in the DGPs.
 
@@
 ```
-```{r figure6, out.width = "100%"}
+```{r figure9, out.width = "100%"}
 knitr::include_graphics("../figures/immigration02.png")
@@
 \end{center}
-\caption*{Figure 6: Fake-Data Simulations for Immigration}
-\footnotesize{\textit{Notes:} The figure depicts the extent to which specific parameter combinations approximate the observed data.
-Each point represents our distance measure, calculated as the Kilmogorov-Smirnov statistics of the empirical CDFs of observed and simulated slopes, where darker colors represent smaller distances.
-We used 5\%, 10\%, and 20\% distance cuts for color shades, along with a second filter using quantiles of 0.05, 0.10, and 0.20 at each level.
-The shading thus reflects both absolute and relative accuracy among the simulations.
-The single yellow point represents the parameter combination that best approximates the observed data in the 26W Study.}
+\caption*{Figure 9: Pair Plots for Immigration Spending}
+\footnotesize{\textit{Notes:} The figure depicts the distribution of, and bivariate relationships among, grid-search parameter values from the best-fitting DGPs for the case of immigration spending (N = 2,000 DGPs from 100,000 simulations).}
 \end{figure}
@@
 
-With this exercise, we extended our grid-search procedure to cases with more than three waves of panel data, but the same procedure is applicable to the three-wave case as well, provided that the analyst believes it is appropriate to use linear slopes estimated from three observations.
-The underlying logic, however, is the same: as long as we have comparable trajectories in the simulated data relative to our observed empirical data, we can scale up the process and perform algorithmic searches.[^6]
+```{=latex}
+\begin{figure}[!ht]
+\begin{center}
+```
+```{r figure10, out.width = "100%"}
+knitr::include_graphics("../figures/immigration03.png")
+```
+```{=latex}
+\end{center}
+\caption*{Figure 10: Best-Fitting DGPs for Immigration Spending}
+\footnotesize{\textit{Notes:} The figure depicts the best fitting 2,000 DGPs (from 100,000 simulations) for the empirical case study.
+Each point represents an accepted draw.
+We rounded the balance parameters to show conditional relationships.}
+\end{figure}
+```
 
-[^6]: Of course, one such procedure may involve a lazy algorithm that goes through a targeted search, where each iteration involves more focused and narrow bands of parameters.
-This is vulnerable to get stuck in local optima, however.
+Once again, Figure 9 presents the pair plots, while Figure 10 depicts the conditional DGP estimates.
+Consistent with the decline in the observed marginals, we find samples in the regions with negative balance (a median of 0.17 with an interquartile range between 0.09 and 0.27).
+We also find similar IQR values of 0.60 for strength and 0.39 for rate, suggesting plausibility bounds in our conditional estimates.
+In contrast to our previous examples, however, these plausibility bounds have much more noise, as seen in Figure 10.
+A tighter acceptance rate reduces the spread, but the main fit still suffers from high KS distance values, suggesting that the data has many plausible DGPs.
 
-# Study 4: Detection of True Changers
+\clearpage
 
-So far, we have presented simulation evidence for recovering plausible generative models (DGPs) that may have produced the observed empirical data.
-But what about *classifying individuals* into those who change and those who do not?
-Suppose that, using the grid-search procedure, we happen to determine the true generative model for our empirical observations.
-Can we also recover an accurate classification of changers and non-changers at the individual level?
-That is, given a vector of predictions, can we measure our success in identifying the true changers [@tharwat2020]?
+# Simulation Studies and Validation Checks
 
-To address this question, we conduct another simulation exercise.
-Similar to earlier simulations, we generate $N$ datasets, this time using parameters fixed to the best-case values identified through our grid-search procedure.
-In each of these datasets, we estimate linear regressions at the individual level with time scaled from 0 to 1---effectively allowing us to extract a model-predicted measure of change for each person.
-We use a straightforward heuristic to decide whether a simulated person changed: if the linear slope is closer to the strength-of-change parameter than it is to 0, the respondent gets classified as a *changer*, otherwise, they are classified as a *non-changer*.
-We then cross-tabulate these predictions with the real change indicator used during the generation of the data, constructing a "confusion matrix" that compares true change classifications with observed change classifications.
+\clearpage
 
-In the binary, three-wave case, these classifications are straightforward.
-Only the 001, 011, 100, and 110 cells can be classified as changers because only they can have non-zero slopes---in fact, all have slopes of 1 or -1.
-In cases where the most plausible DGP is one directional, like for the marijuana question, only two of those cells would be classified as changers.
-The situation is more complicated for a more continuous case, like that of immigration in the 26-wave data.
-Many different slopes are possible.
-In those cases, classifying people as closer to "changers" or "non-changers" relies on choosing a cutoff, as we do here.
-
-In either case, this binary classification allows us to assess how "accurate" we are in a single simulation run.
-Of course, there are multiple ways to conceptualize accuracy, but for our purposes, we focus on two: *sensitivity* and *specificity*.
-The former measures the probability that a true changer will get classified as a changer; the latter is the probability that a true non-changer will be classified as a non-changer.
-These are the formulas for each:
-
-$$
-\begin{aligned}
-\text{Sensitivity} &= \frac{\text{True Positive}}{\text{True Positive + False Negative}} & \text{Specificity} &= \frac{\text{True Negative}}{\text{True Negative + False Positive}}
-\end{aligned}
-$$
-
-We extract the best-case scenarios from our three empirical case studies---marijuana legalization, the legality of abortion, and immigration spending---and perform the classification task across 1,000 fake-data simulations.
-We then calculate the sensitivity and specificity scores in each run.
-
-A concrete example will help clarify what these numbers mean.
-The mean sensitivity for the abortion question is .205.
-This means that only 21% of those who actually changed would be correctly classified as changers, even with full knowledge of the data generating process.
-This is due to several factors, including the shortness of the panel, imperfect reliability---though .86 is quite good---and the low level of response resolution, which is binary.
-
-For sensitivity, the number is higher at .828. This means that 83% of the non-changers are correctly classified. This is largely due to the marginal distribution of change.
-
-The distributions of sensitivity and specificity for all three analyses are presented in Figure 7.
-It is clear that our general ability to identify true changers is quite low, whereas the identification of true non-changers is rather better.
-Across simulations, sensitivity ranges between 17% and 54%, while specificity scores range between 74% and 95%.
-That said, there are also subtle differences among the case studies.
-Sensitivity scores are consistently lower in the legality of abortion case, reflecting our earlier discussion about how the plausible set of scenarios was much higher in that case.
-At the same time, the spread of simulation scores is rather high in the case of immigration spending for both sensitivity and specificity, suggesting that linear trajectories in longer panels might be rather susceptible to randomness in our analyses.
-
-These results suggest that, while the grid-search procedure allows us to identify plausible generative models that may have generated the empirical patterns, detecting true change at the level of individuals occurs under only very narrow conditions.
-The fact that linear trends with longer panels performed better, on average, than the three-wave cases indicates that having longer panels might be essential to dampen the negative effects of response reliability on our estimates.
-This finding emphasizes the importance of collecting panel data over more extended periods to increase the robustness of our proposed DGPs, and the accuracy of identifying true changers in the population.
+# The Detection of True Changers
 
@@
 
+Once again, we emphasize that this identification problem cannot be "solved" with simple methodological considerations.
+As we noted before, our objective was not to eliminate the problem, but rather to provide a concrete approach that can help us understand the issue more clearly.
+More substantively, we need to acknowledge the problem in our empirical applications and be willing to *impose* certain assumptions.
+The fact that multiple, sometimes contradictory, generative processes may very well account for the same data means that researchers should assert theoretical positions when it comes to connecting their hypotheses to empirical data.
+
 This article has several implications for the study of political belief change across the life-course.
-First, our findings suggest that recent debates about the existence or non-existence of belief change in adult populations [@kiley2020; @lersch2023] should tackle the heterogeneous nature of change and stability in individual preferences since group-level measures may hamper our understanding of individual mechanisms.
+First, our findings suggest that recent debates about the existence or non-existence of belief change in adult populations [@kiley2020; @lersch2023] should tackle the heterogeneous nature of change and stability in individual preferences because group-level measures might hamper our understanding of individual mechanisms.
 Instead, scholars need to specify the generative models that aggregate average estimates imply.
@@
 
-Second, the ambivalence in response trajectories and the observation that linear trend models may perform better with longer panels indicate that classical solutions to the measurement problem---ranging from scale construction to structural equation models and principal component analyses [@judd1980; @ansolabehere2008; @alwin1991]---should have an important place in panel designs, as even moderate reliability scores can significantly hinder our ability to specify plausible generative processes.
-We highlighted that a fundamental challenge in the study of belief change is distinguishing whether an observed change is due to true change or measurement error, and reliability and resolution are of utmost importance to resolve this problem.
+Second, the ambivalence in response trajectories indicate that classical solutions to the measurement problem---ranging from scale construction to structural equation models and principal component analyses [@judd1980; @ansolabehere2008; @alwin1991]---should have an important place in panel designs, as even moderate reliability scores can significantly hinder our ability to specify plausible generative processes.
+We note that a fundamental challenge in the study of belief change is distinguishing whether an observed change is due to true change or error, and reliability and resolution are of utmost importance to resolve this problem.
 
@@
 
-Third, our finding that individual classifications of change status are largely ineffective means that future research needs to be more sensitive to the detection of trends in individual survey responses over time.
-As mentioned earlier, varying modeling strategies, ranging from multilevel mixed-effects models that rely on partial pooling to various forms of latent mixture models, could improve the reliability of responses at the individual level.
-We believe that the use of these modeling strategies, along with longer panel designs, could provide important corrections to the study of belief change.
-
-Finally, these results call for granular panel surveys that go beyond the classical 2-wave and 3-wave designs, as reliability concerns significantly limit the plausible conclusions we can derive from these shorter panels.
+Finally, these results call for granular panel surveys that go beyond the classical 2-wave and 3-wave designs because reliability concerns significantly limit the plausible conclusions we may derive from these shorter panels.
 In a 2-wave setup, where there is only one change, we can never properly establish whether a change we observe is due to measurement error or it is indeed a genuine shift, making extended panel designs necessary.
-The differences among our case studies, along with the sensitivity and specificity tests, suggest that panel studies with shorter designs are largely not suited for trend modeling.
 Since we care about generative models rather than one observed sample, strategies that enable us to reliably estimate target quantities are warranted and essential to make progress on longstanding questions about belief change.

### 2025-01-12 scripts/1_ksdist.R

@@ -109,7 +109,7 @@ example_p1 <- data_A |>
     alpha = 0.8
   ) +
   theme_ipsum_rc(grid = "XxY", axis_title_size = 12) +
-  labs(subtitle = "Slopes from DGP 1", 
+  labs(subtitle = "Change Scores from DGP 1", 
        x = "", 
        y = "Relative Frequency") +
   scale_x_continuous(
@@ -129,8 +129,8 @@ example_p2 <- data_B |>
     alpha = 0.8
   ) +
   theme_ipsum_rc(grid = "XxY", axis_title_size = 12) +
-  labs(subtitle = "Slopes from DGP 2", 
-       x = "Individual-Level Slope Coefficients", 
+  labs(subtitle = "Change Scores from DGP 2", 
+       x = "Change Scores", 
        y = "") +
   scale_x_continuous(
     limits = c(-1.15, 1.15),
@@ -147,7 +147,7 @@ example_p3 <-
   stat_ecdf(aes(x = dgp2), geom = "step", col = my_oka[3]) +
   theme_ipsum_rc(grid = "XxY", axis_title_size = 12) +
   labs(subtitle = "Empirical CDF",
-       x = "Individual-Level Slope Coefficients", 
+       x = "Change Scores", 
        y = "ECDF") +
   scale_x_continuous(breaks = scales::pretty_breaks()) +
   scale_y_continuous(breaks = scales::pretty_breaks())

### 2025-01-12 scripts/4_trust.R

@@ -1,159 +1,0 @@
-
-# - promises and pitfalls of panel change ------------------------------------ #
-# - case 3 # trust ----------------------------------------------------------- #
-
-### note: this script
-###       (a) takes `trust` as an examplary BES item,
-###       (b) simulates potential DGPs approximating the observed parameters,
-###       (c) shows the relative fit of the DGPs to observed data.
-
-rm(list = ls()) # clean-up
-
-# install.packages("devtools")
-# devtools::install_github("tkeskinturk/gridsearch")
-
-pacman::p_load(
-  srvyr,
-  gridsearch,
-  tidyverse,
-  hrbrthemes,
-  GGally)
-theme_set(theme_ipsum_rc())
-
-# --- helpers
-my_oka <-
-  c("#eebd64",
-    "#8b3b36", 
-    "#a2b4b7")
-`%nin%` = Negate(`%in%`)
-
-# PART 1: BES DATA ----------------------------------------------------------- #
-
-# --- get data
-data <-
-  haven::read_dta("./data/source_files/bes.dta") |>
-  select(
-    pid = id,
-    trust_1 = genTrustW17,
-    trust_2 = genTrustW20,
-    trust_3 = genTrustW21,
-    trust_4 = genTrustW23,
-    trust_5 = genTrustW27
-  ) |>
-  haven::zap_labels() |>
-  mutate_at(
-    .vars = vars(matches("trust")),
-    .funs = ~ case_when(. == 1 ~ 1, . == 2 ~ 0, TRUE ~ NA_real_)
-  ) |>
-  drop_na() |>
-  pivot_longer(cols = -pid,
-               names_to = "var",
-               values_to = "y") |>
-  separate(var, into = c("var", "t")) |>
-  mutate(t = as.integer(t)) |> select(-var) |> 
-  mutate(
-    year =
-      case_when(
-        t == 1 ~ 2019,
-        t == 2 ~ 2020,
-        t == 3 ~ 2021,
-        t == 4 ~ 2022,
-        t == 5 ~ 2024
-      )
-  )
-
-# --- draw trajectories
-png(
-  "figures/trust01.png",
-  w = 7,
-  h = 5,
-  units = "in",
-  res = 500
-)
-glm(y ~ factor(year), data = data, family = "binomial") |>
-  ggeffects::ggemmeans(terms = "year") |>
-  as_tibble() |>
-  ## draw
-  ggplot(aes(
-    x = x,
-    y = predicted
-  )) +
-  ## statistics
[…]

### 2025-01-12 scripts/5_immig.R

@@ -1,141 +1,0 @@
-
-# - promises and pitfalls of panel change ------------------------------------ #
-# - case 4 # immigration ----------------------------------------------------- #
-
-### note: this script
-###       (a) takes `imm` as an examplary panel data item,
-###       (b) simulates potential DGPs approximating the observed parameters,
-###       (c) shows the relative fit of the DGPs to observed data.
-
-rm(list = ls()) # clean-up
-
-# install.packages("devtools")
-# devtools::install_github("tkeskinturk/gridsearch")
-
-pacman::p_load(
-  srvyr,
-  gridsearch,
-  tidyverse,
-  hrbrthemes,
-  GGally)
-theme_set(theme_ipsum_rc())
-
-# --- helpers
-my_oka <-
-  c("#eebd64",
-    "#8b3b36", 
-    "#a2b4b7")
-`%nin%` = Negate(`%in%`)
-
-# PART 1: 26-WAVE DATA ------------------------------------------------------- #
-
-# --- get data
-data <- 
-  read_csv("./data/source_files/yllanon.csv") |> 
-  ## organize
-  mutate(pid = cur_group_id(), .by = "id") |>
-  ## chop off post-COVID periods
-  filter(wave < 18) |>
-  ## simplify
-  select(pid, t = wave, y = imm) |>
-  ## clean-up the outcome
-  filter(y %nin% c(8, 9)) |> drop_na() |>
-  ## fully balanced
-  mutate(n = n(), .by = "pid") |> filter(n == 17) |>
-  ## scale the variables
-  mutate(y = ifelse(y > 4, 1, 0)) |>
-  ## select and wrap up organization
-  select(pid, t, y) |> arrange(pid, t)
-
-# --- draw trajectories
-png(
-  "figures/immigration01.png",
-  w = 7,
-  h = 5,
-  units = "in",
-  res = 500
-)
-data |>
-  group_by(t) |>
-  summarize(y = mean(y)) |>
-  ## draw
-  ggplot(aes(x = t, y = y)) +
-  ## statistics
-  geom_line(linetype = "dashed", linewidth = 0.2) +
-  geom_point(size = 2) +
-  ## labels
-  labs(x = "Time", y = "Immigration Spending Preference") +
-  scale_x_continuous(breaks = scales::pretty_breaks(n = 17)) +
-  scale_y_continuous(
-    limits = c(0.35, 0.55),
-    labels = scales::label_percent()) +
-  ## structure
[…]

### 2025-01-12 scripts/backups/confusionMat.R

@@ -1,73 +1,0 @@
-
-# - `gridsearch` ------------------------------------------------------------- #
-# - function: confusion matrix ----------------------------------------------- #
-
-### note: this function
-###       (a) takes a confusion matrix of true/simulated counts as its input,
-###       (b) calculates kappa, sensitivity, and specificity.
-
-# PART 1: FUNCTION ----------------------------------------------------------- #
-
-#' Calculate Accuracy Metrics
-#'
-#' This function takes a confusion matrix as its input and calculates sensitivity, specificity, and kappa scores for classification.
-#' @param confusion_matrix A confusion matrix.
-#'
-#' @return A data frame.
-#' @noRd
-#' @keywords internal
-#'
-
-confusionMatrix <-
-
-  function(confusion_matrix) {
-
-    # --- calculate kappa
-
-    # accuracy
-    p_0 =
-      (confusion_matrix[4, 3]
-       +
-       confusion_matrix[1, 3]
-      ) /
-      sum(confusion_matrix$n)
-
-    # expected classification
-    p_e =
-      (
-        (
-        (confusion_matrix[1, 3] + confusion_matrix[2, 3])
-        *
-        (confusion_matrix[1, 3] + confusion_matrix[3, 3])
-        )
-      +
-        (
-        (confusion_matrix[3, 3] + confusion_matrix[4, 3])
-        *
-        (confusion_matrix[2, 3] + confusion_matrix[4, 3])
-        )
-      ) /
-      (sum(confusion_matrix$n) ^ 2)
-
-    # kappa score
-    kappa = (p_0 - p_e) / (1 - p_e)
-
-    # --- calculate sensitivity
-    sensitivity =
-      confusion_matrix[4, 3] /
-      (confusion_matrix[4, 3] + confusion_matrix[3, 3])
-
-    # --- calculate specificity
-    specificity =
-      confusion_matrix[1, 3] /
-      (confusion_matrix[1, 3] + confusion_matrix[2, 3])
-
-    # --- export
-    output <- list(kappa = kappa$n,
-                   sensitivity = sensitivity$n,
-                   specificity = specificity$n)
-
-    return(output)
-  }
-
-# ---------------------------------------------------------------------------- #

### 2025-01-12 scripts/backups/simulateChangers.R

@@ -1,189 +1,0 @@
-
-# - `gridsearch` ------------------------------------------------------------- #
-# - function: simulate changers ---------------------------------------------- #
-
-### note: this function
-###       (a) takes a set of parameters from a specific DGP and simulates data,
-###       (b) calculates accuracy scores for detecting changers.
-
-# FUNCTION CALL -------------------------------------------------------------- #
-
-#' Simulate Changers and Calculate Accuracy
-#'
-#' This function uses parameter values from a given DGP to calculate kappa, sensitivity, and specificity scores in simulated data frames.
-#' @param n The number of individuals in the DGP.
-#' @param t The number of time periods.
-#' @param rate The percent of individuals changing across the panel.
-#' @param balance_dir The direction of change, where 0 codes negative change, 1 codes positive change, and all values in-between codes the percentage of changers changing in the positive direction.
-#' @param balance_res The marginal distribution of the outcome (effectively regulating the percent distribution of 0s and 1s).
-#' @param strength The strength of change in the latent variable.
-#' @param reliable The reliability score of the outcome measurement.
-#' @param nrep The number of simulation runs.
-#' @param seed Seed for reproducibility.
-#' @param workers The number of workers for parallelization (note that parallelization is highly recommended for reasonable duration).
-#'
-#' @return A data frame.
-#'
-#' @details # Examples
-#'
-#' **Basic [simulateChangers()] call:**
-#' ```{r, comment = "#>", collapse = TRUE}
-#' set.seed(11235)
-#' data <- simulateChangers(n = 100,
-#'                          t = 3,
-#'                          rate = 0.5,
-#'                          balance_dir = 1,
-#'                          balance_res = 0.5,
-#'                          strength = 1,
-#'                          reliable = 0.9,
-#'                          nrep = 100,
-#'                          seed = 11234,
-#'                          workers = 10)
-#' print(data)
-#' ```
-#'
-#' @export
-#'
-
-simulateChangers <-
-
-  function(
-
-    # ----------------------------------------------------- #
-    # FUNCTION ARGUMENTS AND WARNINGS                       #
-    # ----------------------------------------------------- #
-
-    n = 1000,
-    t = 3,
-    rate = 0.25,
-    balance_dir = .5,
-    balance_res = .5,
-    strength = .5,
-    reliable = 1,
-    nrep = 1000, seed = 11235, workers = 8
-
-  ) {
-
[…]

### 2025-01-12 scripts/backups/simulate_changers.R

@@ -1,0 +1,189 @@
+
+# - `gridsearch` ------------------------------------------------------------- #
+# - function: simulate changers ---------------------------------------------- #
+
+### note: this function
+###       (a) takes a set of parameters from a specific DGP and simulates data,
+###       (b) calculates accuracy scores for detecting changers.
+
+# FUNCTION CALL -------------------------------------------------------------- #
+
+#' Simulate Changers and Calculate Accuracy
+#'
+#' This function uses parameter values from a given DGP to calculate kappa, sensitivity, and specificity scores in simulated data frames.
+#' @param n The number of individuals in the DGP.
+#' @param t The number of time periods.
+#' @param rate The percent of individuals changing across the panel.
+#' @param balance_dir The direction of change, where 0 codes negative change, 1 codes positive change, and all values in-between codes the percentage of changers changing in the positive direction.
+#' @param balance_res The marginal distribution of the outcome (effectively regulating the percent distribution of 0s and 1s).
+#' @param strength The strength of change in the latent variable.
+#' @param reliable The reliability score of the outcome measurement.
+#' @param nrep The number of simulation runs.
+#' @param seed Seed for reproducibility.
+#' @param workers The number of workers for parallelization (note that parallelization is highly recommended for reasonable duration).
+#'
+#' @return A data frame.
+#'
+#' @details # Examples
+#'
+#' **Basic [simulateChangers()] call:**
+#' ```{r, comment = "#>", collapse = TRUE}
+#' set.seed(11235)
+#' data <- simulateChangers(n = 100,
+#'                          t = 3,
+#'                          rate = 0.5,
+#'                          balance_dir = 1,
+#'                          balance_res = 0.5,
+#'                          strength = 1,
+#'                          reliable = 0.9,
+#'                          nrep = 100,
+#'                          seed = 11234,
+#'                          workers = 10)
+#' print(data)
+#' ```
+#'
+#' @export
+#'
+
+simulateChangers <-
+
+  function(
+
+    # ----------------------------------------------------- #
+    # FUNCTION ARGUMENTS AND WARNINGS                       #
+    # ----------------------------------------------------- #
+
+    n = 1000,
+    t = 3,
+    rate = 0.25,
+    balance_dir = .5,
+    balance_res = .5,
+    strength = .5,
+    reliable = 1,
+    nrep = 1000, seed = 11235, workers = 8
+
+  ) {
+
[…]

### 2025-01-12 scripts/backups/simulation_DGPs.R

@@ -1,0 +1,213 @@
+
+# - promises and pitfalls of panel change ------------------------------------ #
+# - script 1 # parameter space for `gridSearch` ------------------------------ #
+
+### note: this script
+###       (a) generates simulated DGPs with varying parameters,
+###       (b) applies the grid search algorithm to recover error distributions,
+###       (c) evaluates whether the algorithm can recover "true" DGP.
+
+rm(list = ls()) # clean-up
+
+# install.packages("devtools")
+# devtools::install_github("tkeskinturk/gridsearch")
+
+pacman::p_load(
+  purrr,
+  furrr,
+  hrbrthemes,
+  gridsearch,
+  ggforce,
+  tidyverse,
+  broom.helpers)
+theme_set(theme_ipsum_rc())
+
+# --- helpers
+my_oka <-
+  c("#eebd64",
+    "#8b3b36", 
+    "#a2b4b7")
+`%nin%` = Negate(`%in%`)
+
+# PART 1: DATA --------------------------------------------------------------- #
+
+set.seed(11235)
+
+# --- conditions
+d <-
+  expand_grid(
+    ## 500 individuals
+    p_n = 500,
+    ## varying panel waves (restrict to contingency)
+    p_t = c(3, 5, 7),
+    ## the rate of change in the population
+    p_rate = c(.1, .25, .5),
+    ## the direction of change (halfsies versus secular)
+    p_dir = c(.5, 1),
+    ## the fixed level of marginal distributions
+    p_res = c(.5, .75),
+    ## the strength of change in latent variable
+    p_strength = c(.5, 1),
+    ## the reliability score of the outcome measurement
+    p_reliable = .9,
+    ## the number of simulated datasets
+    sims = 1:30)
+
+# --- dataframes
+d <- d |>
+  dplyr::mutate(
+    data =
+      furrr::future_pmap(
+        ## mapping list
+        .l = list(p_n,
+                  p_t,
+                  p_rate,
+                  p_dir,
+                  p_res,
+                  p_strength,
+                  p_reliable),
+        ## refer to list based on index
+        .f = purrr::possibly(
+          ~ buildDGP(
+            # varying parameters
+            n = ..1,
+            t = ..2,
+            rate = ..3,
+            balance_dir = ..4,
+            balance_res = ..5,
+            strength = ..6,
+            reliable = ..7,
+            export = TRUE,
[…]

### 2025-01-12 scripts/backups/simulations_reliability.R

@@ -1,246 +1,0 @@
-
-# - promises and pitfalls of panel change ------------------------------------ #
-# - script 2 # reliability space for `gridSearch` ---------------------------- #
-
-### note: this script
-###       (a) generates simulated DGPs with varying reliability scores,
-###       (b) applies the grid search algorithm to recover error distributions,
-###       (c) evaluates whether the algorithm can recover "true" DGP.
-
-rm(list = ls()) # clean-up
-
-# install.packages("devtools")
-# devtools::install_github("tkeskinturk/gridsearch")
-
-pacman::p_load(
-  purrr,
-  furrr,
-  hrbrthemes,
-  gridsearch,
-  ggforce,
-  tidyverse,
-  marginaleffects)
-theme_set(theme_ipsum_rc())
-
-# --- helpers
-my_oka <-
-  c("#eebd64",
-    "#8b3b36", 
-    "#a2b4b7")
-`%nin%` = Negate(`%in%`)
-
-# PART 1: DATA --------------------------------------------------------------- #
-
-set.seed(11235)
-
-# --- conditions
-d <-
-  expand_grid(
-    ## 500 individuals
-    p_n = 500,
-    ## varying panel waves (restrict to contingency)
-    p_t = 3,
-    ## the rate of change in the population
-    p_rate = c(.1, .25, .5),
-    ## the direction of change (halfsies versus secular)
-    p_dir = c(.5, 1),
-    ## the fixed level of marginal distributions
-    p_res = .5,
-    ## the strength of change in latent variable
-    p_strength = c(.5, 1),
-    ## the reliability score of the outcome measurement
-    p_reliable = c(.7, .9),
-    ## the number of simulated datasets
-    sims = 1:30)
-
-# --- dataframes
-d <- d |>
-  dplyr::mutate(
-    data =
-      furrr::future_pmap(
-        ## mapping list
-        .l = list(p_n,
-                  p_t,
-                  p_rate,
-                  p_dir,
-                  p_res,
-                  p_strength,
-                  p_reliable),
-        ## refer to list based on index
-        .f = purrr::possibly(
-          ~ buildDGP(
-            # varying parameters
-            n = ..1,
-            t = ..2,
-            rate = ..3,
-            balance_dir = ..4,
-            balance_res = ..5,
-            strength = ..6,
[…]
